### 2016-04-06 swsc.tex
Initial ShareLaTeX Import

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                
%%      SWSC LaTeX class for Journal of Space Weather and Space Climate
%%      
%%                                      (c) Springer-Verlag HD
%%                                      revised by EDP Sciences
%%                                      further revised by J. Watermann 
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%      This demonstration file was derived from aa.dem
%%  
%%      AA vers. 7.0, LaTeX class for Astronomy & Astrophysics
%%      demonstration file
%%                                                (c) Springer-Verlag HD
%%                                                revised by EDP Sciences
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%      modified for Journal of Space Weather and Space Climate
%%      by Jurgen Watermann, Editorial Advisor to SWSC
%%
%%      01-04-2012
%%      02-04-2012 revision 1
%%      12-07-2012 revision 2
%%      06-12-2012 revision 3 
%%      01-01-2014 revision 4
%%      06-03-2014 revision 4.1
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%      The two sub-figures referenced in this template are of eps and png type,
%%      respectively, in order to demonstrate the usepackages subfigure and
%%      epstopdf and thus create pdf-only output 
%%
%%      If you want to use TexLive or MikTex together with a bibtex bibliography 
%%      file you may run Latex2e from the command line 
%%          pdflatex -shell-escape swsc.tex
%%          bibtex swsc (do not include an extension such as .tex or .bib)
%%          pdflatex -shell-escape swsc.tex
%%          pdflatex -shell-escape swsc.tex
%%
%%      A double call to pdflatex after calling bibtex is necessary in order to
%%      set citations and references correctly and insure that foreward/backward  
%%      linkage (backref option) is properly applied
%%      If you use MikTex you may need to make a triple call to pdflatex
%%
%%      If you are using TexLive or MikTex but not a bibtex type of bibliography
%%      you may simply run Latex2e twice from the command line 
%%          pdflatex -shell-escape swsc.tex
%%          pdflatex -shell-escape swsc.tex
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%   single column 12-point version for review
%%

%%  with traditional abstract
\documentclass[referee,a4paper,12pt,traditabstract]{swsc}

%%  with structured abstract 
%\documentclass[referee,a4paper,12pt,structabstract]{swsc}

\usepackage{graphicx}
\usepackage{txfonts}
\usepackage{subfigure}
\usepackage{epstopdf}
\usepackage{lineno}
\usepackage[authoryear,round]{natbib}
\usepackage[backref]{hyperref}
\usepackage{url}
\usepackage[inline]{enumitem}
%%    This version assumes using bibtex with the swsc bibliography style file
\bibliographystyle{swsc.bst}

\hypersetup{colorlinks=true,citecolor=cyan,urlcolor=cyan,linkcolor=blue}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{linenumbers}

   \title{Gaussian Processes Autoregressive Models for the Disturbance Storm Time Index}

   %\subtitle{I. Overviewing the $\kappa$-mechanism}

   \titlerunning{Gaussian Process $D_{st}$ models}

   \authorrunning{Chandorkar and Camporeale}

   \author{M. Chandorkar
          \inst{1}
          \and
          E. Camporeale\inst{1}
          }

   \institute{Centrum Wiskunde Informatica (CWI), Amsterdam,
              1098XG Amsterdam\\
              \email{\href{mailto:m.h.chandorkar@cwi.nl}{m.h.chandorkar@cwi.nl} \href{mailto:e.camporeale@cwi.nl}{e.camporeale@cwi.nl}} 
             }

%%   \date{Received September 15, 1996; accepted March 16, 1997}

  % \abstract{}{}{}{}{}        %% uncomment if structured abstract is desired
 %% 5 {} token are mandatory

   %% replace by pair of curly brackets, {}, if structured abstract is selected

   \abstract{
    We aim to improve the state of the art of \emph{One Step Ahead}(OSA) prediction of the $D_{st}$ geomagnetic activity index. In the domain of OSA prediction, it is well known that $D_{st}(t)$ values correlate strongly with values one hour in the past (i.e. $D_{st}(t-1)$). While a number of forecasting models have been proposed and compared, none of them have been compared with the \emph{Persistence} model to quantify their performance improvements above this most trivial of predictors. In this study we train Gaussian Process regression models for One Step Ahead (OSE) prediction of the Disturbance Storm Time ($D_{st}$) geomagnetic index. We propose three variants of the Gaussian Process model.
    \begin{enumerate*}
      \item Gaussian Process Nonlinear Auto regressive GP-NAR 
      \item Gaussian Process Auto regressive with eXogenous inputs GP-NARX. 
    \end{enumerate*}
    We compare the performance of these models with the current state of the art in one step ahead $D_{st}$ prediction on a set of 
    63 benchmark storms from 1998-2006. The proposed Gaussian Process models not only outperform the rest but also the \emph{Persistence} model.
    }

   \keywords{Geomagnetic indices --
            $D_{st}$ OSA Prediction --
            Gaussian Processes --
            Machine Learning
            }

   \maketitle
%%
%%________________________________________________________________

\section{Introduction}

We use the hourly resolution Omni data set (see \cite{OmniPaper}) as extracted from NASA/GSFC's OMNI data set through OMNIWeb.

\section{Methodology: Gaussian Process}

We restrict ourselves to the supervised learning problem in statistics, which is inductive. In other words, we aim to model functions of single or multiple variables (i.e. fields $f(\mathbf{x}), \ x \in \mathbb{R}^d $), using a set of labeled data points ${(\mathbf{x}_i, y_i), \forall i \in 1 \cdots N}$. These problems have been studied extensively in statistics over the past century.

\emph{Stochastic Processes} extend the idea of probability distributions to the space of functions of a single or multiple variables. A \emph{Gaussian Process} is a stochastic process whose finite dimensional distributions are multivariate Gaussian \ref{eq:sto}.
\vspace{-\baselineskip}
\begin{eqnarray}
 \left(f(\mathbf{x}_1), f(\mathbf{x}_2) \cdots f(\mathbf{x}_N) \right)^T & \sim & \mathcal{N}\left(\mathbf{\mu}, \mathbf{\Lambda} \right) \label{eq:sto} \\
 \mathbf{\Lambda}_{ij} & = & K(\mathbf{x}_i, \mathbf{x}_j) \label{eq:kern}\\
 f & \sim & \mathcal{GP}(m(\mathbf{x}), K(\mathbf{x},\mathbf{x}'))
\end{eqnarray}

The covariance structure of a \emph{stochastic process} is represented by a symmetric positive definite function as given in \ref{eq:kern}, by \emph{Kolmogorov's extension theorem} (see \citet{tao2011introduction}), the function $K: \mathbb{R}^d \times \mathbb{R}^d \rightarrow \mathbb{R}$ uniquely determines the process. Therefore this covariance function (also called kernel function in machine learning) helps us to specify the class of functions represented by the process.

\emph{Gaussian Processes} first appeared in machine learning research as the limiting case of Bayesian inference performed on neural networks with infinitely many neurons in the hidden layers \citet{Neal:1996:BLN:525544}. It was noted by the author that "there may be simpler ways to do inference in this case". Although their inception in the machine learning community is recent, they were studied in statistics/mathematics community as special cases of \emph{Stochastic Processes}. The reader is requested to refer to \cite{Rasmussen:2005:GPM:1162254} for an in depth treatment of Gaussian Processes in Machine Learning.

\begin{eqnarray}
      \mathbf{x} & \in & \mathbb{R}^d \\
      \mathbf{X} & = & (\mathbf{x}_1, \mathbf{x}_2, \cdots, \mathbf{x}_N) \\
      \mathbf{y} & = & (y_1, y_2, \cdots, y_N) \\
      y & = & f(\mathbf{x}) + \epsilon  \\
      f & \sim & \mathcal{GP}(m(\mathbf{x}), K(\mathbf{x},\mathbf{x}'))  \\
      \left(\mathbf{y} \ \ \mathbf{f_*} \right)^\intercal & \sim & 
\mathcal{N}\left(\mathbf{0}, \left[ \begin{array}{ccc} \mathbf{K}(\mathbf{X}, \mathbf{X}) + \sigma^{2} \mathbf{I} & \mathbf{K}(\mathbf{X}, \mathbf{X_*}) \\ \mathbf{K}(\mathbf{X_*}, \mathbf{X}) & \mathbf{K}(\mathbf{X_*}, \mathbf{X_*}) \end{array} \right ] \right) \\
      \mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*} & \sim & \mathcal{N}(\mathbf{\bar{f}_*}, \Sigma_*)  \label{eq:posterior} \\
      \mathbf{\bar{f}_*} & = & \mathbb{E}[\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}]  \\
      & = & K(\mathbf{X_*},\mathbf{X})^\intercal [K(\mathbf{X},\mathbf{X}) + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y} \\
      \Sigma_* & = & K(\mathbf{X_*}, \mathbf{X_*}) - K(\mathbf{X_*},\mathbf{X})^\intercal \left(K(\mathbf{X},\mathbf{X}) + \sigma^{2} \mathbf{I}\right)^{-1}K(\mathbf{X},\mathbf{X_*})
\end{eqnarray}

\subsection{Kernel Functions}

For the success of a Gaussian Process model an appropriate choice of kernel function is paramount. It is important to note that every kernel function generates a \emph{Reproducing Kernel Hilbert Space} (RKHS) of functions, and the nature of the kernel dictates what restrictions exist on the functions belonging to its RKHS. The interested reader is suggested to refer to \cite{Berlinet2004} for a thorough treatment on RKHS and the rich theory behind them.

Some common kernel functions used in machine learning are.

\subsubsection{Radial Basis Function (RBF)}

The RBF kernel is the most commonly used kernel in machine learning, in its core it

\begin{equation}
    K(\mathbf{x}, \mathbf{y}) = \frac{1}{2} exp(-||\mathbf{x} - \mathbf{y}||^2/\sigma^2)
\end{equation}

\subsubsection{Polynomial}

\begin{equation}
    K(\mathbf{x}, \mathbf{y}) = (\mathbf{x}^\intercal \mathbf{y} + b)^d 
\end{equation}

\subsection{Optimizing Hyper-parameters}

\section{One Step Ahead Prediction}

\subsection{Gaussian Process Nonlinear Auto-Regressive (NAR)}

The input vector $x_t$ at each time step is the $p$ order time history of $D_{st}$.
  \begin{eqnarray}
    D_{st}(t) & = & f(x_t) + \epsilon\\
    x_t & = & \left(D_{st}(t-1), D_{st}(t-2), \cdots , D_{st}(t-p)\right) \\
    f(x_t) & \sim & \mathcal{GP}(m(x_t), K(x_t,x_s)) \label{eq:DstGP}
\end{eqnarray}

\subsection{Gaussian Process Nonlinear Auto-Regressive with eXogenous inputs (NARX)}

Solar wind velocity $V$ and interplanetary magnetic field $B_z$ recorded at the ACE satellite are used as the exogenous inputs of the model, along with the time history of $D_{st}$.

  \begin{eqnarray}
    & D_{st}(t) & = f(\mathbf{x}_t) + \epsilon \\
        & \mathbf{x}_t & = (D_{st}(t-1), D_{st}(t-2), \cdots , D_{st}(t-p), \\
        & \ \ \ \ \ & V(t-1), V(t-2), \cdots, V(t-p),\\
        & \ \ \ \ \ & B_{z}(t-1), B_{z}(t-2), \cdots, B_{z}(t-p))\\
    & f(x_t) & \sim \mathcal{GP}(m(\mathbf{x}_t), K(\mathbf{x}_t,\mathbf{x}_s)) \label{eq:DstGP1}
\end{eqnarray}

\section{Experiments}

   \begin{table}
      \caption[]{Opacity sources.}
         \label{KapSou}
     $$ 
         \begin{array}{p{0.5\linewidth}l}
            \hline
            \noalign{\smallskip}
            Source      &  T / {[\mathrm{K}]} \\
            \noalign{\smallskip}
            \hline
            \noalign{\smallskip}
            \citet{Ji2012} & \leq 1700^{\mathrm{a}}     \\
            \citet{kocijan2015modelling}                     & 1700 \leq T \leq 5000 \\
            \citet{Berlinet2004}                  & 5000 \leq             \\
            \noalign{\smallskip}
            \hline
         \end{array}
     $$ 
\begin{list}{}{}
\item[$^{\mathrm{a}}$] This is footnote a
\end{list}
\end{table}

\section{Results}
%%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages

\begin{figure}
   \centering
   \subfigure{\includegraphics[width=\columnwidth]{Compare_RMSE.png}}
   \vspace{12pt}
   \subfigure{\includegraphics[width=\columnwidth]{Compare_CC.png}}
   \caption{\small  Performance Comparison} 
   \label{fig:rmsecc}
   \end{figure}

\begin{figure}
   \centering
   \subfigure{\includegraphics[width=\columnwidth]{Compare_deltaDst.png}}
   \vspace{12pt}
   \subfigure{\includegraphics[width=\columnwidth]{Compare_timingerr.png}}
   \caption{\small  Performance Comparison} 
   \label{fig:deltaDst}
   \end{figure}

\section{Conclusions}

   \begin{enumerate}
      \item Persistence model due to \emph{One Step Ahead} restriction 
      \item Superior performance of \emph{Gaussian Process} NAR and NARX models
      \item Multi step ahead prediction should be the preferred gold standard with respect to forecasting of geomagnetic indices.
   \end{enumerate}

\begin{acknowledgements}
      We acknowledge use of NASA/GSFC's Space Physics Data Facility's OMNIWeb (or CDAWeb or ftp) service, and OMNI data.
\end{acknowledgements}

%%    This version assumes use of bibtex with the swsc.bib file being present
%%    If your bib file has a different name you need to change the following line

\bibliography{swsc}

\end{linenumbers}

\end{document}

%%    If you wish to include your bibliography items in your tex file 
%%    using {thebibliography} as shown below you must out-comment the 
%%    three lines above (insert % at the start of each line)

\begin{thebibliography}{}

   \bibitem[Baker(1966)]{baker} Baker, N., Solar Evolution,
      in Stellar Evolution, ed.\ R. F. Stein and A. G. W. Cameron,
      Plenum, New York, 333, 1966.

   \bibitem[Balluch(1988)]{balluch} Balluch, M., A paper on something, 
      {\it Astron. Astrophys}, {\bf 200}, 58-63, 1988.

   \bibitem[Cox(1980)]{cox} Cox, J. P.,
      Theory of Stellar Pulsation,
      Princeton University Press, Princeton, 1980.

   \bibitem[Cox and Stewart(1969)]{cox69} Cox, A. N., and J. N. Stewart,
      Academia Nauk, {\it Scientific Information}, {\bf 15}, 1-11, 1969.

   \bibitem[Dumm and Kopf(2014)]{dummkopf} Dumm, Z., and Z. Kopf, A new self-inconsistent 
      core stability model, 
      {\it J. Space Weather Space Clim.}, {\bf 4}, A00, 2014, DOI: 10.1051/swsc/2014000,
      \url{http://dx.doi.org/10.1051/swsc/2014000}.

   \bibitem[Fool(2012)]{fool} Fool, X. Y., All earlier results are wrong, 
      {\it J. Space Weather Space Clim.}, {\bf 2}, A00, 2012, DOI: 10.1051/swsc/2012000,
      \url{http://dx.doi.org/10.1051/swsc/2012000}.

   \bibitem[Fou(2013)]{fou} Fou, Y. X., No earlier results were right, 
      {\it J. Space Weather Space Clim.}, {\bf 3}, A00, 2013, DOI: 10.1051/swsc/2013000,
      \url{http://dx.doi.org/10.1051/swsc/2013000}.

   \bibitem[Kr\"ugel et al.(1971)]{krugel71} Kr\"ugel, A. H., A. F. Davidsen, D. Tytler, and G. A. Kriss, 
      Summary of Project Achievements, Technical Report 08-11, 
      Fake Company, Nowherecity, Nowhereland, 1971.

   \bibitem[Mizuno(1980)]{mizuno} Mizuno, H., A further step backward in scientific understanding, 
      {\it Prog. Theor. Phys.}, {\bf 64}, 544-545, 1980.

   \bibitem[Tscharnuter(1987)]{tscharnuter} Tscharnuter, W. M., Who knows about astronomy?, 
      {\it Astron. Astrophys}, {\bf 188}, 55-57, 1987.

   \bibitem[Yorke(1980a)]{yorke80a} Yorke, H. W., A paper on nothing, 
      {\it Astron. Astrophys}, {\bf 86}, 286, 1980a.

   \bibitem[Yorke(1980b)]{yorke80b} Yorke, H. W., A paper on everything, 
      {\it Astron. Astrophys}, {\bf 86}, 291, 1980b.

\end{thebibliography}

\end{linenumbers}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%   Below are examples of how to use particular document layout features
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Examples for figures using graphicx
%%  A guide "Using Imported Graphics in LaTeX2e"  (Keith Reckdahl)
%%  is available on a lot of LaTeX public servers or ctan mirrors.
%%  The file is : epslatex.pdf 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%_____________________________________________________________
%                 A figure as large as the width of the column
%-------------------------------------------------------------
   \begin{figure}
   \centering
   \includegraphics[width=\textwidth]{empty.eps}
      \caption{Vibrational stability equation of state
               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
               $>0$ means vibrational stability.
              }
         \label{FigVibStab}
   \end{figure}
%
%_____________________________________________________________
%                                    One column rotated figure
%-------------------------------------------------------------
   \begin{figure}
   \centering
   \includegraphics[angle=-90,width=3cm]{empty.eps}
      \caption{Vibrational stability equation of state
               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
               $>0$ means vibrational stability.
              }
         \label{FigVibStab}
   \end{figure}
%
%_____________________________________________________________
%                        Figure with caption on the right side 
%-------------------------------------------------------------
   \begin{figure}
   \centering
   \includegraphics[width=3cm]{empty.eps}
      \caption{Vibrational stability equation of state
               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
               $>0$ means vibrational stability.
              }
         \label{FigVibStab}
   \end{figure}
%
%_____________________________________________________________
%
%_____________________________________________________________
%                                Figure with a new BoundingBox 
%-------------------------------------------------------------
   \begin{figure}
   \centering
   \includegraphics[bb=10 20 100 300,width=3cm,clip]{empty.eps}
      \caption{Vibrational stability equation of state
               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
               $>0$ means vibrational stability.
              }
         \label{FigVibStab}
   \end{figure}
%
%_____________________________________________________________
%
%_____________________________________________________________
%                                      The "resizebox" command 
%-------------------------------------------------------------
   \begin{figure}
   \resizebox{\textwidth}{!}
            {\includegraphics[bb=10 20 100 300,clip]{empty.eps}
      \caption{Vibrational stability equation of state
               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
               $>0$ means vibrational stability.
              }
         \label{FigVibStab}
   \end{figure}
%
%______________________________________________________________
%
%_____________________________________________________________
%                                             Simple A&A Table
%_____________________________________________________________
%
\begin{table}
\caption{Nonlinear Model Results}             % title of Table
\label{table:1}      % is used to refer this table in the text
\centering                          % used for centering table
\begin{tabular}{c c c c}        % centered columns (4 columns)
\hline\hline                 % inserts double horizontal lines
HJD & $E$ & Method\#2 & Method\#3 \\    % table heading 
\hline                        % inserts single horizontal line
   1 & 50 & $-837$ & 970 \\      % inserting body of the table
   2 & 47 & 877    & 230 \\
   3 & 31 & 25     & 415 \\
   4 & 35 & 144    & 2356 \\
   5 & 45 & 300    & 556 \\ 
\hline                                   %inserts single line
\end{tabular}
\end{table}
%
%_____________________________________________________________
%                                             Two column Table 
%_____________________________________________________________
%
\begin{table*}
\caption{Nonlinear Model Results}             
\label{table:1}      
\centering          
\begin{tabular}{c c c c l l l }     % 7 columns 
\hline\hline       
                      % To combine 4 columns into a single one 
HJD & $E$ & Method\#2 & \multicolumn{4}{c}{Method\#3}\\ 
\hline                    
   1 & 50 & $-837$ & 970 & 65 & 67 & 78\\  
   2 & 47 & 877    & 230 & 567& 55 & 78\\
   3 & 31 & 25     & 415 & 567& 55 & 78\\
   4 & 35 & 144    & 2356& 567& 55 & 78 \\
   5 & 45 & 300    & 556 & 567& 55 & 78\\
\hline                  
\end{tabular}
\end{table*}
%
%-------------------------------------------------------------
%                                          Table with notes 
%-------------------------------------------------------------
%
% A single note
\begin{table}[h]
\caption{\label{t7}Spectral types and photometry for stars in the
  region.}
\centering
\begin{tabular}{lccc}
\hline\hline
Star&Spectral type&RA(J2000)&Dec(J2000)\\
\hline
69           &B1\,V     &09 15 54.046 & $-$50 00 26.67\\
49           &B0.7\,V   &*09 15 54.570& $-$50 00 03.90\\
LS~1267~(86) &O8\,V     &09 15 52.787&11.07\\
24.6         &7.58      &1.37 &0.20\\
\hline
LS~1262      &B0\,V     &09 15 05.17&11.17\\
MO 2-119     &B0.5\,V   &09 15 33.7 &11.74\\
LS~1269      &O8.5\,V   &09 15 56.60&10.85\\
\hline
\end{tabular}
\tablefoot{ The top panel shows likely members of Pismis~11. The second
panel contains likely members of Alicante~5. The bottom panel
displays stars outside the clusters.}
\end{table}
%
% More notes
%
\begin{table}[h]
\caption{\label{t7}Spectral types and photometry for stars in the
  region.}
\centering
\begin{tabular}{lccc}
\hline\hline
Star&Spectral type&RA(J2000)&Dec(J2000)\\
\hline
69           &B1\,V     &09 15 54.046 & $-$50 00 26.67\\
49           &B0.7\,V   &*09 15 54.570& $-$50 00 03.90\\
LS~1267~(86) &O8\,V     &09 15 52.787&11.07\tablefootmark{a}\\
24.6         &7.58\tablefootmark{1}&1.37\tablefootmark{a}   &0.20\tablefootmark{a}\\
\hline
LS~1262      &B0\,V     &09 15 05.17&11.17\tablefootmark{b}\\
MO 2-119     &B0.5\,V   &09 15 33.7 &11.74\tablefootmark{c}\\
LS~1269      &O8.5\,V   &09 15 56.60&10.85\tablefootmark{d}\\
\hline
\end{tabular}
\tablefoot{ The top panel shows likely members of Pismis~11. The second
panel contains likely members of Alicante~5. The bottom panel
displays stars outside the clusters.\\
\tablefoottext{a}{Photometry for MF13, LS~1267 and HD~80077 from
Dupont et al.}
\tablefoottext{b}{Photometry for LS~1262, LS~1269 from
Durand et al.}
\tablefoottext{c}{Photometry for MO2-119 from
Mathieu et al.}
}
\end{table}
%
%-------------------------------------------------------------
%                                       Table with references 
%-------------------------------------------------------------
%
\begin{table*}[h]
 \caption[]{\label{nearbylistaa2}List of nearby SNe used in this work.}
\begin{tabular}{lccc}
 \hline \hline
  SN name &
  Epoch &
 Bands &
  References \\
 &
  (with respect to $B$ maximum) &
 &
 \\ \hline
1981B   & 0 & {\it UBV} & 1\\
1986G   &  $-$3, $-$1, 0, 1, 2 & {\it BV}  & 2\\
1989B   & $-$5, $-$1, 0, 3, 5 & {\it UBVRI}  & 3, 4\\
1990N   & 2, 7 & {\it UBVRI}  & 5\\
1991M   & 3 & {\it VRI}  & 6\\
\hline
\noalign{\smallskip}
\multicolumn{4}{c}{ SNe 91bg-like} \\
\noalign{\smallskip}
\hline
1991bg   & 1, 2 & {\it BVRI}  & 7\\
1999by   & $-$5, $-$4, $-$3, 3, 4, 5 & {\it UBVRI}  & 8\\
\hline
\noalign{\smallskip}
\multicolumn{4}{c}{ SNe 91T-like} \\
\noalign{\smallskip}
\hline
1991T   & $-$3, 0 & {\it UBVRI}  &  9, 10\\
2000cx  & $-$3, $-$2, 0, 1, 5 & {\it UBVRI}  & 11\\ %
\hline
\end{tabular}
\tablebib{(1)~\citet{branch83};
(2) \citet{phillips87}; (3) \citet{barbon90}; (4) \citet{wells94};
(5) \citet{mazzali93}; (6) \citet{gomez98}; (7) \citet{kirshner93};
(8) \citet{patat96}; (9) \citet{salvo01}; (10) \citet{branch03};
(11) \citet{jha99}.
}
\end{table}
%_____________________________________________________________
%                                 A rotated Table in landscape  
%  In the preamble, use:   \usepackage{lscape}
%-------------------------------------------------------------
\begin{landscape}
\begin{table*}
\caption{Summary for ISOCAM sources with mid-IR excess 
(YSO candidates).}\label{YSOtable}
\centering
\begin{tabular}{crrlcl} 
\hline\hline             
ISO-L1551 & $F_{6.7}$~[mJy] & $\alpha_{6.7-14.3}$ 
& YSO type$^{d}$ & Status & Comments\\
\hline
  \multicolumn{6}{c}{\it New YSO candidates}\\ % To combine 6 columns into a single one
\hline
  1 & 1.56 $\pm$ 0.47 & --    & Class II$^{c}$ & New & Mid\\
  2 & 0.79:           & 0.97: & Class II ?     & New & \\
  3 & 4.95 $\pm$ 0.68 & 3.18  & Class II / III & New & \\
  5 & 1.44 $\pm$ 0.33 & 1.88  & Class II       & New & \\
\hline
  \multicolumn{6}{c}{\it Previously known YSOs} \\
\hline
  61 & 0.89 $\pm$ 0.58 & 1.77 & Class I & \object{HH 30} & Circumstellar disk\\
  96 & 38.34 $\pm$ 0.71 & 37.5& Class II& MHO 5          & Spectral type\\
\hline
\end{tabular}
\end{table*}
\end{landscape}
%
%_____________________________________________________________
%                              Table longer than a single page  
%  In the preamble, use:              \usepackage{longtable}
%-------------------------------------------------------------
%          All long tables have to be placed at the end, after 
%                                        \end{thebibliography}
%
% In the text, at the place where the large table should appear
% add the command:
\addtocounter{table}{1}
% Tables counters will be well numbered.
%
\end{thebibliography}
% If table 2
\longtab{2}{
\begin{longtable}{lllrrr}
\caption{\label{kstars} Sample stars with absolute magnitude}\\
\hline\hline
Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
\hline
\endfirsthead
\caption{continued.}\\
\hline\hline
Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
\hline
\endhead
\hline
\endfoot
%%
Gl 33    & 6.37 & K2 V & 7.46 & S & 0.043170\\
Gl 66AB  & 6.26 & K2 V & 8.15 & S & 0.260478\\
Gl 68    & 5.87 & K1 V & 7.47 & P & 0.026610\\
         &      &      &      & H & 0.008686\\
Gl 86 
\footnote{Source not included in the HRI catalog. See Sect.~5.4.2 for details.}
         & 5.92 & K0 V & 10.91& S & 0.058230\\
\end{longtable}
}% End \longtab
%
%_____________________________________________________________
%                              Table longer than a single page
%                                             and in landscape 
%  In the preamble, use:       \usepackage{longtable,lscape}
%-------------------------------------------------------------
%          All long tables have to be placed at the end, after
%                                        \end{thebibliography}
%
% In the text, at the place where the large table should appear
% add the command:                        
\addtocounter{table}{1} 
% Tables counters will be well numbered.
%
\end{thebibliography}
% If table 2
\longtabL{2}{
\begin{landscape}
\begin{longtable}{lllrrr}
\caption{\label{kstars} Sample stars with absolute magnitude}\\
\hline\hline
Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
\hline
\endfirsthead
\caption{continued.}\\
\hline\hline
Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
\hline
\endhead
\hline
\endfoot
%%
Gl 33    & 6.37 & K2 V & 7.46 & S & 0.043170\\
Gl 66AB  & 6.26 & K2 V & 8.15 & S & 0.260478\\
Gl 68    & 5.87 & K1 V & 7.47 & P & 0.026610\\
         &      &      &      & H & 0.008686\\
Gl 86
\footnote{Source not included in the HRI catalog. See Sect.~5.4.2 for details.}
         & 5.92 & K0 V & 10.91& S & 0.058230\\
\end{longtable}
\end{landscape}
}% End \longtabL
%
% Online Material
%_____________________________________________________________
%        Online appendices have to be placed at the end, after
%                                        \end{thebibliography}
%-------------------------------------------------------------
\end{thebibliography}

\Online

\begin{appendix} %First online appendix
\section{Background galaxy number counts and shear noise-levels}
Because the optical images used in this analysis...

\begin{figure*}
\centering
\includegraphics[width=16.4cm,clip]{1787f24.ps}
\caption{Plotted above...}
\label{appfig}
\end{figure*}

Because the optical images...
\end{appendix}

\begin{appendix} %Second online appendix
These studies, however, have faced...
\end{appendix}

\end{document}
%
%_____________________________________________________________
%        Some tables or figures are in the printed version and
%                      some are only in the electronic version
%-------------------------------------------------------------
%
% Leave all the tables or figures in the text, at their right place 
% and use the commands \onlfig{}{} and \onltab{}{}. These elements
% will be automatically placed at the end, in the section
% Online material.

\documentclass{swsc}
...
\begin{document}
text of the paper...
\begin{figure*}%f1
\includegraphics[width=10.9cm]{1787f01.eps}
\caption{Shown in greyscale is a...}
\label{cl12301}}
\end{figure*}
...
from the intrinsic ellipticity distribution.
% Figure 2 available electronically only
\onlfig{2}{
\begin{figure*}%f2
\includegraphics[width=11.6cm]{1787f02.eps}
\caption {Shown in greyscale...}
\label{cl1018}
\end{figure*}
}

% Figure 3 available electronically only
\onlfig{3}{
\begin{figure*}%f3
\includegraphics[width=11.2cm]{1787f03.eps}
\caption{Shown in panels...}
\label{cl1059}
\end{figure*}
}

\begin{figure*}%f4
\includegraphics[width=10.9cm]{1787f04.eps}
\caption{Shown in greyscale is...}
\label{cl1232}}
\end{figure*}

\begin{table}%t1
\caption{Complexes characterisation.}\label{starbursts}
\centering
\begin{tabular}{lccc}
\hline \hline
Complex & $F_{60}$ & 8.6 &  No. of  \\
...
\hline
\end{tabular}
\end{table}
The second method produces...

% Figure 5 available electronically only
\onlfig{5}{
\begin{figure*}%f5
\includegraphics[width=11.2cm]{1787f05.eps}
\caption{Shown in panels...}
\label{cl1238}}
\end{figure*}
}

As can be seen, in general the deeper...
% Table 2 available electronically only
\onltab{2}{
\begin{table*}%t2
\caption{List of the LMC stellar complexes...}\label{Properties}
\centering
\begin{tabular}{lccccccccc}
\hline  \hline
Stellar & RA & Dec & ...
...
\hline
\end{tabular}
\end{table*}
}

% Table 3 available electronically only
\onltab{3}{
\begin{table*}%t3
\caption{List of the derived...}\label{IrasFluxes}
\centering
\begin{tabular}{lcccccccccc}
\hline \hline
Stellar & $f12$ & $L12$ &...
...
\hline
\end{tabular}
\end{table*}
}
%
%-------------------------------------------------------------
%     For the online material, table longer than a single page
%                 In the preamble, use: \usepackage{longtable}
%       or for landscape option: \usepackage{longtable,lscape}
%-------------------------------------------------------------
\documentclass{swsc}
\usepackage[varg]{txfonts}
\usepackage{graphicx}
\usepackage{longtable}

\begin{document}
text of the paper
% Table will be print automatically at the end, in the section Online material.
\onllongtab{3}{
\begin{longtable}{lrcrrrrrrrrl}
\caption{Line data and abundances ...}\\
\hline
\hline
Def & mol & Ion & $\lambda$ & $\chi$ & $\log gf$ & N & e &  rad & $\delta$ & $\delta$ 
red & References \\
\hline
\endfirsthead
\caption{Continued.} \\
\hline
Def & mol & Ion & $\lambda$ & $\chi$ & $\log gf$ & B & C &  rad & $\delta$ & $\delta$ 
red & References \\
\hline
\endhead
\hline
\endfoot
\hline
\endlastfoot
A & CH & 1 &3638 & 0.002 & $-$2.551 &  &  &  & $-$150 & 150 &  Jorgensen et al. (1996) \\                    
\end{longtable}
}% End onllongtab

% Or for landscape, large table:

\onllongtabL{3}{
\begin{landscape}
\begin{longtable}{lrcrrrrrrrrl}
...
\end{longtable}
\end{landscape}
}% End onllongtabL

### 2016-04-06 swsc.tex
Subsections added in methodology

@@
 
+\subsection{Formulation}
+
 \begin{eqnarray}
@@
       \left(\mathbf{y} \ \ \mathbf{f_*} \right)^\intercal & \sim & 
-\mathcal{N}\left(\mathbf{0}, \left[ \begin{array}{ccc} \mathbf{K}(\mathbf{X}, \mathbf{X}) + \sigma^{2} \mathbf{I} & \mathbf{K}(\mathbf{X}, \mathbf{X_*}) \\ \mathbf{K}(\mathbf{X_*}, \mathbf{X}) & \mathbf{K}(\mathbf{X_*}, \mathbf{X_*}) \end{array} \right ] \right) \\
+\mathcal{N}\left(\mathbf{0}, \left[ \begin{array}{cc} \mathbf{K}(\mathbf{X}, \mathbf{X}) + \sigma^{2} \mathbf{I} & \mathbf{K}(\mathbf{X}, \mathbf{X_*}) \\ \mathbf{K}(\mathbf{X_*}, \mathbf{X}) & \mathbf{K}(\mathbf{X_*}, \mathbf{X_*}) \end{array} \right ] \right)
+\end{eqnarray}
+
+\subsection{Inference and Predictions}
+
+\begin{eqnarray}
       \mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*} & \sim & \mathcal{N}(\mathbf{\bar{f}_*}, \Sigma_*)  \label{eq:posterior} \\
@@
 It is important to note that every kernel function generates a \emph{Reproducing Kernel Hilbert Space} (RKHS) of functions, and the nature of the kernel dictates what restrictions exist on the functions belonging to its RKHS.
+The Gaussian Process model essentially uses basis functions from this RKHS to represent the unknown function.
 The interested reader is suggested to refer to \cite{Berlinet2004} for a thorough treatment on RKHS and the rich theory behind them.
@@
 
-The RBF kernel is the most commonly used kernel in machine learning, in its core it
+The RBF kernel is the most commonly used kernel in machine learning, in its core it encodes the idea of \emph{smoothness} because functions drawn from its Hilbert space are infinitely differentiable.
+The RBF kernel is also an example of a \emph{stationary} kernel i.e.
+its value depends only on $||\mathbf{x} - \mathbf{y}||$.
+The quantity $\sigma$ is known as the characteristic \emph{length scale} of this kernel, it is a free parameter (also known as hyper-parameter) which must be determined by the modeller, refer to section \ref{sec:hyp} on how suitable values of kernel hyper-parameters.
 
@@
 
-\subsection{Optimizing Hyper-parameters}
+\subsection{Optimizing Hyper-parameters} \label{sec:hyp}
 

### 2016-04-07 swsc.tex
Added polynomial kernel paragraph

@@
     \begin{enumerate*}
-      \item Gaussian Process Nonlinear Auto regressive GP-NAR 
+      \item Gaussian Process Auto regressive GP-AR 
       \item Gaussian Process Auto regressive with eXogenous inputs GP-NARX. 
@@
 \subsubsection{Polynomial}
+
+The polynomial kernel is also a common kernel function employed in research, it has two hyper-parameters the degree $d$ and an intercept $b$.
+The polynomial kernel formulates the function approximation in terms of polynomial expansions of the input dimensions $\mathbf{x}$, more specifically it considers all monomial combination of the input dimensions up to degree $d$.
+When $d = 1$ it corresponds to a Bayesian linear regression model.
 

### 2016-04-07 swsc.tex
Added global opt section

@@
       \item Gaussian Process Auto regressive GP-AR 
-      \item Gaussian Process Auto regressive with eXogenous inputs GP-NARX. 
+      \item Gaussian Process Auto regressive with eXogenous inputs GP-ARX. 
     \end{enumerate*}
@@
 
-The covariance structure of a \emph{stochastic process} is represented by a symmetric positive definite function as given in \ref{eq:kern}, by \emph{Kolmogorov's extension theorem} (see \citet{tao2011introduction}), the function $K: \mathbb{R}^d \times \mathbb{R}^d \rightarrow \mathbb{R}$ uniquely determines the process.
+The covariance structure of a \emph{stochastic process} is represented by a symmetric positive definite function as given in \ref{eq:kern}.
+By \emph{Kolmogorov's extension theorem} (\citet{tao2011introduction}), the function $K: \mathbb{R}^d \times \mathbb{R}^d \rightarrow \mathbb{R}$ uniquely determines the process.
 Therefore this covariance function (also called kernel function in machine learning) helps us to specify the class of functions represented by the process.
@@
 
+In this study, we construct Gaussian Process models with polynomial kernels for \emph{one step ahead} prediction of the $D_{st}$ index.
+
 \subsection{Optimizing Hyper-parameters} \label{sec:hyp}
+
+With kernel based prediction models one always encounters the crucial question of what values to assign to the kernel hyper-parameters.
+This is known as \emph{model tuning} in kernel research parlance.
+There are many methods to perform \emph{model tuning}, one generally formulates an objective function and aims to optimize it in terms of the hyper-parameters.
+Common examples of objective functions include data likelihood, cross validation performance scores and validation set performance scores among others.
+Since the objective function for the model tuning problem is non-convex and non-smooth in general, gradient free approaches are used to find approximate values of local optima.
+
+The simplest way to perform model tuning is called \emph{grid search} which entails evaluating the above objective function on a fixed grid which is constructed by making a list of values for each hyper-parameter and taking the combinations of all such values.
+There are more sophisticated methods for \emph{model tuning} such as \emph{coupled simulated annealing} (\citet{Xavier-De-Souza2010}) and \emph{simplex search} (\citet{Nelder1965}) but they are computationally more expensive.
+In this study we chose \emph{grid search} as the model tuning routine.
 
@@
 
-\subsection{Gaussian Process Nonlinear Auto-Regressive (NAR)}
+\subsection{Gaussian Process Nonlinear Auto-Regressive (GP-AR)}
 
@@
 
-\subsection{Gaussian Process Nonlinear Auto-Regressive with eXogenous inputs (NARX)}
+\subsection{Gaussian Process Nonlinear Auto-Regressive with eXogenous inputs (GP-ARX)}
 

### 2016-04-19 swsc.tex
Changes 19 April

@@
 \usepackage[inline]{enumitem}
+\usepackage{verbatim}
 %%    This version assumes using bibtex with the swsc bibliography style file
@@
 
+The Sun is the principal driver of Space Weather, the steady solar wind and intermittent bursts of hot plasma from its corona (Coronal Mass Ejections \emph{CME}) are the causes of several observed phenomena in the Earths Magnetosphere such as but not limited to \emph{Aurora Borealis}, \emph{Aurora Australis}, \emph{Van Allen Currents}, geomagnetic storms etc.
+
+Although the qualitative bifurcation of these phenomena is well understood and each has been studied in depth, the magnetosphere's dynamics and its associated solar wind driver form a complex nonlinear coupled dynamical system.
+It is therefore not only instructive but also greatly simplifying to use representative indices to quantify the state of geomagnetic activity.
+
+Geomagnetic indices come in various forms, they may take continuous or discreet values.
+Their values are often calculated by averaging or combining a number of readings taken by instruments around the Earth.
+Each geomagnetic index is a proxy for a particular kind of phenomena.
+Some popular indices are.
+
+\begin{enumerate}
+    \item $K_p$: The Kp-index is a discreet valued global geomagnetic storm index and is based on 3 hour measurements of the K-indices, for a given value, for each of the past days.
+The K-index itself is a three hour long quasi-logarithmic local index of the geomagnetic activity, relative to a calm day curve for the given location.
+
+    \item AE: The Auroral Electrojet Index, AE, is designed to provide a global, quantitative measure of auroral zone magnetic activity produced by enhanced Ionospheric currents flowing below and within the auroral oval.
+
+    \item $D_{st}$: A continuous hourly index which measures the weakening of the Earths magnetic field due to ring currents. 
+\end{enumerate}
+
+We focus on prediction of the hourly $D_{st}$ index which is an important problem because $D_{st}$ is a straight forward indicator of the occurrence of geomagnetic storms.
+More specifically we focus on the \emph{one step ahead} prediction of $D_{st}$ because it is the first step towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions and because it has been the subject of several modeling efforts in the past.
+
 We use the hourly resolution Omni data set (see \cite{OmniPaper}) as extracted from NASA/GSFC's OMNI data set through OMNIWeb.
@@
 
-We restrict ourselves to the supervised learning problem in statistics, which is inductive.
-In other words, we aim to model functions of single or multiple variables (i.e.
-fields $f(\mathbf{x}), \ x \in \mathbb{R}^d $), using a set of labeled data points ${(\mathbf{x}_i, y_i), \forall i \in 1 \cdots N}$.
-These problems have been studied extensively in statistics over the past century.
+\emph{Gaussian Processes} first appeared in machine learning research as the limiting case of Bayesian inference performed on neural networks with infinitely many neurons in the hidden layers \citet{Neal:1996:BLN:525544}.
+It was noted by the author that "there may be simpler ways to do inference in this case".
+Although their inception in the machine learning community is recent, they were studied in statistics/mathematics community as special cases of \emph{Stochastic Processes}.
+The reader is requested to refer to \cite{Rasmussen:2005:GPM:1162254} for an in depth treatment of Gaussian Processes in Machine Learning.
 
+\begin{comment}
 \emph{Stochastic Processes} extend the idea of probability distributions to the space of functions of a single or multiple variables.
@@
 
-\emph{Gaussian Processes} first appeared in machine learning research as the limiting case of Bayesian inference performed on neural networks with infinitely many neurons in the hidden layers \citet{Neal:1996:BLN:525544}.
-It was noted by the author that "there may be simpler ways to do inference in this case".
-Although their inception in the machine learning community is recent, they were studied in statistics/mathematics community as special cases of \emph{Stochastic Processes}.
-The reader is requested to refer to \cite{Rasmussen:2005:GPM:1162254} for an in depth treatment of Gaussian Processes in Machine Learning.
+\end{comment}
 
 \subsection{Formulation}
+
+We restrict ourselves to the supervised learning problem in statistics, which is inductive.
+In other words, we aim to model functions of single or multiple variables (i.e.
+fields $f(\mathbf{x}), \ x \in \mathbb{R}^d $), using a set of labeled data points ${(\mathbf{x}_i, y_i), \forall i \in 1 \cdots N}$.
+These problems have been studied extensively in statistics over the past century.
 

### 2016-04-21 swsc.tex
Updates: April 21, 2016

@@
 \emph{Gaussian Processes} first appeared in machine learning research as the limiting case of Bayesian inference performed on neural networks with infinitely many neurons in the hidden layers \citet{Neal:1996:BLN:525544}.
-It was noted by the author that "there may be simpler ways to do inference in this case".
-Although their inception in the machine learning community is recent, they were studied in statistics/mathematics community as special cases of \emph{Stochastic Processes}.
+Although their inception in the machine learning community is recent, their origins can be traced back to the geo-statistics research community where they are known as \emph{Kriging} methods (\citet{krige1951statistical}).
+In pure mathematics area \emph{Gaussian Processes} have been studied extensively and their existence was first proven by Kolmogorov's extension theorem (\citet{tao2011introduction}).
 The reader is requested to refer to \cite{Rasmussen:2005:GPM:1162254} for an in depth treatment of Gaussian Processes in Machine Learning.
 
-\begin{comment}
-\emph{Stochastic Processes} extend the idea of probability distributions to the space of functions of a single or multiple variables.
-A \emph{Gaussian Process} is a stochastic process whose finite dimensional distributions are multivariate Gaussian \ref{eq:sto}.
-\vspace{-\baselineskip}
+A \emph{Gaussian Process} is a stochastic process $f(\mathbf{x}), \ \mathbf{x} \in \mathbb{R}^d$ whose finite dimensional distributions over a set of points $\mathbf{x}_1, \cdots, \mathbf{x}_n$ are multivariate Gaussian.
+Formally this is expressed as.
+
 \begin{eqnarray}
- \left(f(\mathbf{x}_1), f(\mathbf{x}_2) \cdots f(\mathbf{x}_N) \right)^T & \sim & \mathcal{N}\left(\mathbf{\mu}, \mathbf{\Lambda} \right) \label{eq:sto} \\
- \mathbf{\Lambda}_{ij} & = & K(\mathbf{x}_i, \mathbf{x}_j) \label{eq:kern}\\
- f & \sim & \mathcal{GP}(m(\mathbf{x}), K(\mathbf{x},\mathbf{x}'))
+ \mathbf{f} & = & \left( \begin{array}{c} f(\mathbf{x}_1) \\ f(\mathbf{x}_2) \\ \vdots \\ f(\mathbf{x}_n) \end{array} \right) \\
+ \vspace{2\baselineskip}
+ \mathbf{f} | \mathbf{x}_1, \cdots, \mathbf{x}_n & \sim & \mathcal{N}\left( \mathbf{\mu}, \mathbf{\Lambda} \right)  \\
+ \vspace{2\baselineskip}
+ p( \mathbf{f} \ | \ \mathbf{x}_1, \cdots, \mathbf{x}_n) & = & \frac{1}{(2\pi)^{n/2} det(\mathbf{\Lambda})^{1/2}} exp \left(-\frac{1}{2} (\mathbf{f} - \mathbf{\mu})^T \mathbf{\Lambda}^{-1} (\mathbf{f} - \mathbf{\mu}) \right) \label{eq:sto}
 \end{eqnarray}
 
-The covariance structure of a \emph{stochastic process} is represented by a symmetric positive definite function as given in \ref{eq:kern}.
-By \emph{Kolmogorov's extension theorem} (\citet{tao2011introduction}), the function $K: \mathbb{R}^d \times \mathbb{R}^d \rightarrow \mathbb{R}$ uniquely determines the process.
-Therefore this covariance function (also called kernel function in machine learning) helps us to specify the class of functions represented by the process.
+From equation \ref{eq:sto}, one can observe that in order to uniquely define the distribution of the process, it is required to specify the mean vector $\mathbf{\mu}$ and the covariance matrix $\mathbf{\Lambda}$.
+For this probability density to be valid, there is a further requirements imposed on $\mathbf{\Lambda}$:
 
-\end{comment}
+\begin{enumerate}
+      \item Symmetry: $\mathbf{\Lambda}_{ij} = \mathbf{\Lambda}_{ij} \ \forall i,j \in {1, \cdots, n} $ 
+      \item Positive Semi-definiteness: $\mathbf{y}^T \mathbf{\Lambda} \mathbf{y} \geq 0 \ \forall \mathbf{y} \in \mathbb{R}^n$  
+\end{enumerate}
 
-\subsection{Formulation}
-
-We restrict ourselves to the supervised learning problem in statistics, which is inductive.
-In other words, we aim to model functions of single or multiple variables (i.e.
-fields $f(\mathbf{x}), \ x \in \mathbb{R}^d $), using a set of labeled data points ${(\mathbf{x}_i, y_i), \forall i \in 1 \cdots N}$.
-These problems have been studied extensively in statistics over the past century.
+Inspecting the individual elements of $\mathbf{\mu}$ and $\mathbf{\Lambda}$, we realise that they take the following form.
 
 \begin{eqnarray}
-      \mathbf{x} & \in & \mathbb{R}^d \\
-      \mathbf{X} & = & (\mathbf{x}_1, \mathbf{x}_2, \cdots, \mathbf{x}_N) \\
-      \mathbf{y} & = & (y_1, y_2, \cdots, y_N) \\
+      \mu_i & = & \mathbb{E}[f(\mathbf{x}_i)] := m(\mathbf{x}_i) \\
+      \Lambda_{ij} & = & \mathbb{E}[(f(\mathbf{x}_i) - \mu_i)(f(\mathbf{x}_j) - \mu_j)] := K(\mathbf{x}_i, \mathbf{x}_j)
+\end{eqnarray}
+
+Here $\mathbb{E}$ denotes the expectation (average).
+The elements of $\mathbf{\mu}$ and $\mathbf{\Lambda}$ are expressed as functions $m(\mathbf{x}_i)$ and $K(\mathbf{x}_i, \mathbf{x}_j)$ of the inputs $\mathbf{x}_i,\ \mathbf{x}_j$.
+Specifying the functions $m(\mathbf{x})$ and $K(\mathbf{x}, \mathbf{x}')$ completely specifies each element of $\mathbf{\mu}$ and $\mathbf{\Lambda}$ and subsequently the finite dimensional distribution of $\mathbf{f} | \mathbf{x}_1, \cdots, \mathbf{x}_n $.
+In machine learning literature \emph{Gaussian Processes} are represented using the following notation.
+
+\begin{equation}
+    f(\mathbf{x}) \sim \mathcal{GP}(m(\mathbf{x}), K(\mathbf{x}, \mathbf{x}'))
+\end{equation}
+
+In most practical applications of \emph{Gaussian Processes} the mean function is often defined as $m(\mathbf{x}) = 0$, which is not unreasonable if the data is standardized to have zero mean.
+The central design issue now becomes choice of the function $K(\mathbf{x}, \mathbf{x}')$.
+The same constraints that apply to $\mathbf{\Lambda}$ also apply to the function $K(.,.)$.
+In machine learning, these symmetric positive definite functions of two variables are known as \emph{kernels}.
+Kernel based methods are applied extensively in data analysis i.e.
+regression, clustering, classification, density estimation (see \citet{Scholkopf:2001:LKS:559923}, \citet{hofmann2008}).
+
+\subsection{Gaussian Process Regression: Formulation}
+
+\begin{eqnarray}
+  \mathbf{X} & = & \left( \begin{array}{c} \mathbf{x}^{T}_1 \\ \mathbf{x}^{T}_2 \\ \vdots \\ \mathbf{x}^{T}_n \end{array} \right)_{n \times d} \label{eq:feat} \\
+  \vspace{2\baselineskip}
+  \mathbf{y} & = & \left( \begin{array}{c} y_1 \\ y_2 \\ \vdots \\ y_N \end{array} \right) _{n \times 1} \label{eq:labels}
+\end{eqnarray}
+
+We start with a set of labeled data points $ {(\mathbf{x}_i, y_i): \forall i \in 1 \cdots N, \ \mathbf{x}_i \in \mathbb{R}^d \ y_i \in \mathbb{R}} $ condensed into a data matrix $\mathbf{X}$ and response vector $\mathbf{y}$ in equations \ref{eq:feat} and \ref{eq:labels}.
+The values $y_i$ are assumed to be measurements of an unknown function $f(\mathbf{x}_i)$ corrupted by an independent and identically distributed zero mean Gaussian noise.
+The \emph{Gaussian Process} regression framework is the defined by the following components.
+
+\begin{eqnarray}
       y & = & f(\mathbf{x}) + \epsilon  \\
-      f & \sim & \mathcal{GP}(m(\mathbf{x}), K(\mathbf{x},\mathbf{x}'))  \\
-      \left(\mathbf{y} \ \ \mathbf{f_*} \right)^\intercal & \sim & 
-\mathcal{N}\left(\mathbf{0}, \left[ \begin{array}{cc} \mathbf{K}(\mathbf{X}, \mathbf{X}) + \sigma^{2} \mathbf{I} & \mathbf{K}(\mathbf{X}, \mathbf{X_*}) \\ \mathbf{K}(\mathbf{X_*}, \mathbf{X}) & \mathbf{K}(\mathbf{X_*}, \mathbf{X_*}) \end{array} \right ] \right)
+      \epsilon & \sim & \mathcal{N}(0, \sigma^2) \\
+      f(\mathbf{x}) & \sim & \mathcal{GP}(m(\mathbf{x}), K(\mathbf{x},\mathbf{x}')) 
 \end{eqnarray}
@@
 
+Our aim is to generate infer the function $f(\mathbf{x})$ from the noisy training data and generate predictions $f(\mathbf{x}^{*}_i)$ for a set of test points $ {\mathbf{x}^{*}_i : \forall i \in 1, \cdots, m} $.
+We start from equation \ref{eq:sto} and use it to construct the distribution of $f(\mathbf{x})$ over the training and test points, how ever since we have noisy measurements of $f$ over the training data, we add the noise variance $\sigma^2$ to the variance of $f$ as shown in \ref{eq:dist}.
+
+\begin{eqnarray}
+    \mathbf{X}_* & = & \left( \begin{array}{c} (\mathbf{x}^{*}_1)^T \\ (\mathbf{x}^{*}_2)^T \\ \vdots \\ (\mathbf{x}^{*}_m)^T \end{array} \right)_{m \times d} \\
+    \mathbf{f}_* & = & \left( \begin{array}{c} f(\mathbf{x^{*}_1}) \\ f(\mathbf{x^{*}_2}) \\ \vdots \\ f(\mathbf{x^{*}_m}) \end{array} \right)_{m \times 1} \\
+     \vspace{4\baselineskip}
+    \left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right) | \ \ \mathbf{X}, \mathbf{X}_* & \sim & 
+    \mathcal{N}\left(\mathbf{0}, \left[ \begin{array}{cc} \mathbf{K} + \sigma^{2} \mathbf{I} & \mathbf{K}_{*} \\ \mathbf{K}_{*}^T & \mathbf{K}_{**} \end{array} \right ] \right) \label{eq:dist}
+\end{eqnarray}
+
+In equation \ref{eq:dist} the block matrix components of the $n+m \times n+m$ covariance matrix have the following structure.
+
+\begin{enumerate}
+      \item $\mathbf{I}$: The $n \times n$ identity matrix.
+      \item $\mathbf{K} = [K(\mathbf{x}_i, \mathbf{x}_j)], \ i,j \in 1,\cdots,n$ : Kernel matrix constructed from all couples of the training data.
+      \item $\mathbf{K}_{*} = [K(\mathbf{x}_i, \mathbf{x}^{*}_j)], \ i \in 1,\cdots,n ; j \in 1,\cdots,m$ : Cross kernel matrix constructed from all couples between training and test data points.
+      \item $\mathbf{K}_{**} = [K(\mathbf{x}^{*}_i, \mathbf{x}^{*}_j)], \ i,j \in 1,\cdots,m$: Kernel matrix constructed from all couples of the test data.
+\end{enumerate}
+
+With the multivariate normal distribution defined in equation \ref{eq:dist}, probabilistic predictions $f_*$ can be generated by constructing the conditional distribution $\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}$, since the original distribution of $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right) | \ \ \mathbf{X}, \mathbf{X}_*$ is a multivariate Gaussian, conditioning on a subset of elements $\mathbf{y}$ yields another Gaussian distribution with mean and covariance given below \ref{eq:posterior}.
+
 \begin{eqnarray}
     \mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*} & \sim & \mathcal{N}(\mathbf{\bar{f}_*}, \Sigma_*)  \label{eq:posterior} \\
-      \mathbf{\bar{f}_*} & = & \mathbb{E}[\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}]  \\
-      & = & K(\mathbf{X_*},\mathbf{X})^\intercal [K(\mathbf{X},\mathbf{X}) + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y} \\
-      \Sigma_* & = & K(\mathbf{X_*}, \mathbf{X_*}) - K(\mathbf{X_*},\mathbf{X})^\intercal \left(K(\mathbf{X},\mathbf{X}) + \sigma^{2} \mathbf{I}\right)^{-1}K(\mathbf{X},\mathbf{X_*})
+    \mathbf{\bar{f}_*} & = & \mathbf{K}^T_{*} [\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y} \\
+    \Sigma_* & = & \mathbf{K}_{**} - \mathbf{K}^T_{*} \left(\mathbf{K} + \sigma^{2} \mathbf{I}\right)^{-1} \mathbf{K}_{*}
 \end{eqnarray}
@@
 For the success of a Gaussian Process model an appropriate choice of kernel function is paramount.
-It is important to note that every kernel function generates a \emph{Reproducing Kernel Hilbert Space} (RKHS) of functions, and the nature of the kernel dictates what restrictions exist on the functions belonging to its RKHS.
+Every kernel function generates a \emph{Reproducing Kernel Hilbert Space} (RKHS) of functions, and the nature of the kernel dictates what restrictions exist on the functions belonging to its RKHS.
 The Gaussian Process model essentially uses basis functions from this RKHS to represent the unknown function.
 The interested reader is suggested to refer to \cite{Berlinet2004} for a thorough treatment on RKHS and the rich theory behind them.
-
 Some common kernel functions used in machine learning are.
@@
 its value depends only on $||\mathbf{x} - \mathbf{y}||$.
-The quantity $\sigma$ is known as the characteristic \emph{length scale} of this kernel, it is a free parameter (also known as hyper-parameter) which must be determined by the modeller, refer to section \ref{sec:hyp} on how suitable values of kernel hyper-parameters.
+The quantity $l$ is known as the characteristic \emph{length scale} of this kernel, it is a free parameter (also known as hyper-parameter) which must be determined by the modeller, refer to section \ref{sec:hyp} on how suitable values of kernel hyper-parameters.
 
 \begin{equation}
-    K(\mathbf{x}, \mathbf{y}) = \frac{1}{2} exp(-||\mathbf{x} - \mathbf{y}||^2/\sigma^2)
+    K(\mathbf{x}, \mathbf{y}) = \frac{1}{2} exp(-||\mathbf{x} - \mathbf{y}||^2/l^2)
 \end{equation}
@@
 
-In this study, we construct Gaussian Process models with polynomial kernels for \emph{one step ahead} prediction of the $D_{st}$ index.
+In this study, we construct Gaussian Process regression models with polynomial kernels for \emph{one step ahead} prediction of the $D_{st}$ index.
 

### 2016-04-22 swsc.tex
Changes April 22, 2016

@@
 We use the hourly resolution Omni data set (see \cite{OmniPaper}) as extracted from NASA/GSFC's OMNI data set through OMNIWeb.
+The rest of the paper is organized as follows.
+Section \ref{sec:method} gives an introduction to the \emph{Gaussian Process} methodology for regression and the various issues surrounding their application.
+In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res} bench marked against the \emph{Persistence} $D_{st}$ model as well as other models outlined in \citet{Ji2012}.
 
-\section{Methodology: Gaussian Process}
+\section{Methodology: Gaussian Process} \label{sec:method}
 
@@
 For the success of a Gaussian Process model an appropriate choice of kernel function is paramount.
-Every kernel function generates a \emph{Reproducing Kernel Hilbert Space} (RKHS) of functions, and the nature of the kernel dictates what restrictions exist on the functions belonging to its RKHS.
-The Gaussian Process model essentially uses basis functions from this RKHS to represent the unknown function.
-The interested reader is suggested to refer to \cite{Berlinet2004} for a thorough treatment on RKHS and the rich theory behind them.
+The symmetry and positive semi-definiteness of \emph{Gaussian Process} kernels implies that they represent inner-products between some basis function representation of the data.
+This is analogous to a Fourier series decomposition in Physics.
+The interested reader is suggested to refer to \cite{Berlinet2004}, \citet{Scholkopf:2001:LKS:559923} and \citet{hofmann2008} for a thorough treatment on kernel functions and the rich theory behind them.
 Some common kernel functions used in machine learning are.
@@
 
-With kernel based prediction models one always encounters the crucial question of what values to assign to the kernel hyper-parameters.
-This is known as \emph{model tuning} in kernel research parlance.
-There are many methods to perform \emph{model tuning}, one generally formulates an objective function and aims to optimize it in terms of the hyper-parameters.
-Common examples of objective functions include data likelihood, cross validation performance scores and validation set performance scores among others.
-Since the objective function for the model tuning problem is non-convex and non-smooth in general, gradient free approaches are used to find approximate values of local optima.
+When fitting \emph{Gaussian Process} models to data, the question to be answered following choice of a kernel function is the assignment of appropriate values to the kernel hyper-parameters and noise level ($\sigma^2$).
+This is known as the problem of \emph{model selection}.
 
-The simplest way to perform model tuning is called \emph{grid search} which entails evaluating the above objective function on a fixed grid which is constructed by making a list of values for each hyper-parameter and taking the combinations of all such values.
+Common examples of criteria used for \emph{model selection} include data likelihood, cross validation performance scores and validation set performance scores among others.
+Since model selection criteria are non-convex and non-smooth in general, gradient free approaches are used to find approximate values of local optima.
+
+The simplest way to perform \emph{model selection} is called \emph{grid search} which entails evaluating the above criteria on a fixed grid which is constructed by making a list of values for each hyper-parameter and taking the combinations of all such values.
 There are more sophisticated methods for \emph{model tuning} such as \emph{coupled simulated annealing} (\citet{Xavier-De-Souza2010}) and \emph{simplex search} (\citet{Nelder1965}) but they are computationally more expensive.
-In this study we chose \emph{grid search} as the model tuning routine.
+In this study we employ \emph{grid search} to find the best performing set of hyper-parameter assignments inside a predefined grid of possible hyper-parameter values.
 
-\section{One Step Ahead Prediction}
+\section{One Step Ahead Prediction} \label{sec:osa}
 
-\subsection{Gaussian Process Nonlinear Auto-Regressive (GP-AR)}
+Below in equations \ref{eq:Dst}, \ref{eq:GPNoise}, \ref{eq:DstGP}, \ref{eq:GPPoly} we outline a \emph{Gaussian Process} formulation for \emph{OSA} prediction of $D_{st}$.
+A vector of features $x_{t-1}$ is used as input to an unknown function $f(x_{t-1})$ which drawn from a \emph{Gaussian Process} driven by a linear polynomial kernel.
 
-The input vector $x_t$ at each time step is the $p$ order time history of $D_{st}$.
+The features $x_{t-1}$ can be any collection of quantities in the hourly resolution Omni data set.
+Generally $x_{t-1}$ are time histories of $D_{st}$ or other important variables examples being plasma pressure $p(t)$, solar wind speed $V(t)$, z component of the interplanetary magnetic field $B_z(t)$.
+
 \begin{eqnarray}
-    D_{st}(t) & = & f(x_t) + \epsilon\\
-    x_t & = & \left(D_{st}(t-1), D_{st}(t-2), \cdots , D_{st}(t-p)\right) \\
-    f(x_t) & \sim & \mathcal{GP}(m(x_t), K(x_t,x_s)) \label{eq:DstGP}
+    D_{st}(t) & = & f(\mathbf{x}_{t-1}) + \epsilon \label{eq:Dst} \\
+    \epsilon & \sim & \mathcal{N}(0, \sigma^2) \label{eq:GPNoise} \\
+    f(x_t) & \sim & \mathcal{GP}(m(\mathbf{x}_t), K(\mathbf{x}_t, \mathbf{x}_s)) \label{eq:DstGP} \\
+    K(\mathbf{x}, \mathbf{y}) & = & \mathbf{x}^\intercal \mathbf{y} + b \label{eq:GPPoly}
 \end{eqnarray}
 
-\subsection{Gaussian Process Nonlinear Auto-Regressive with eXogenous inputs (GP-ARX)}
+In \ref{sec:gpar} and \ref{sec:gparx} below we consider two choices for the input features $\mathbf{x}_{t-1}$ leading to two variants of \emph{Gaussian Process} regression for $D_{st}$ time series prediction.
 
-Solar wind velocity $V$ and interplanetary magnetic field $B_z$ recorded at the ACE satellite are used as the exogenous inputs of the model, along with the time history of $D_{st}$.
+\subsection{Gaussian Process Auto-Regressive (GP-AR)} \label{sec:gpar}
 
-  \begin{eqnarray}
-    & D_{st}(t) & = f(\mathbf{x}_t) + \epsilon \\
-        & \mathbf{x}_t & = (D_{st}(t-1), D_{st}(t-2), \cdots , D_{st}(t-p), \\
-        & \ \ \ \ \ & V(t-1), V(t-2), \cdots, V(t-p),\\
-        & \ \ \ \ \ & B_{z}(t-1), B_{z}(t-2), \cdots, B_{z}(t-p))\\
-    & f(x_t) & \sim \mathcal{GP}(m(\mathbf{x}_t), K(\mathbf{x}_t,\mathbf{x}_s)) \label{eq:DstGP1}
-\end{eqnarray}
+The simplest auto-regressive models for \emph{OSA} prediction of $D_{st}$ are those that use only the history of $D_{st}$ to construct input features for model training.
+The input features $\mathbf{x}_{t-1}$ at each time step is the history of $D_{st}(t)$ until a time lag of $p$ steps.
 
-\section{Experiments}
+\begin{eqnarray*}
+    \mathbf{x}_t & = & \left(D_{st}(t), D_{st}(t-1), \cdots , D_{st}(t-p+1)\right)
+\end{eqnarray*}
 
+\subsection{Gaussian Process Auto-Regressive with eXogenous inputs (GP-ARX)} \label{sec:gparx}
+
+Auto-regressive models can be improved by including exogenous quantities that can help in improving predictions.
+It is well known that the solar wind and interplanetary magnetic field are important factors that influence the Earth's geomagnetic state.
+
+Hourly solar wind velocity $V(t)$ and interplanetary magnetic field $B_z(t)$ recorded at the ACE satellite are used as the exogenous inputs of the model, along with the time history of $D_{st}$.
+In this model we choose distinct time lags $p$ and $p_{ex}$ for auto-regressive and exogenous variables respectively.
+
+\begin{eqnarray*}
+        & \mathbf{x}_t & = (D_{st}(t), D_{st}(t-1), \cdots , D_{st}(t-p+1), \\
+        & \ \ \ \ \ & V(t), V(t-1), \cdots, V(t-p_{ex}+1),\\
+        & \ \ \ \ \ & B_{z}(t), B_{z}(t-1), \cdots, B_{z}(t-p_{ex}+1),\\
+        & \ \ \ \ \ & V B_{z}(t), V B_{z}(t-1), \cdots, V B_{z}(t-p_{ex}+1))
+\end{eqnarray*}
+
+\section{Model Training and Validation}
+
+\section{Experiments} \label{sec:exp}
+
+\begin{comment}
+%%
+%%
+%%
 \begin{table}
@@
 \end{table}
+%%
+%%
+%%
+\end{comment}
 
-\section{Results}
+\section{Results} \label{sec:res}
 %%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages
+
+\begin{figure}
+   \centering
+   \subfigure{\includegraphics[width=\columnwidth]{Compare_RelProb.png}}
+   \caption{\small  Comparison of empirical distribution of relative errors for various models on the list of 63 storms in \citet{Ji2012}.} 
+   \label{fig:relprob}
+   \end{figure}
 

### 2016-04-22 swsc.tex
Updates from ShareLaTeX

@@
 
+Before performance bench marks for \emph{OSA} $D_{st}$ prediction are run on the 63 geomagnetic storm events from 1998-2006 in \cite{Ji2012}, training and model selection of \emph{GP-AR} and \emph{GP-ARX} models on independent data sets must be performed.
+For this purpose we choose segments 00:00 1 January 2008 - 10:00 11 January 2008 and 00:00 15 November 2014 - 23:00 1 December 2014 for training and model selection respectively.
+
+The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$ are arrived by experimenting with values starting from one and choosing the value of $p$ yielding best performance on the 63 events in the bench mark.
+
 \section{Experiments} \label{sec:exp}
 
-\begin{comment}
-%%
-%%
-%%
+\subsection{Models Compared}
+
 \begin{table}
-      \caption[]{Opacity sources.}
-         \label{KapSou}
-     $$ 
-         \begin{array}{p{0.5\linewidth}l}
+      \caption[]{One Step Ahead $D_{st}$ prediction models}
+         \label{table:DstModels}
+
+         \begin{tabular}{lll}
             \hline
             \noalign{\smallskip}
-            Source      &  T / {[\mathrm{K}]} \\
+            Model     &  Reference & Description \\
             \noalign{\smallskip}
@@
             \noalign{\smallskip}
-            \citet{Ji2012} & \leq 1700^{\mathrm{a}}     \\
-            \citet{kocijan2015modelling}                     & 1700 \leq T \leq 5000 \\
-            \citet{Berlinet2004}                  & 5000 \leq             \\
+            TL & \citet{JGRA:JGRA16300} & Auto regressive model decomposable into additive terms.      \\
+            NM & \citet{balikhin:narmax} & Non linear auto regressive with exogenous inputs. \\
+            B & \citet{JGR:JGR10260} & Prediction of $D_{st}$ by solving ODE having injection and decay terms. \\
+            W & \citet{Wang:Dst} & Obtained by modification of injection term in \citet{JGR:JGR10260}. \\
+            FL & \citet{GRL:GRL11549} & Uses polarity of magnetic clouds to predict geomagnetic response.\\
+            OM & \citet{JGRA:JGRA14856} & Modification of injection term in \citet{JGR:JGR10260}.\\
             \noalign{\smallskip}
             \hline
-         \end{array}
-     $$ 
-\begin{list}{}{}
-\item[$^{\mathrm{a}}$] This is footnote a
-\end{list}
+         \end{tabular}
 \end{table}
-%%
-%%
-%%
-\end{comment}
 

### 2016-04-24 swsc.tex
Updates from ShareLaTeX

@@
 
-Before performance bench marks for \emph{OSA} $D_{st}$ prediction are run on the 63 geomagnetic storm events from 1998-2006 in \cite{Ji2012}, training and model selection of \emph{GP-AR} and \emph{GP-ARX} models on independent data sets must be performed.
+Before performance bench marks for \emph{OSA} $D_{st}$ prediction are run on the storm events in \cite{Ji2012}, training and model selection of \emph{GP-AR} and \emph{GP-ARX} models on independent data sets must be performed.
 For this purpose we choose segments 00:00 1 January 2008 - 10:00 11 January 2008 and 00:00 15 November 2014 - 23:00 1 December 2014 for training and model selection respectively.
@@
 
+\subsection{Storm Events}
+\citet{Ji2012}} compare a six OSA $D_{st}$ models in their paper by compiling a list of 63 geomagnetic storm events of varying strengths which occurred from 1998-2006.
+This serves as an important stepping stone for a systematic comparison of existing and new prediction techniques because performance metrics are averaged over a large number of storm events that occurred over a 8 year period.
+
 \subsection{Models Compared}
@@
 \begin{table}
-      \caption[]{One Step Ahead $D_{st}$ prediction models}
+      \caption[]{One Step Ahead $D_{st}$ prediction models compared in \citet{Ji2012}}
          \label{table:DstModels}

### 2016-04-24 swsc.tex
Updates from ShareLaTeX

@@
 There are more sophisticated methods for \emph{model tuning} such as \emph{coupled simulated annealing} (\citet{Xavier-De-Souza2010}) and \emph{simplex search} (\citet{Nelder1965}) but they are computationally more expensive.
+
 In this study we employ \emph{grid search} to find the best performing set of hyper-parameter assignments inside a predefined grid of possible hyper-parameter values.
@@
 
+Model selection of \emph{GP-AR} and \emph{GP-ARX} is performed by grid search methodology and the root mean square error (\emph{RMSE}) on the validation set is used to select the values of the model hyper-parameters.
 The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$ are arrived by experimenting with values starting from one and choosing the value of $p$ yielding best performance on the 63 events in the bench mark.
@@
 
-\subsection{Storm Events}
-\citet{Ji2012}} compare a six OSA $D_{st}$ models in their paper by compiling a list of 63 geomagnetic storm events of varying strengths which occurred from 1998-2006.
+\citet{Ji2012} compare a six OSA $D_{st}$ models in their paper by compiling a list of 63 geomagnetic storm events of varying strengths which occurred from 1998-2006.
 This serves as an important stepping stone for a systematic comparison of existing and new prediction techniques because performance metrics are averaged over a large number of storm events that occurred over a 8 year period.
+They compare the averaged performance of these models on four key performance metrics.
+
+\begin{enumerate}
+    \item $RMSE$: The root mean square error
+    \item $CC$: Correlation coefficient between the predicted and actual value of $D_{st}$
+    \item $\Delta Dst_{min}$: The error in prediction of the peak value of $D_{st}$ for a storm event.
+    \item $|\Delta t_{peak}|$: The error in predicting the timing of a storm peak, also called \emph{timing error}. 
+\end{enumerate}
+
+Apart from the metrics above, we also generate the empirical distribution of absolute relative errors $|\frac{D_{st} - \hat{D_{st}}}{D_{st}}|$ for the \emph{GP-AR}, \emph{GP-ARX}, \emph{NM} and \emph{TL} models.
+These empirical cumulative distributions enable us to upper bound the percentage error in $D_{st}$ prediction and attach probability estimates for said bounds.
 
 \subsection{Models Compared}
+
+Table \ref{table:DstModels} gives a brief enumeration of the models compared, we use the performance benchmark results published in \citet{Ji2012} and to them add results obtained from testing \emph{GP-AR}, \emph{GP-ARX} as well as the \emph{Persistence} model on the same list of storms.
+
+For the purpose of generating the cumulative probability plots of the model errors, we must obtain hourly predictions for all the storm events using the \emph{NARMAX} and \emph{TL} models.
+In the case of the \emph{NM} model, the formula outlined in \citet{balikhin:narmax} is used to generate hourly predictions.
+For the \emph{TL} model we use real time predictions listed on their website corresponding to all the storm events.
+The real time predictions are at a frequency of ten minutes, hourly predictions are generated by averaging the $D_{st}$ predictions for each hour.
 

### 2016-04-24 swsc.tex
Updates from ShareLaTeX

@@
     \begin{enumerate*}
-      \item Gaussian Process Auto regressive GP-AR 
-      \item Gaussian Process Auto regressive with eXogenous inputs GP-ARX. 
+      \item Gaussian Process Auto regressive (GP-AR) 
+      \item Gaussian Process Auto regressive with eXogenous inputs (GP-ARX). 
     \end{enumerate*}
@@
 
+\begin{figure}[ht]
+   \centering
+   \subfigure{\includegraphics[width=\columnwidth]{Compare_RMSE.png}}
+   \vspace{12pt}
+   \subfigure{\includegraphics[width=\columnwidth]{Compare_CC.png}}
+   \caption{\small  Performance Comparison; left: Averaged RMSE right: Averaged Cross correlation coefficient} 
+   \label{fig:rmsecc}
+   \end{figure}
+
 \section{Experiments} \label{sec:exp}
@@
 These empirical cumulative distributions enable us to upper bound the percentage error in $D_{st}$ prediction and attach probability estimates for said bounds.
+
+\begin{figure}
+   \centering
+   \subfigure{\includegraphics[width=\columnwidth]{Compare_deltaDst.png}}
+   \vspace{12pt}
+   \subfigure{\includegraphics[width=\columnwidth]{Compare_timingerr.png}}
+   \caption{\small  Performance Comparison; left: Averaged $\Delta Dst_{min}$ right: Averaged absolute timing error} 
+   \label{fig:deltaDst}
+   \end{figure}
 
@@
 
+With respect to the \emph{TL} model one important concern must be expressed.
+The training of the \emph{TL} model was done on Omni data from 1996-2002 which overlaps with a large number of the storm events, therefore the experimental procedure has a strong bias towards \emph{TL} model.
+
 \section{Results} \label{sec:res}
 %%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages
+
+Figures \ref{fig:rmsecc} and \ref{fig:deltaDst} compare the performance of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model with the existing results of \citet{Ji2012}.
+Two central themes come out of these charts.
 
@@
 
-\begin{figure}
-   \centering
-   \subfigure{\includegraphics[width=\columnwidth]{Compare_RMSE.png}}
-   \vspace{12pt}
-   \subfigure{\includegraphics[width=\columnwidth]{Compare_CC.png}}
-   \caption{\small  Performance Comparison} 
-   \label{fig:rmsecc}
-   \end{figure}
-
-\begin{figure}
-   \centering
-   \subfigure{\includegraphics[width=\columnwidth]{Compare_deltaDst.png}}
-   \vspace{12pt}
-   \subfigure{\includegraphics[width=\columnwidth]{Compare_timingerr.png}}
-   \caption{\small  Performance Comparison} 
-   \label{fig:deltaDst}
-   \end{figure}
+\subsection{Persistence Behavior}
+
+It is clear that the \emph{persistence} behavior in the $D_{st}$ values is very strong i.e.
+the trivial predictive model $\hat{D_{st}}(t) = D_{st}(t-1)$ gives excellent performance.
+In fact it can be seen in these figures that the \emph{Persistence} model is outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $D_{st}$.
+
+\subsection{Assessment of Gaussian Process Models}
+
+The results indicate that there are many benefits to be gained by the application of Gaussian Process models for prediction of $D_{st}$.
+The \emph{GP-AR} and \emph{GP-ARX} not only outperform the models compared in \citet{Ji2012} but also the \emph{Persistence} model.
+Figure \ref{fig:relprob} compares the empirical distributions of absolute relative model errors across all storm events in the experiments.
+If one chooses a $95\%$ probability bound then it can be seen that the \emph{GP-ARX} model bounds the relative error to $20\%$ of the actual $D_{st}$ value, while in the case of the \emph{NM} model it bounds the relative error to $50\%$ of the actual $D_{st}$ value.
+Therefore the \emph{GP-ARX} model gives an approximate benefit of $25\%$ over the other compared models on the test data set chosen.
+
+If one compares the training sets used by the models, one can appreciate that they are able to learn robust models for prediction with relatively small training and validations sets (training set contains 250 instances, while the validation set contains 432 instances) on the other hand the \emph{NM} model uses one year (8760 instances) of data to learn the formula outlined in \citet{balikhin:narmax} and the \emph{TL} model uses six years of data (52560 instances) for training.
+
+The encouraging results of Gaussian Processes illustrate the strengths of the Bayesian approach to predictive modeling.
+Since the GP models generate predictive distributions for test data and not just point predictions they lend themselves to the requirements of space weather prediction very well because of the need to generate error bars on predictions.
 
@@
 
+We conclude this paper with three key conclusions/insights learned from our activities so far.
    \begin{enumerate}
-      \item Persistence model due to \emph{One Step Ahead} restriction 
-      \item Superior performance of \emph{Gaussian Process} NAR and NARX models
-      \item Multi step ahead prediction should be the preferred gold standard with respect to forecasting of geomagnetic indices.
+      \item \emph{Persistence} model must be central in the model evaluation process in the context of \emph{One Step Ahead} prediction of the $D_{st}$ index.
+This is due to its excellent performance in terms of errors and correlations.
+      \item \emph{Gaussian Process} AR and ARX models give encouraring benefits in OSA prediction even when compared to the \emph{Persistence} model.
+Leveraging the strengths of the Bayesian approach, they are able to learn robust predictors from data.
+      \item Multi-step ahead prediction should be the preferred gold standard with respect to forecasting of geomagnetic indices.
+Also important is the usage of more informative performance metrics for evaluating and comparing empirical predictive models.
    \end{enumerate}

### 2016-04-25 swsc.tex
Updates from ShareLaTeX

@@
     \mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*} & \sim & \mathcal{N}(\mathbf{\bar{f}_*}, \Sigma_*)  \label{eq:posterior} \\
-    \mathbf{\bar{f}_*} & = & \mathbf{K}^T_{*} [\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y} \\
-    \Sigma_* & = & \mathbf{K}_{**} - \mathbf{K}^T_{*} \left(\mathbf{K} + \sigma^{2} \mathbf{I}\right)^{-1} \mathbf{K}_{*}
+    \mathbf{\bar{f}_*} & = & \mathbf{K}^T_{*} [\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y} \label{eq:posteriormean} \\
+    \Sigma_* & = & \mathbf{K}_{**} - \mathbf{K}^T_{*} \left(\mathbf{K} + \sigma^{2} \mathbf{I}\right)^{-1} \mathbf{K}_{*} \label{eq:posteriorcov}
 \end{eqnarray}
+
+The distribution of $\mathbf{f_*}| \mathbf{X},\mathbf{y},\mathbf{X_*}$ is known in Bayesian analysis research as the \emph{Posterior Predictive Distribution}.
+This illustrates a key difference between \emph{Gaussian Processes} and other regression models such as \emph{Neural Networks}, \emph{Linear Models} and \emph{Support Vector Machines}, that a \emph{Gaussian Process} model does not generate point predictions for new data but outputs a predictive distribution for the quantity sought, this allows us to construct error bars on the predictions.
+This property of Bayesian models such as \emph{Gaussian Processes} makes them suitable for Space Weather forecasting applications.
 
@@
 
-\begin{figure}[ht]
-   \centering
-   \subfigure{\includegraphics[width=\columnwidth]{Compare_RMSE.png}}
-   \vspace{12pt}
-   \subfigure{\includegraphics[width=\columnwidth]{Compare_CC.png}}
-   \caption{\small  Performance Comparison; left: Averaged RMSE right: Averaged Cross correlation coefficient} 
-   \label{fig:rmsecc}
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Compare_RMSE.png}
+      \caption{Averaged RMSE}
+         \label{fig:rmse}
    \end{figure}
@@
 
+\begin{figure}
+   \centering
+   \subfigure{\includegraphics[width=\columnwidth]{Compare_RelProb.png}}
+   \caption{\small  Comparison of empirical distribution of relative errors for various models on the list of 63 storms in \citet{Ji2012}.} 
+   \label{fig:relprob}
+\end{figure}
+
 Apart from the metrics above, we also generate the empirical distribution of absolute relative errors $|\frac{D_{st} - \hat{D_{st}}}{D_{st}}|$ for the \emph{GP-AR}, \emph{GP-ARX}, \emph{NM} and \emph{TL} models.
@@
 
-\begin{figure}
-   \centering
-   \subfigure{\includegraphics[width=\columnwidth]{Compare_deltaDst.png}}
-   \vspace{12pt}
-   \subfigure{\includegraphics[width=\columnwidth]{Compare_timingerr.png}}
-   \caption{\small  Performance Comparison; left: Averaged $\Delta Dst_{min}$ right: Averaged absolute timing error} 
-   \label{fig:deltaDst}
-   \end{figure}
-
 \subsection{Models Compared}
@@
 In the case of the \emph{NM} model, the formula outlined in \citet{balikhin:narmax} is used to generate hourly predictions.
-For the \emph{TL} model we use real time predictions listed on their website corresponding to all the storm events.
+For the \emph{TL} model we use real time predictions listed on their website (\url{http://lasp.colorado.edu/home/spaceweather/}) corresponding to all the storm events.
 The real time predictions are at a frequency of ten minutes, hourly predictions are generated by averaging the $D_{st}$ predictions for each hour.
@@
 
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Compare_CC.png}
+      \caption{Averaged Cross correlation coefficient}
+         \label{fig:cc}
+   \end{figure}
+
 With respect to the \emph{TL} model one important concern must be expressed.
@@
 
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Compare_deltaDst.png}
+      \caption{Averaged $\Delta Dst_{min}$}
+         \label{fig:deltaDst}
+   \end{figure}
+
 \section{Results} \label{sec:res}
 %%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages
-
-Figures \ref{fig:rmsecc} and \ref{fig:deltaDst} compare the performance of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model with the existing results of \citet{Ji2012}.
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Compare_timingerr.png}
+      \caption{Averaged absolute timing error}
+         \label{fig:timingErr}
+   \end{figure}
+
+Figures \ref{fig:rmse}, \ref{fig:cc}, \ref{fig:deltaDst} and \ref{fig:timingErr} compare the performance of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model with the existing results of \citet{Ji2012}.
 Two central themes come out of these charts.
-
-\begin{figure}
-   \centering
-   \subfigure{\includegraphics[width=\columnwidth]{Compare_RelProb.png}}
-   \caption{\small  Comparison of empirical distribution of relative errors for various models on the list of 63 storms in \citet{Ji2012}.} 
-   \label{fig:relprob}
-\end{figure}
 
@@
 
-If one compares the training sets used by the models, one can appreciate that they are able to learn robust models for prediction with relatively small training and validations sets (training set contains 250 instances, while the validation set contains 432 instances) on the other hand the \emph{NM} model uses one year (8760 instances) of data to learn the formula outlined in \citet{balikhin:narmax} and the \emph{TL} model uses six years of data (52560 instances) for training.
-
-The encouraging results of Gaussian Processes illustrate the strengths of the Bayesian approach to predictive modeling.
-Since the GP models generate predictive distributions for test data and not just point predictions they lend themselves to the requirements of space weather prediction very well because of the need to generate error bars on predictions.
-
 \section{Conclusions}
 
-We conclude this paper with three key conclusions/insights learned from our activities so far.
+We conclude this paper with three key conclusions/insights learned from our research so far.
    \begin{enumerate}
@@
 This is due to its excellent performance in terms of errors and correlations.
-      \item \emph{Gaussian Process} AR and ARX models give encouraring benefits in OSA prediction even when compared to the \emph{Persistence} model.
+      \item \emph{Gaussian Process} AR and ARX models give encouraging benefits in OSA prediction even when compared to the \emph{Persistence} model.
 Leveraging the strengths of the Bayesian approach, they are able to learn robust predictors from data.
+
+      If one compares the training sets used by the models, one can appreciate that they are able to learn robust models for prediction with relatively small training and validations sets (training set contains 250 instances, while the validation set contains 432 instances) on the other hand the \emph{NM} model uses one year (8760 instances) of data to learn the formula outlined in \citet{balikhin:narmax} and the \emph{TL} model uses six years of data (52560 instances) for training.
+
+      The encouraging results of Gaussian Processes illustrate the strengths of the Bayesian approach to predictive modeling.
+Since the GP models generate predictive distributions for test data and not just point predictions they lend themselves to the requirements of space weather prediction very well because of the need to generate error bars on predictions.
+
       \item Multi-step ahead prediction should be the preferred gold standard with respect to forecasting of geomagnetic indices.

### 2016-04-25 swsc.tex
Updates from ShareLaTeX

@@
 
-   %\subtitle{I. Overviewing the $\kappa$-mechanism}
+   \subtitle{Extracting predictive performance over and above persistence behavior}
 
@@
 Figures \ref{fig:rmse}, \ref{fig:cc}, \ref{fig:deltaDst} and \ref{fig:timingErr} compare the performance of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model with the existing results of \citet{Ji2012}.
+Figure \ref{fig:rmse} compares the average RMSE of the models over the list of 63 storms, we can see that the \emph{GP-ARX} model gives an improvement of approximately $38\%$ in RMSE with respect to the \emph{TL} model and $60\%$ improvement with respect to the \emph{NM} model.
+
 Two central themes come out of these charts.

### 2016-04-25 swsc.tex
Updates from ShareLaTeX

@@
    \centering
-   \subfigure{\includegraphics[width=\columnwidth]{Compare_RelProb.png}}
-   \caption{\small  Comparison of empirical distribution of relative errors for various models on the list of 63 storms in \citet{Ji2012}.} 
+   \includegraphics[width=\textwidth]{Compare_RelProb.png}
+   \caption{Comparison of empirical distribution of relative errors for various models on the list of 63 storms in \citet{Ji2012}.}
    \label{fig:relprob}
@@
 
+Figure \ref{fig:relprob} compares the empirical distributions of absolute relative model errors across all storm events in the experiments.
+If one chooses a $95\%$ probability bound then it can be seen that the \emph{GP-ARX} model bounds the relative error to $20\%$ of the actual $D_{st}$ value, while in the case of the \emph{NM} model it bounds the relative error to $50\%$ of the actual $D_{st}$ value.
+Therefore the \emph{GP-ARX} model gives an approximate improvement of $25\%$ over the other compared models on the test data set chosen.
+
+Figure \ref{fig:cc} shows the comparison of correlation coefficients between model predictions and actual $D_{st}$ values.
+From the results of \citet{Ji2012}, the \emph{TL} model gives the highest correlation coefficient of $94\%$, but that is not surprising considering the fact that there is a large overlap between the training data used to train it and storm events.
+This is also the case in \ref{fig:relprob} where we average predictions generated from a pre-trained instance of the \emph{TL} model.
+Even so the \emph{GP-ARX} model gives comparable correlation coefficient to \emph{TL} a testimony to its performance given that its training set has no overlap with the storm events.
+
+Figure \ref{fig:deltaDst} shows how the different models compare with respect to accuracy in predicting the peak value of $D_{st}$ during storm events ($\Delta Dst_{min}$ averaged over all storms).
+In the context of this metric the \emph{GP-ARX} model gives a $31\%$ improvement over the \emph{TL} model and $40\%$ over the \emph{NM} model.
+It is trivial to note that by definition, the \emph{Persistence} model ($\hat{D_{st}}(t) = D_{st}(t-1)$) will have $\Delta Dst_{min} = 0$ for every storm instance.
+
+Figure \ref{fig:timingErr} compares the time discrepancy of predicting the storm peak, again as seen above the \emph{Persistence} model will always have $\Delta t_{peak} = -1$.
 Two central themes come out of these charts.
 
-\subsection{Persistence Behavior}
-
-It is clear that the \emph{persistence} behavior in the $D_{st}$ values is very strong i.e.
+\begin{enumerate}
+    \item Persistence Behavior: It is clear that the \emph{persistence} behavior in the $D_{st}$ values is very strong i.e.
 the trivial predictive model $\hat{D_{st}}(t) = D_{st}(t-1)$ gives excellent performance.
 In fact it can be seen in these figures that the \emph{Persistence} model is outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $D_{st}$.
-
-\subsection{Assessment of Gaussian Process Models}
-
-The results indicate that there are many benefits to be gained by the application of Gaussian Process models for prediction of $D_{st}$.
+    \item Assessment of Gaussian Process Models: The results indicate that there are many benefits to be gained by the application of Gaussian Process models for prediction of $D_{st}$.
 The \emph{GP-AR} and \emph{GP-ARX} not only outperform the models compared in \citet{Ji2012} but also the \emph{Persistence} model.
-Figure \ref{fig:relprob} compares the empirical distributions of absolute relative model errors across all storm events in the experiments.
-If one chooses a $95\%$ probability bound then it can be seen that the \emph{GP-ARX} model bounds the relative error to $20\%$ of the actual $D_{st}$ value, while in the case of the \emph{NM} model it bounds the relative error to $50\%$ of the actual $D_{st}$ value.
-Therefore the \emph{GP-ARX} model gives an approximate benefit of $25\%$ over the other compared models on the test data set chosen.
+\end{enumerate}
 

### 2016-04-25 swsc.tex
Updates from ShareLaTeX

@@
 Each geomagnetic index is a proxy for a particular kind of phenomena.
-Some popular indices are.
-
+Some popular indices are the $K_p$, $D_{st}$ and the AE index.
+
+\begin{comment}
 \begin{enumerate}
@@
 \end{enumerate}
+\end{comment}
 
 We focus on prediction of the hourly $D_{st}$ index which is an important problem because $D_{st}$ is a straight forward indicator of the occurrence of geomagnetic storms.
-More specifically we focus on the \emph{one step ahead} prediction of $D_{st}$ because it is the first step towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions and because it has been the subject of several modeling efforts in the past.
+More specifically we focus on the \emph{one step ahead} (OSA) prediction of $D_{st}$ because it is the first step towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions.
+This problem has been the subject of several modeling efforts so far, some of the earliest are by \citet{JGR:JGR10260} who outlined a prediction framework that has been used as a base line for many $D_{st}$ prediction models.
+\citet{JGR:JGR10260} model $D_{st}(t)$ a function of time using an \emph{Ordinary Differential Equation} (ODE) which modeled it's rate of change as a combination of two terms: decay and injection $\frac{d D_{st}(t)}{dt} = Q(t) - \frac{D_{st}(t)}{\tau}$.
+The \citet{JGR:JGR10260} model has proven to be very influential in the Space Weather research effort and many subsequent works like \citet{Wang:Dst}, \citet{JGRA:JGRA14856} modified the proposed ODE by defining expressions for the injection term $Q(t)$.
 
@@
 In fact it can be seen in these figures that the \emph{Persistence} model is outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $D_{st}$.
+Therefore any model that is proposed to tackle the OSA prediction problem for $D_{st}$ must be compared to and show visible gains over the \emph{Persistence} model.
     \item Assessment of Gaussian Process Models: The results indicate that there are many benefits to be gained by the application of Gaussian Process models for prediction of $D_{st}$.

### 2016-04-26 swsc.tex
Updates from ShareLaTeX

@@
 
+%Talk about Burton and friends
 We focus on prediction of the hourly $D_{st}$ index which is an important problem because $D_{st}$ is a straight forward indicator of the occurrence of geomagnetic storms.
 More specifically we focus on the \emph{one step ahead} (OSA) prediction of $D_{st}$ because it is the first step towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions.
-This problem has been the subject of several modeling efforts so far, some of the earliest are by \citet{JGR:JGR10260} who outlined a prediction framework that has been used as a base line for many $D_{st}$ prediction models.
+
+The $D_{st}$ OSA prediction problem has been the subject of several modeling efforts so far, some of the earliest are by \citet{JGR:JGR10260} who outlined a prediction framework that has been used as a base line for many $D_{st}$ models.
 \citet{JGR:JGR10260} model $D_{st}(t)$ a function of time using an \emph{Ordinary Differential Equation} (ODE) which modeled it's rate of change as a combination of two terms: decay and injection $\frac{d D_{st}(t)}{dt} = Q(t) - \frac{D_{st}(t)}{\tau}$.
+
 The \citet{JGR:JGR10260} model has proven to be very influential in the Space Weather research effort and many subsequent works like \citet{Wang:Dst}, \citet{JGRA:JGRA14856} modified the proposed ODE by defining expressions for the injection term $Q(t)$.
-
+More recently \citet{Ballatore2014} have tried to generate estimates for the injection and decay terms in Burton's equation.
+
+%Talk about NARMAX Dst
+Among the other major empirical models used to predict $D_{st}$ has been the \emph{Nonlinear Auto-Regessive Moving Average with eXogenous inputs} (NARMAX) methodology developed in \citet{doi:10.1080/00207178908559767}, \citet{GRL:GRL13494}, \citet{GRL:GRL20944}, \citet{JGRA:JGRA18657}, \citet{JGRA:JGRA20661}, \citet{JGRA:JGRA50192} and \citet{balikhin:narmax}.
+The NARMAX methodology builds models by constructing polynomial expansions of inputs and determines the best combinations or monomials to include in the refined model using the \emph{error reduction ratio} (ERR) criterion.
+The parameters of the NARMAX OLS-ERR model are calculated using the standard \emph{ordinary least squares} (OLS) method hence these methods are also denoted by the acronym NARMAX OLS-ERR.
+The reader may refer to \citet{billings2013nonlinear} for a detailed exposition on the NARMAX methodology.
+
+%Talk about neural networks
+Application of \emph{Machine Learning} methods for prediction of geomagnetic indices is fairly recent (1990's onwards).
+\emph{Neural Networks} have been a popular choice for building predictive models and researchers employed the standard \emph{feed forward} and the more specialized \emph{recurrent} architectures.
+\citet{Lund} proposed an \emph{Elman} recurrent network architecture for prediction of $D_{st}$.
+\citet{SWE:SWE286} originally proposed a \emph{feed forward} network for predicting the $K_p$ index which employed the \emph{Boyle coupling function} \citet{boyle1997empirical}, in \citet{SWE:SWE286} the same architecture is adapted for prediction of $D_{st}$ and $AE$ indices as well.
+
+%Talk about comparison papers, number of storms covered and Persistence
+
+% Talk about Gaussian Process
 We use the hourly resolution Omni data set (see \cite{OmniPaper}) as extracted from NASA/GSFC's OMNI data set through OMNIWeb.

### 2016-04-26 swsc.tex
Updates from ShareLaTeX

@@
 %Talk about Burton and friends
-We focus on prediction of the hourly $D_{st}$ index which is an important problem because $D_{st}$ is a straight forward indicator of the occurrence of geomagnetic storms.
+We focus on prediction of the hourly $D_{st}$ index which is an important problem because $D_{st}$ is a straight forward indicator of geomagnetic storms.
 More specifically we focus on the \emph{one step ahead} (OSA) prediction of $D_{st}$ because it is the first step towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions.
@@
 %Talk about neural networks
-Application of \emph{Machine Learning} methods for prediction of geomagnetic indices is fairly recent (1990's onwards).
 \emph{Neural Networks} have been a popular choice for building predictive models and researchers employed the standard \emph{feed forward} and the more specialized \emph{recurrent} architectures.
-\citet{Lund} proposed an \emph{Elman} recurrent network architecture for prediction of $D_{st}$.
-\citet{SWE:SWE286} originally proposed a \emph{feed forward} network for predicting the $K_p$ index which employed the \emph{Boyle coupling function} \citet{boyle1997empirical}, in \citet{SWE:SWE286} the same architecture is adapted for prediction of $D_{st}$ and $AE$ indices as well.
+\citet{Lund} proposed an \emph{Elman} recurrent network architecture called Lund $D_{st}$, which used the solar wind, \emph{interplanetary magnetic field} (IMF) and historical $D_{st}$ data.
+\citet{SWE:SWE286} originally proposed a \emph{feed forward} network for predicting the $K_p$ index which used the \emph{Boyle coupling function} (\citet{boyle1997empirical}), in \citet{SWE:SWE286} the same architecture is adapted for prediction of $D_{st}$, which is popularly known as the Rice $D_{st}$ model.
+\citet{pallocchia:hal-00318011} proposed a \emph{neural network} model called EDDA to predict $D_{st}$ using only the IMF data.
 
 %Talk about comparison papers, number of storms covered and Persistence
+In light of the extensive list of modeling techniques employed for prediction of geomagnetic indices and in particular $D_{st}$, model comparison and evaluation becomes a crucial direction in advancing the research domain.
+\citet{Rastatter2013} compare several physics based (convection, kinetic and magneto hydrodynamic) and empirical prediction models such as Narmax, Rice $D_{st}$ on 4 storm events which occurred between 2001 and 2006.
+\citet{Amata2008} compares the EDDA and the Lund $D_{st}$ models over the 2003-2005 period.
+But the most extensive model comparison in terms of storm events was conducted in \citet{Ji2012} which compared seven $D_{st}$ models (see table \ref{table:DstModels}) on a list of 63 geomagnetic storm events of varying intensities which occurred between 1998 and 2006.
+
+%Vision
+In $D_{st}$ prediction a seemingly trivial yet highly informative prediction method known as the \emph{Persistence} model has been largely ignored in its crucial role in model evaluation.
+The \emph{Persistence} model is stated simply as using the previous value of $D_{st}$ as the prediction for the next time step ($\hat{D_{st}}(t) = D_{st}(t-1)$).
+Due to high auto correlation between $D_{st}(t)$ and $D_{st}(t-1)$ in the Omni data, the \emph{Persistence} model gives excellent predictive performance despite its apparent lack of sophistication.
+In essence it is a trivial hypothesis and hence must always be used as a base line to compare the performance of any proposed $D_{st}$ algorithm.
+Therefore in the context of OSA prediction, a model must outperform the \emph{Persistence} model in order to be a viable candidate for advancing the science of space weather prediction.
+
+%Assertion
+It is with the above vision in mind, we propose \emph{Gaussian Process} models for OSA prediction of $D_{st}$.
+We use the results of \citet{Ji2012} as a starting point and compare our proposed models with the methods outlined in the same paper while using the performance of the \emph{Persistence} model as a litmus test.
 

### 2016-04-26 swsc.tex
Updates from ShareLaTeX

@@
 
-   \subtitle{Extracting predictive performance over and above persistence behavior}
+   \subtitle{Attaining performance over and above persistence behavior}
 
@@
 %Talk about Burton and friends
-We focus on prediction of the hourly $D_{st}$ index which is an important problem because $D_{st}$ is a straight forward indicator of geomagnetic storms.
+We focus on prediction of the hourly $D_{st}$ index which is an important problem because $D_{st}$ is a straightforward indicator of geomagnetic storms.
 More specifically we focus on the \emph{one step ahead} (OSA) prediction of $D_{st}$ because it is the first step towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions.
 
-The $D_{st}$ OSA prediction problem has been the subject of several modeling efforts so far, some of the earliest are by \citet{JGR:JGR10260} who outlined a prediction framework that has been used as a base line for many $D_{st}$ models.
-\citet{JGR:JGR10260} model $D_{st}(t)$ a function of time using an \emph{Ordinary Differential Equation} (ODE) which modeled it's rate of change as a combination of two terms: decay and injection $\frac{d D_{st}(t)}{dt} = Q(t) - \frac{D_{st}(t)}{\tau}$.
-
-The \citet{JGR:JGR10260} model has proven to be very influential in the Space Weather research effort and many subsequent works like \citet{Wang:Dst}, \citet{JGRA:JGRA14856} modified the proposed ODE by defining expressions for the injection term $Q(t)$.
-More recently \citet{Ballatore2014} have tried to generate estimates for the injection and decay terms in Burton's equation.
+The $D_{st}$ OSA prediction problem has been the subject of several modeling efforts so far, some of the earliest are by \citet{JGR:JGR10260} who outlined a prediction framework that has been used as a starting point for many $D_{st}$ models.
+\citet{JGR:JGR10260} model $D_{st}(t)$ a function of time using an \emph{Ordinary Differential Equation} (ODE) which expressed it's rate of change as a combination of two terms: decay and injection $\frac{d D_{st}(t)}{dt} = Q(t) - \frac{D_{st}(t)}{\tau}$.
+
+The \citet{JGR:JGR10260} model has proven to be very influential in the Space Weather research effort and many subsequent works like \citet{Wang:Dst}, \citet{JGRA:JGRA14856} modified the proposed ODE by proposing expressions for the injection term $Q(t)$.
+More recently \citet{Ballatore2014} have tried to generate empirical estimates for the injection and decay terms in Burton's equation.
 
 %Talk about NARMAX Dst
-Among the other major empirical models used to predict $D_{st}$ has been the \emph{Nonlinear Auto-Regessive Moving Average with eXogenous inputs} (NARMAX) methodology developed in \citet{doi:10.1080/00207178908559767}, \citet{GRL:GRL13494}, \citet{GRL:GRL20944}, \citet{JGRA:JGRA18657}, \citet{JGRA:JGRA20661}, \citet{JGRA:JGRA50192} and \citet{balikhin:narmax}.
+Among the other major empirical models used to predict $D_{st}$ has been the \emph{Nonlinear Auto-Regessive Moving Average with eXogenous inputs} (NARMAX) methodology developed in \citet{doi:10.1080/00207178908559767}, \citet{GRL:GRL13494}, \citet{GRL:GRL20944}, \citet{JGRA:JGRA18657}, \citet{balikhin:narmax}, \citet{JGRA:JGRA20661} and \citet{JGRA:JGRA50192}.
 The NARMAX methodology builds models by constructing polynomial expansions of inputs and determines the best combinations or monomials to include in the refined model using the \emph{error reduction ratio} (ERR) criterion.
 The parameters of the NARMAX OLS-ERR model are calculated using the standard \emph{ordinary least squares} (OLS) method hence these methods are also denoted by the acronym NARMAX OLS-ERR.
-The reader may refer to \citet{billings2013nonlinear} for a detailed exposition on the NARMAX methodology.
+The reader may refer to \citet{billings2013nonlinear} for a detailed exposition of the NARMAX methodology.
 
@@
 %Assertion
-It is with the above vision in mind, we propose \emph{Gaussian Process} models for OSA prediction of $D_{st}$.
+It is with the above observation in mind, we propose \emph{Gaussian Process} models for OSA prediction of $D_{st}$.
 We use the results of \citet{Ji2012} as a starting point and compare our proposed models with the methods outlined in the same paper while using the performance of the \emph{Persistence} model as a litmus test.
@@
 Section \ref{sec:method} gives an introduction to the \emph{Gaussian Process} methodology for regression and the various issues surrounding their application.
-In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res} bench marked against the \emph{Persistence} $D_{st}$ model as well as other models outlined in \citet{Ji2012}.
+In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res}, bench marked against the \emph{Persistence} $D_{st}$ model as well as other models outlined in \citet{Ji2012}.
 
@@
 
-\emph{Gaussian Processes} first appeared in machine learning research as the limiting case of Bayesian inference performed on neural networks with infinitely many neurons in the hidden layers \citet{Neal:1996:BLN:525544}.
+\emph{Gaussian Processes} first appeared in machine learning research in \citet{Neal:1996:BLN:525544}, as the limiting case of Bayesian inference performed on neural networks with infinitely many neurons in the hidden layers.
 Although their inception in the machine learning community is recent, their origins can be traced back to the geo-statistics research community where they are known as \emph{Kriging} methods (\citet{krige1951statistical}).
 In pure mathematics area \emph{Gaussian Processes} have been studied extensively and their existence was first proven by Kolmogorov's extension theorem (\citet{tao2011introduction}).
-The reader is requested to refer to \cite{Rasmussen:2005:GPM:1162254} for an in depth treatment of Gaussian Processes in Machine Learning.
+The reader is requested to refer to \cite{Rasmussen:2005:GPM:1162254} for an in depth treatment of Gaussian Processes in machine learning.
 

### 2016-05-02 swsc.tex
Updates from ShareLaTeX

@@
 
-\emph{Gaussian Processes} first appeared in machine learning research in \citet{Neal:1996:BLN:525544}, as the limiting case of Bayesian inference performed on neural networks with infinitely many neurons in the hidden layers.
-Although their inception in the machine learning community is recent, their origins can be traced back to the geo-statistics research community where they are known as \emph{Kriging} methods (\citet{krige1951statistical}).
-In pure mathematics area \emph{Gaussian Processes} have been studied extensively and their existence was first proven by Kolmogorov's extension theorem (\citet{tao2011introduction}).
-The reader is requested to refer to \cite{Rasmussen:2005:GPM:1162254} for an in depth treatment of Gaussian Processes in machine learning.
-
 A \emph{Gaussian Process} is a stochastic process $f(\mathbf{x}), \ \mathbf{x} \in \mathbb{R}^d$ whose finite dimensional distributions over a set of points $\mathbf{x}_1, \cdots, \mathbf{x}_n$ are multivariate Gaussian.
@@
 
+\emph{Gaussian Processes} first appeared in machine learning research in \citet{Neal:1996:BLN:525544}, as the limiting case of Bayesian inference performed on neural networks with infinitely many neurons in the hidden layers.
+Although their inception in the machine learning community is recent, their origins can be traced back to the geo-statistics research community where they are known as \emph{Kriging} methods (\citet{krige1951statistical}).
+In pure mathematics area \emph{Gaussian Processes} have been studied extensively and their existence was first proven by Kolmogorov's extension theorem (\citet{tao2011introduction}).
+The reader is requested to refer to \cite{Rasmussen:2005:GPM:1162254} for an in depth treatment of Gaussian Processes in machine learning.
+
 \subsection{Gaussian Process Regression: Formulation}
@@
 
-The distribution of $\mathbf{f_*}| \mathbf{X},\mathbf{y},\mathbf{X_*}$ is known in Bayesian analysis research as the \emph{Posterior Predictive Distribution}.
+The distribution of $\mathbf{f_*}| \mathbf{X},\mathbf{y},\mathbf{X_*}$ is known in Bayesian analysis as the \emph{Posterior Predictive Distribution}.
 This illustrates a key difference between \emph{Gaussian Processes} and other regression models such as \emph{Neural Networks}, \emph{Linear Models} and \emph{Support Vector Machines}, that a \emph{Gaussian Process} model does not generate point predictions for new data but outputs a predictive distribution for the quantity sought, this allows us to construct error bars on the predictions.
@@
 
-Figure \ref{fig:timingErr} compares the time discrepancy of predicting the storm peak, again as seen above the \emph{Persistence} model will always have $\Delta t_{peak} = -1$.
-Two central themes come out of these charts.
-
+Figure \ref{fig:timingErr} compares the time discrepancy of predicting the storm peak, again as observed above the \emph{Persistence} model will always have $\Delta t_{peak} = -1$, the GP-AR and GP-ARX models give better performance than the other models in terms of timing error in prediction of storm peaks.
+
+In figure \ref{fig:predictions} we show the OSA predictions generated by the TL, NM and GP-ARX models for one event from the list of storms in \citet{Ji2012}.
+This storm is interesting due to its strength ($Dst_{min} = -289 nT$) and the presence of two distinct peaks.
+In this particular event the GP-ARX model approximates the twin peaks as well as the onset and decay phases quite faithfully which is lacking in the NM and TL model predictions.
+The TL model recognises the presence of twin peaks but over estimates the second one as well as the decay phase, while the NM model is much delayed in the prediction of the initial peak and fails to approximate the time and intensity of the second peak.
+
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Compare_pred.png}
+      \caption{Comparison of OSA predictions generated by the NM, TL and GP-ARX models for the storm event $9^{th}$ November $2004$ 11:00 UTC - $11^{th}$ November $2004$ 09:00 UTC}
+         \label{fig:predictions}
+   \end{figure}
+
+\begin{comment}
 \begin{enumerate}
@@
 \end{enumerate}
+\end{comment}
 
@@
       \item \emph{Persistence} model must be central in the model evaluation process in the context of \emph{One Step Ahead} prediction of the $D_{st}$ index.
-This is due to its excellent performance in terms of errors and correlations.
+It is clear that the \emph{persistence} behavior in the $D_{st}$ values is very strong i.e.
+the trivial predictive model $\hat{D_{st}}(t) = D_{st}(t-1)$ gives excellent performance.
+In fact it can be seen in \ref{sec:res} that the \emph{Persistence} model is outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $D_{st}$.
+Therefore any model that is proposed to tackle the OSA prediction problem for $D_{st}$ must be compared to the \emph{Persistence} model and show visible gains above it.
+
       \item \emph{Gaussian Process} AR and ARX models give encouraging benefits in OSA prediction even when compared to the \emph{Persistence} model.

### 2016-05-02 swsc.tex
Updates from ShareLaTeX

@@
 The interested reader is suggested to refer to \cite{Berlinet2004}, \citet{Scholkopf:2001:LKS:559923} and \citet{hofmann2008} for a thorough treatment on kernel functions and the rich theory behind them.
-Some common kernel functions used in machine learning are.
-
+Some common kernel functions used in machine learning include the radial basis function (RBF) kernel $K(\mathbf{x}, \mathbf{y}) = \frac{1}{2} exp(-||\mathbf{x} - \mathbf{y}||^2/l^2)$ and the polynomial kernel $K(\mathbf{x}, \mathbf{y}) = (\mathbf{x}^\intercal \mathbf{y} + b)^d$.
+
+The quantities $l$ in the RBF, $b$ and $d$ in the polynomial kernel are known as \emph{hyper-parameters}.
+Hyper-parameters give flexibility to particular kernel structure, for example $d = 1, 2, 3, \cdots$ in the polynomial kernel represents linear, quadratic, cubic and higher order phenomena respectively.
+The method of assigning values to the \emph{hyper-parameter} is crucial in the model building process (see section \ref{sec:hyp} below).
+In this study, we construct Gaussian Process regression models with polynomial kernels for \emph{one step ahead} prediction of the $D_{st}$ index.
+
+\begin{comment}
 \subsubsection{Radial Basis Function (RBF)}
@@
 
-In this study, we construct Gaussian Process regression models with polynomial kernels for \emph{one step ahead} prediction of the $D_{st}$ index.
+\end{comment}
 
@@
 Before performance bench marks for \emph{OSA} $D_{st}$ prediction are run on the storm events in \cite{Ji2012}, training and model selection of \emph{GP-AR} and \emph{GP-ARX} models on independent data sets must be performed.
-For this purpose we choose segments 00:00 1 January 2008 - 10:00 11 January 2008 and 00:00 15 November 2014 - 23:00 1 December 2014 for training and model selection respectively.
+For this purpose we choose segments 00:00 1 January 2008 - 10:00 11 January 2008 for training and 00:00 15 November 2014 - 23:00 1 December 2014 for model selection.
 

### 2016-05-03 swsc.tex
Updates from ShareLaTeX

@@
 
-In most practical applications of \emph{Gaussian Processes} the mean function is often defined as $m(\mathbf{x}) = 0$, which is not unreasonable if the data is standardized to have zero mean.
-The central design issue now becomes choice of the function $K(\mathbf{x}, \mathbf{x}')$.
-The same constraints that apply to $\mathbf{\Lambda}$ also apply to the function $K(.,.)$.
-In machine learning, these symmetric positive definite functions of two variables are known as \emph{kernels}.
-Kernel based methods are applied extensively in data analysis i.e.
-regression, clustering, classification, density estimation (see \citet{Scholkopf:2001:LKS:559923}, \citet{hofmann2008}).
-
-\emph{Gaussian Processes} first appeared in machine learning research in \citet{Neal:1996:BLN:525544}, as the limiting case of Bayesian inference performed on neural networks with infinitely many neurons in the hidden layers.
-Although their inception in the machine learning community is recent, their origins can be traced back to the geo-statistics research community where they are known as \emph{Kriging} methods (\citet{krige1951statistical}).
-In pure mathematics area \emph{Gaussian Processes} have been studied extensively and their existence was first proven by Kolmogorov's extension theorem (\citet{tao2011introduction}).
-The reader is requested to refer to \cite{Rasmussen:2005:GPM:1162254} for an in depth treatment of Gaussian Processes in machine learning.
-
 \subsection{Gaussian Process Regression: Formulation}
@@
 
+\emph{Gaussian Processes} first appeared in machine learning research in \citet{Neal:1996:BLN:525544}, as the limiting case of Bayesian inference performed on neural networks with infinitely many neurons in the hidden layers.
+Although their inception in the machine learning community is recent, their origins can be traced back to the geo-statistics research community where they are known as \emph{Kriging} methods (\citet{krige1951statistical}).
+In pure mathematics area \emph{Gaussian Processes} have been studied extensively and their existence was first proven by Kolmogorov's extension theorem (\citet{tao2011introduction}).
+The reader is requested to refer to \cite{Rasmussen:2005:GPM:1162254} for an in depth treatment of Gaussian Processes in machine learning.
+
+In most practical applications of \emph{Gaussian Processes} the mean function is often defined as $m(\mathbf{x}) = 0$, which is not unreasonable if the data is standardized to have zero mean.
+The central design issue now becomes choice of the function $K(\mathbf{x}, \mathbf{x}')$.
+The same constraints that apply to $\mathbf{\Lambda}$ also apply to the function $K(.,.)$.
+In machine learning, these symmetric positive definite functions of two variables are known as \emph{kernels}.
+Kernel based methods are applied extensively in data analysis i.e.
+regression, clustering, classification, density estimation (see \citet{Scholkopf:2001:LKS:559923}, \citet{hofmann2008}).
+
 \subsection{Kernel Functions}
@@
    \includegraphics[width=\textwidth]{Compare_RMSE.png}
-      \caption{Averaged RMSE}
+      \caption{Averaged RMSE performance of GP-AR, GP-ARX and the Persistence model versus results in \citet{Ji2012}.}
          \label{fig:rmse}
@@
    \includegraphics[width=\textwidth]{Compare_RelProb.png}
-   \caption{Comparison of empirical distribution of relative errors for various models on the list of 63 storms in \citet{Ji2012}.}
+   \caption{Comparison of empirical distribution of relative errors for GP-AR, GP-ARX, NM and TL on the test set of \citet{Ji2012}.}
    \label{fig:relprob}
@@
    \includegraphics[width=\textwidth]{Compare_CC.png}
-      \caption{Averaged Cross correlation coefficient}
+      \caption{Averaged cross correlation performance of GP-AR, GP-ARX and the Persistence model versus results in \citet{Ji2012}.}
          \label{fig:cc}
@@
    \includegraphics[width=\textwidth]{Compare_deltaDst.png}
-      \caption{Averaged $\Delta Dst_{min}$}
+      \caption{Averaged $\Delta Dst_{min}$ of GP-AR, GP-ARX and the Persistence model versus results in \citet{Ji2012}.}
          \label{fig:deltaDst}
@@
 %%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages
-\begin{figure}
-   \centering
-   \includegraphics[width=\textwidth]{Compare_timingerr.png}
-      \caption{Averaged absolute timing error}
-         \label{fig:timingErr}
-   \end{figure}
 
@@
 Therefore the \emph{GP-ARX} model gives an approximate improvement of $25\%$ over the other compared models on the test data set chosen.
+
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Compare_timingerr.png}
+      \caption{Averaged absolute timing error of GP-AR, GP-ARX and the Persistence model versus results in \citet{Ji2012}.}
+         \label{fig:timingErr}
+   \end{figure}
 
@@
 
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Compare_pred.png}
+      \caption{Comparison of OSA predictions generated by the NM, TL and GP-ARX models for the storm event $9^{th}$ November $2004$ 11:00 UTC - $11^{th}$ November $2004$ 09:00 UTC}
+         \label{fig:predictions}
+\end{figure}
+
 Figure \ref{fig:deltaDst} shows how the different models compare with respect to accuracy in predicting the peak value of $D_{st}$ during storm events ($\Delta Dst_{min}$ averaged over all storms).
@@
    \centering
-   \includegraphics[width=\textwidth]{Compare_pred.png}
-      \caption{Comparison of OSA predictions generated by the NM, TL and GP-ARX models for the storm event $9^{th}$ November $2004$ 11:00 UTC - $11^{th}$ November $2004$ 09:00 UTC}
-         \label{fig:predictions}
+   \includegraphics[width=\textwidth]{Compare_pred_err_bar.png}
+      \caption{OSA predictions with error bars $\pm 1$ standard deviation, generated by the GP-ARX model for the storm event $9^{th}$ November $2004$ 11:00 UTC - $11^{th}$ November $2004$ 09:00 UTC}
+         \label{fig:predictionswitherrorbar}
 \end{figure}

### 2016-05-04 swsc.tex
Updates from ShareLaTeX

@@
 
-\subsection{Inference and Predictions}
+\subsection{Inference and Predictions} \label{sec:inference}
 
@@
 The TL model recognises the presence of twin peaks but over estimates the second one as well as the decay phase, while the NM model is much delayed in the prediction of the initial peak and fails to approximate the time and intensity of the second peak.
+
+As noted in section \ref{sec:inference}, \emph{Gaussian Process} models generate probabilistic predictions instead of point predictions, giving the modeller the ability to specify error bars on the predictions.
+In figure \ref{fig:predictionswitherrorbar} OSA predictions along with lower and upper error bars generated by the GP-ARX model are depicted for the event used as a reference in figure \ref{fig:predictions}.
+The error bars are calculated at one standard deviation on either side of the mean prediction, using the posterior predictive distribution in equations \ref{eq:posteriormean} and \ref{eq:posteriorcov}.
 

### 2016-05-18 swsc.tex
Changes May 18th

@@
 
-   \title{Gaussian Processes Autoregressive Models for the Disturbance Storm Time Index}
-
-   \subtitle{Attaining performance over and above persistence behavior}
-
-   \titlerunning{Gaussian Process $D_{st}$ models}
+   \title{Gaussian Processes Autoregressive Models for Forecasting the Disturbance Storm Time Index}
+
+   \titlerunning{Gaussian Process $Dst$ models}
 
@@
    \abstract{
-    We aim to improve the state of the art of \emph{One Step Ahead}(OSA) prediction of the $D_{st}$ geomagnetic activity index.
-In the domain of OSA prediction, it is well known that $D_{st}(t)$ values correlate strongly with values one hour in the past (i.e.
-$D_{st}(t-1)$).
+    We present two models that improve the state of the art of \emph{One Step Ahead}(OSA) prediction of the $Dst$ geomagnetic activity index.
+In the domain of OSA prediction, it is well known that $Dst(t)$ values correlate strongly with values one hour in the past (i.e.
+$Dst(t-1)$).
 While a number of forecasting models have been proposed and compared, none of them have been compared with the \emph{Persistence} model to quantify their performance improvements above this most trivial of predictors.
-In this study we train Gaussian Process regression models for One Step Ahead (OSE) prediction of the Disturbance Storm Time ($D_{st}$) geomagnetic index.
-We propose three variants of the Gaussian Process model.
+In this study we train Gaussian Process regression models for One Step Ahead (OSE) prediction of the Disturbance Storm Time ($Dst$) geomagnetic index.
+We propose two variants of the Gaussian Process model.
     \begin{enumerate*}
@@
     \end{enumerate*}
-    We compare the performance of these models with the current state of the art in one step ahead $D_{st}$ prediction on a set of 
+    We compare the performance of these models with the current state of the art in one step ahead $Dst$ prediction on a set of 
     63 benchmark storms from 1998-2006. The proposed Gaussian Process models not only outperform the rest but also the \emph{Persistence} model.
@@
    \keywords{Geomagnetic indices --
-            $D_{st}$ OSA Prediction --
+            $Dst$ OSA Prediction --
             Gaussian Processes --
@@
 
-The Sun is the principal driver of Space Weather, the steady solar wind and intermittent bursts of hot plasma from its corona (Coronal Mass Ejections \emph{CME}) are the causes of several observed phenomena in the Earths Magnetosphere such as but not limited to \emph{Aurora Borealis}, \emph{Aurora Australis}, \emph{Van Allen Currents}, geomagnetic storms etc.
-
-Although the qualitative bifurcation of these phenomena is well understood and each has been studied in depth, the magnetosphere's dynamics and its associated solar wind driver form a complex nonlinear coupled dynamical system.
+The magnetosphere's dynamics and its associated solar wind driver form a complex nonlinear coupled dynamical system.
 It is therefore not only instructive but also greatly simplifying to use representative indices to quantify the state of geomagnetic activity.
 
-Geomagnetic indices come in various forms, they may take continuous or discreet values.
+Geomagnetic indices come in various forms, they may take continuous or discrete values.
 Their values are often calculated by averaging or combining a number of readings taken by instruments around the Earth.
 Each geomagnetic index is a proxy for a particular kind of phenomena.
-Some popular indices are the $K_p$, $D_{st}$ and the AE index.
+Some popular indices are the $K_p$, $Dst$ and the $AE$ index.
 
@@
 \begin{enumerate}
-    \item $K_p$: The Kp-index is a discreet valued global geomagnetic storm index and is based on 3 hour measurements of the K-indices, for a given value, for each of the past days.
+    \item $K_p$: The Kp-index is a discrete valued global geomagnetic storm index and is based on 3 hour measurements of the K-indices, for a given value, for each of the past days.
 The K-index itself is a three hour long quasi-logarithmic local index of the geomagnetic activity, relative to a calm day curve for the given location.
@@
 
-    \item $D_{st}$: A continuous hourly index which measures the weakening of the Earths magnetic field due to ring currents. 
+    \item $Dst$: A continuous hourly index which measures the weakening of the Earths magnetic field due to ring currents. 
 \end{enumerate}
@@
 %Talk about Burton and friends
-We focus on prediction of the hourly $D_{st}$ index which is an important problem because $D_{st}$ is a straightforward indicator of geomagnetic storms.
-More specifically we focus on the \emph{one step ahead} (OSA) prediction of $D_{st}$ because it is the first step towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions.
-
-The $D_{st}$ OSA prediction problem has been the subject of several modeling efforts so far, some of the earliest are by \citet{JGR:JGR10260} who outlined a prediction framework that has been used as a starting point for many $D_{st}$ models.
-\citet{JGR:JGR10260} model $D_{st}(t)$ a function of time using an \emph{Ordinary Differential Equation} (ODE) which expressed it's rate of change as a combination of two terms: decay and injection $\frac{d D_{st}(t)}{dt} = Q(t) - \frac{D_{st}(t)}{\tau}$.
+We focus on prediction of the hourly $Dst$ index which is an important problem because $Dst$ is a straightforward indicator of geomagnetic storms.
+More specifically we focus on the \emph{one step ahead} (OSA) prediction of $Dst$ because it is the first step towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions.
+
+The $Dst$ OSA prediction problem has been the subject of several modeling efforts so far, some of the earliest are by \citet{JGR:JGR10260} who outlined a prediction framework that has been used as a starting point for many $Dst$ models.
+\citet{JGR:JGR10260} model $Dst(t)$ a function of time using an \emph{Ordinary Differential Equation} (ODE) which expressed it's rate of change as a combination of two terms: decay and injection $\frac{d Dst(t)}{dt} = Q(t) - \frac{Dst(t)}{\tau}$.
 
@@
 %Talk about NARMAX Dst
-Among the other major empirical models used to predict $D_{st}$ has been the \emph{Nonlinear Auto-Regessive Moving Average with eXogenous inputs} (NARMAX) methodology developed in \citet{doi:10.1080/00207178908559767}, \citet{GRL:GRL13494}, \citet{GRL:GRL20944}, \citet{JGRA:JGRA18657}, \citet{balikhin:narmax}, \citet{JGRA:JGRA20661} and \citet{JGRA:JGRA50192}.
+Among the other major empirical models used to predict $Dst$ has been the \emph{Nonlinear Auto-Regessive Moving Average with eXogenous inputs} (NARMAX) methodology developed in \citet{doi:10.1080/00207178908559767}, \citet{GRL:GRL13494}, \citet{GRL:GRL20944}, \citet{JGRA:JGRA18657}, \citet{balikhin:narmax}, \citet{JGRA:JGRA20661} and \citet{JGRA:JGRA50192}.
 The NARMAX methodology builds models by constructing polynomial expansions of inputs and determines the best combinations or monomials to include in the refined model using the \emph{error reduction ratio} (ERR) criterion.
@@
 \emph{Neural Networks} have been a popular choice for building predictive models and researchers employed the standard \emph{feed forward} and the more specialized \emph{recurrent} architectures.
-\citet{Lund} proposed an \emph{Elman} recurrent network architecture called Lund $D_{st}$, which used the solar wind, \emph{interplanetary magnetic field} (IMF) and historical $D_{st}$ data.
-\citet{SWE:SWE286} originally proposed a \emph{feed forward} network for predicting the $K_p$ index which used the \emph{Boyle coupling function} (\citet{boyle1997empirical}), in \citet{SWE:SWE286} the same architecture is adapted for prediction of $D_{st}$, which is popularly known as the Rice $D_{st}$ model.
-\citet{pallocchia:hal-00318011} proposed a \emph{neural network} model called EDDA to predict $D_{st}$ using only the IMF data.
+\citet{Lund} proposed an \emph{Elman} recurrent network architecture called Lund $Dst$, which used the solar wind, \emph{interplanetary magnetic field} (IMF) and historical $Dst$ data.
+\citet{SWE:SWE286} originally proposed a \emph{feed forward} network for predicting the $K_p$ index which used the \emph{Boyle coupling function} (\citet{boyle1997empirical}), in \citet{SWE:SWE286} the same architecture is adapted for prediction of $Dst$, which is popularly known as the Rice $Dst$ model.
+\citet{pallocchia:hal-00318011} proposed a \emph{neural network} model called EDDA to predict $Dst$ using only the IMF data.
 
 %Talk about comparison papers, number of storms covered and Persistence
-In light of the extensive list of modeling techniques employed for prediction of geomagnetic indices and in particular $D_{st}$, model comparison and evaluation becomes a crucial direction in advancing the research domain.
-\citet{Rastatter2013} compare several physics based (convection, kinetic and magneto hydrodynamic) and empirical prediction models such as Narmax, Rice $D_{st}$ on 4 storm events which occurred between 2001 and 2006.
-\citet{Amata2008} compares the EDDA and the Lund $D_{st}$ models over the 2003-2005 period.
-But the most extensive model comparison in terms of storm events was conducted in \citet{Ji2012} which compared seven $D_{st}$ models (see table \ref{table:DstModels}) on a list of 63 geomagnetic storm events of varying intensities which occurred between 1998 and 2006.
+In light of the extensive list of modeling techniques employed for prediction of geomagnetic indices and in particular $Dst$, model comparison and evaluation becomes a crucial direction in advancing the research domain.
+\citet{Rastatter2013} compare several physics based (convection, kinetic and magneto hydrodynamic) and empirical prediction models such as Narmax, Rice $Dst$ on 4 storm events which occurred between 2001 and 2006.
+\citet{Amata2008} compares the EDDA and the Lund $Dst$ models over the 2003-2005 period.
+But the most extensive model comparison in terms of storm events was conducted in \citet{Ji2012} which compared seven $Dst$ models (see table \ref{table:DstModels}) on a list of 63 geomagnetic storm events of varying intensities which occurred between 1998 and 2006.
 
 %Vision
-In $D_{st}$ prediction a seemingly trivial yet highly informative prediction method known as the \emph{Persistence} model has been largely ignored in its crucial role in model evaluation.
-The \emph{Persistence} model is stated simply as using the previous value of $D_{st}$ as the prediction for the next time step ($\hat{D_{st}}(t) = D_{st}(t-1)$).
-Due to high auto correlation between $D_{st}(t)$ and $D_{st}(t-1)$ in the Omni data, the \emph{Persistence} model gives excellent predictive performance despite its apparent lack of sophistication.
-In essence it is a trivial hypothesis and hence must always be used as a base line to compare the performance of any proposed $D_{st}$ algorithm.
+In $Dst$ prediction a seemingly trivial yet highly informative prediction method known as the \emph{Persistence} model has been largely ignored in its crucial role in model evaluation.
+The \emph{Persistence} model is stated simply as using the previous value of $Dst$ as the prediction for the next time step ($\hat{Dst}(t) = Dst(t-1)$).
+Due to high auto correlation between $Dst(t)$ and $Dst(t-1)$ in the Omni data, the \emph{Persistence} model gives excellent predictive performance despite its apparent lack of sophistication.
+In essence it is a trivial hypothesis and hence must always be used as a base line to compare the performance of any proposed $Dst$ algorithm.
 Therefore in the context of OSA prediction, a model must outperform the \emph{Persistence} model in order to be a viable candidate for advancing the science of space weather prediction.
@@
 %Assertion
-It is with the above observation in mind, we propose \emph{Gaussian Process} models for OSA prediction of $D_{st}$.
+It is with the above observation in mind, we propose \emph{Gaussian Process} models for OSA prediction of $Dst$.
 We use the results of \citet{Ji2012} as a starting point and compare our proposed models with the methods outlined in the same paper while using the performance of the \emph{Persistence} model as a litmus test.
@@
 Section \ref{sec:method} gives an introduction to the \emph{Gaussian Process} methodology for regression and the various issues surrounding their application.
-In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res}, bench marked against the \emph{Persistence} $D_{st}$ model as well as other models outlined in \citet{Ji2012}.
+In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res}, bench marked against the \emph{Persistence} $Dst$ model as well as other models outlined in \citet{Ji2012}.
 
@@
 The method of assigning values to the \emph{hyper-parameter} is crucial in the model building process (see section \ref{sec:hyp} below).
-In this study, we construct Gaussian Process regression models with polynomial kernels for \emph{one step ahead} prediction of the $D_{st}$ index.
+In this study, we construct Gaussian Process regression models with polynomial kernels for \emph{one step ahead} prediction of the $Dst$ index.
 
@@
 
-Below in equations \ref{eq:Dst}, \ref{eq:GPNoise}, \ref{eq:DstGP}, \ref{eq:GPPoly} we outline a \emph{Gaussian Process} formulation for \emph{OSA} prediction of $D_{st}$.
+Below in equations \ref{eq:Dst}, \ref{eq:GPNoise}, \ref{eq:DstGP}, \ref{eq:GPPoly} we outline a \emph{Gaussian Process} formulation for \emph{OSA} prediction of $Dst$.
 A vector of features $x_{t-1}$ is used as input to an unknown function $f(x_{t-1})$ which drawn from a \emph{Gaussian Process} driven by a linear polynomial kernel.
@@
 The features $x_{t-1}$ can be any collection of quantities in the hourly resolution Omni data set.
-Generally $x_{t-1}$ are time histories of $D_{st}$ or other important variables examples being plasma pressure $p(t)$, solar wind speed $V(t)$, z component of the interplanetary magnetic field $B_z(t)$.
+Generally $x_{t-1}$ are time histories of $Dst$ or other important variables examples being plasma pressure $p(t)$, solar wind speed $V(t)$, z component of the interplanetary magnetic field $B_z(t)$.
 
 \begin{eqnarray}
-    D_{st}(t) & = & f(\mathbf{x}_{t-1}) + \epsilon \label{eq:Dst} \\
+    Dst(t) & = & f(\mathbf{x}_{t-1}) + \epsilon \label{eq:Dst} \\
     \epsilon & \sim & \mathcal{N}(0, \sigma^2) \label{eq:GPNoise} \\
@@
 
-In \ref{sec:gpar} and \ref{sec:gparx} below we consider two choices for the input features $\mathbf{x}_{t-1}$ leading to two variants of \emph{Gaussian Process} regression for $D_{st}$ time series prediction.
+In \ref{sec:gpar} and \ref{sec:gparx} below we consider two choices for the input features $\mathbf{x}_{t-1}$ leading to two variants of \emph{Gaussian Process} regression for $Dst$ time series prediction.
 
@@
 
-The simplest auto-regressive models for \emph{OSA} prediction of $D_{st}$ are those that use only the history of $D_{st}$ to construct input features for model training.
-The input features $\mathbf{x}_{t-1}$ at each time step is the history of $D_{st}(t)$ until a time lag of $p$ steps.
+The simplest auto-regressive models for \emph{OSA} prediction of $Dst$ are those that use only the history of $Dst$ to construct input features for model training.
+The input features $\mathbf{x}_{t-1}$ at each time step is the history of $Dst(t)$ until a time lag of $p$ steps.
 
 \begin{eqnarray*}
-    \mathbf{x}_t & = & \left(D_{st}(t), D_{st}(t-1), \cdots , D_{st}(t-p+1)\right)
+    \mathbf{x}_t & = & \left(Dst(t), Dst(t-1), \cdots , Dst(t-p+1)\right)
 \end{eqnarray*}
@@
 
-Hourly solar wind velocity $V(t)$ and interplanetary magnetic field $B_z(t)$ recorded at the ACE satellite are used as the exogenous inputs of the model, along with the time history of $D_{st}$.
+Hourly solar wind velocity $V(t)$ and interplanetary magnetic field $B_z(t)$ recorded at the ACE satellite are used as the exogenous inputs of the model, along with the time history of $Dst$.
 In this model we choose distinct time lags $p$ and $p_{ex}$ for auto-regressive and exogenous variables respectively.
@@
 \begin{eqnarray*}
-        & \mathbf{x}_t & = (D_{st}(t), D_{st}(t-1), \cdots , D_{st}(t-p+1), \\
+        & \mathbf{x}_t & = (Dst(t), Dst(t-1), \cdots , Dst(t-p+1), \\
         & \ \ \ \ \ & V(t), V(t-1), \cdots, V(t-p_{ex}+1),\\
@@
 
-Before performance bench marks for \emph{OSA} $D_{st}$ prediction are run on the storm events in \cite{Ji2012}, training and model selection of \emph{GP-AR} and \emph{GP-ARX} models on independent data sets must be performed.
+Before performance bench marks for \emph{OSA} $Dst$ prediction are run on the storm events in \cite{Ji2012}, training and model selection of \emph{GP-AR} and \emph{GP-ARX} models on independent data sets must be performed.
 For this purpose we choose segments 00:00 1 January 2008 - 10:00 11 January 2008 for training and 00:00 15 November 2014 - 23:00 1 December 2014 for model selection.
@@
 
-\citet{Ji2012} compare a six OSA $D_{st}$ models in their paper by compiling a list of 63 geomagnetic storm events of varying strengths which occurred from 1998-2006.
+\citet{Ji2012} compare a six OSA $Dst$ models in their paper by compiling a list of 63 geomagnetic storm events of varying strengths which occurred from 1998-2006.
 This serves as an important stepping stone for a systematic comparison of existing and new prediction techniques because performance metrics are averaged over a large number of storm events that occurred over a 8 year period.
@@
     \item $RMSE$: The root mean square error
-    \item $CC$: Correlation coefficient between the predicted and actual value of $D_{st}$
-    \item $\Delta Dst_{min}$: The error in prediction of the peak value of $D_{st}$ for a storm event.
+    \item $CC$: Correlation coefficient between the predicted and actual value of $Dst$
+    \item $\Delta Dst_{min}$: The error in prediction of the peak value of $Dst$ for a storm event.
     \item $|\Delta t_{peak}|$: The error in predicting the timing of a storm peak, also called \emph{timing error}. 
@@
 
-Apart from the metrics above, we also generate the empirical distribution of absolute relative errors $|\frac{D_{st} - \hat{D_{st}}}{D_{st}}|$ for the \emph{GP-AR}, \emph{GP-ARX}, \emph{NM} and \emph{TL} models.
-These empirical cumulative distributions enable us to upper bound the percentage error in $D_{st}$ prediction and attach probability estimates for said bounds.
+Apart from the metrics above, we also generate the empirical distribution of absolute relative errors $|\frac{Dst - \hat{Dst}}{Dst}|$ for the \emph{GP-AR}, \emph{GP-ARX}, \emph{NM} and \emph{TL} models.
+These empirical cumulative distributions enable us to upper bound the percentage error in $Dst$ prediction and attach probability estimates for said bounds.
 
@@
 For the \emph{TL} model we use real time predictions listed on their website (\url{http://lasp.colorado.edu/home/spaceweather/}) corresponding to all the storm events.
-The real time predictions are at a frequency of ten minutes, hourly predictions are generated by averaging the $D_{st}$ predictions for each hour.
+The real time predictions are at a frequency of ten minutes, hourly predictions are generated by averaging the $Dst$ predictions for each hour.
 
 \begin{table}
-      \caption[]{One Step Ahead $D_{st}$ prediction models compared in \citet{Ji2012}}
+      \caption[]{One Step Ahead $Dst$ prediction models compared in \citet{Ji2012}}
          \label{table:DstModels}
@@
             NM & \citet{balikhin:narmax} & Non linear auto regressive with exogenous inputs. \\
-            B & \citet{JGR:JGR10260} & Prediction of $D_{st}$ by solving ODE having injection and decay terms. \\
+            B & \citet{JGR:JGR10260} & Prediction of $Dst$ by solving ODE having injection and decay terms. \\
             W & \citet{Wang:Dst} & Obtained by modification of injection term in \citet{JGR:JGR10260}. \\
@@
 Figure \ref{fig:relprob} compares the empirical distributions of absolute relative model errors across all storm events in the experiments.
-If one chooses a $95\%$ probability bound then it can be seen that the \emph{GP-ARX} model bounds the relative error to $20\%$ of the actual $D_{st}$ value, while in the case of the \emph{NM} model it bounds the relative error to $50\%$ of the actual $D_{st}$ value.
+If one chooses a $95\%$ probability bound then it can be seen that the \emph{GP-ARX} model bounds the relative error to $20\%$ of the actual $Dst$ value, while in the case of the \emph{NM} model it bounds the relative error to $50\%$ of the actual $Dst$ value.
 Therefore the \emph{GP-ARX} model gives an approximate improvement of $25\%$ over the other compared models on the test data set chosen.
@@
 
-Figure \ref{fig:cc} shows the comparison of correlation coefficients between model predictions and actual $D_{st}$ values.
+Figure \ref{fig:cc} shows the comparison of correlation coefficients between model predictions and actual $Dst$ values.
 From the results of \citet{Ji2012}, the \emph{TL} model gives the highest correlation coefficient of $94\%$, but that is not surprising considering the fact that there is a large overlap between the training data used to train it and storm events.
@@
 
-Figure \ref{fig:deltaDst} shows how the different models compare with respect to accuracy in predicting the peak value of $D_{st}$ during storm events ($\Delta Dst_{min}$ averaged over all storms).
+Figure \ref{fig:deltaDst} shows how the different models compare with respect to accuracy in predicting the peak value of $Dst$ during storm events ($\Delta Dst_{min}$ averaged over all storms).
 In the context of this metric the \emph{GP-ARX} model gives a $31\%$ improvement over the \emph{TL} model and $40\%$ over the \emph{NM} model.
-It is trivial to note that by definition, the \emph{Persistence} model ($\hat{D_{st}}(t) = D_{st}(t-1)$) will have $\Delta Dst_{min} = 0$ for every storm instance.
+It is trivial to note that by definition, the \emph{Persistence} model ($\hat{Dst}(t) = Dst(t-1)$) will have $\Delta Dst_{min} = 0$ for every storm instance.
 
@@
 \begin{enumerate}
-    \item Persistence Behavior: It is clear that the \emph{persistence} behavior in the $D_{st}$ values is very strong i.e.
-the trivial predictive model $\hat{D_{st}}(t) = D_{st}(t-1)$ gives excellent performance.
-In fact it can be seen in these figures that the \emph{Persistence} model is outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $D_{st}$.
-Therefore any model that is proposed to tackle the OSA prediction problem for $D_{st}$ must be compared to and show visible gains over the \emph{Persistence} model.
-    \item Assessment of Gaussian Process Models: The results indicate that there are many benefits to be gained by the application of Gaussian Process models for prediction of $D_{st}$.
+    \item Persistence Behavior: It is clear that the \emph{persistence} behavior in the $Dst$ values is very strong i.e.
+the trivial predictive model $\hat{Dst}(t) = Dst(t-1)$ gives excellent performance.
+In fact it can be seen in these figures that the \emph{Persistence} model is outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $Dst$.
+Therefore any model that is proposed to tackle the OSA prediction problem for $Dst$ must be compared to and show visible gains over the \emph{Persistence} model.
+    \item Assessment of Gaussian Process Models: The results indicate that there are many benefits to be gained by the application of Gaussian Process models for prediction of $Dst$.
 The \emph{GP-AR} and \emph{GP-ARX} not only outperform the models compared in \citet{Ji2012} but also the \emph{Persistence} model.
@@
    \begin{enumerate}
-      \item \emph{Persistence} model must be central in the model evaluation process in the context of \emph{One Step Ahead} prediction of the $D_{st}$ index.
-It is clear that the \emph{persistence} behavior in the $D_{st}$ values is very strong i.e.
-the trivial predictive model $\hat{D_{st}}(t) = D_{st}(t-1)$ gives excellent performance.
-In fact it can be seen in \ref{sec:res} that the \emph{Persistence} model is outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $D_{st}$.
-Therefore any model that is proposed to tackle the OSA prediction problem for $D_{st}$ must be compared to the \emph{Persistence} model and show visible gains above it.
+      \item \emph{Persistence} model must be central in the model evaluation process in the context of \emph{One Step Ahead} prediction of the $Dst$ index.
+It is clear that the \emph{persistence} behavior in the $Dst$ values is very strong i.e.
+the trivial predictive model $\hat{Dst}(t) = Dst(t-1)$ gives excellent performance.
+In fact it can be seen in \ref{sec:res} that the \emph{Persistence} model is outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $Dst$.
+Therefore any model that is proposed to tackle the OSA prediction problem for $Dst$ must be compared to the \emph{Persistence} model and show visible gains above it.
 

### 2016-05-18 swsc.tex
Updates from ShareLaTeX

@@
 
-The magnetosphere's dynamics and its associated solar wind driver form a complex nonlinear coupled dynamical system.
-It is therefore not only instructive but also greatly simplifying to use representative indices to quantify the state of geomagnetic activity.
-
-Geomagnetic indices come in various forms, they may take continuous or discrete values.
+The magnetosphere's dynamics and its associated solar wind driver form a complex dynamical system.
+It is therefore instructive and greatly simplifying to use representative indices to quantify the state of geomagnetic activity.
+
+Geomagnetic indices come in various forms, they may take continuous or discrete values, indices may be calculated at varying time frequencies.
 Their values are often calculated by averaging or combining a number of readings taken by instruments around the Earth.
@@
 
-\begin{comment}
 \begin{enumerate}
@@
 
-    \item AE: The Auroral Electrojet Index, AE, is designed to provide a global, quantitative measure of auroral zone magnetic activity produced by enhanced Ionospheric currents flowing below and within the auroral oval.
+    \item $AE$: The Auroral Electrojet Index, $AE$, is designed to provide a global, quantitative measure of auroral zone magnetic activity produced by enhanced Ionospheric currents flowing below and within the auroral oval.
+It is a continuous index which is calculated every hour.
 
@@
 \end{enumerate}
-\end{comment}
 
 %Talk about Burton and friends
-We focus on prediction of the hourly $Dst$ index which is an important problem because $Dst$ is a straightforward indicator of geomagnetic storms.
-More specifically we focus on the \emph{one step ahead} (OSA) prediction of $Dst$ because it is the first step towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions.
-
-The $Dst$ OSA prediction problem has been the subject of several modeling efforts so far, some of the earliest are by \citet{JGR:JGR10260} who outlined a prediction framework that has been used as a starting point for many $Dst$ models.
-\citet{JGR:JGR10260} model $Dst(t)$ a function of time using an \emph{Ordinary Differential Equation} (ODE) which expressed it's rate of change as a combination of two terms: decay and injection $\frac{d Dst(t)}{dt} = Q(t) - \frac{Dst(t)}{\tau}$.
-
-The \citet{JGR:JGR10260} model has proven to be very influential in the Space Weather research effort and many subsequent works like \citet{Wang:Dst}, \citet{JGRA:JGRA14856} modified the proposed ODE by proposing expressions for the injection term $Q(t)$.
+We focus on prediction of the hourly $Dst$ index which is a straightforward indicator of geomagnetic storms.
+More specifically, we focus on the \emph{one step ahead} (OSA) prediction of $Dst$ because it is the first piece towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions.
+
+The $Dst$ OSA prediction problem has been the subject of several modeling efforts in the literature.
+One the earliest models has been presented by \citet{JGR:JGR10260} who calculated $Dst(t)$ as the solution of an \emph{Ordinary Differential Equation} (ODE) which expressed the rate of change of $Dst(t)$ as a combination of two terms: decay and injection $\frac{d Dst(t)}{dt} = Q(t) - \frac{Dst(t)}{\tau}$.
+
+The \citet{JGR:JGR10260} model has proven to be very influential particularly due to its simplicity.
+Many subsequent works have modified the proposed ODE by proposing alternative expressions for the injection term $Q(t)$ [see \citet{Wang:Dst}, \citet{JGRA:JGRA14856}].
 More recently \citet{Ballatore2014} have tried to generate empirical estimates for the injection and decay terms in Burton's equation.
@@
 %Talk about NARMAX Dst
-Among the other major empirical models used to predict $Dst$ has been the \emph{Nonlinear Auto-Regessive Moving Average with eXogenous inputs} (NARMAX) methodology developed in \citet{doi:10.1080/00207178908559767}, \citet{GRL:GRL13494}, \citet{GRL:GRL20944}, \citet{JGRA:JGRA18657}, \citet{balikhin:narmax}, \citet{JGRA:JGRA20661} and \citet{JGRA:JGRA50192}.
-The NARMAX methodology builds models by constructing polynomial expansions of inputs and determines the best combinations or monomials to include in the refined model using the \emph{error reduction ratio} (ERR) criterion.
-The parameters of the NARMAX OLS-ERR model are calculated using the standard \emph{ordinary least squares} (OLS) method hence these methods are also denoted by the acronym NARMAX OLS-ERR.
+Another important empirical model used to predict $Dst$ is the \emph{Nonlinear Auto-Regessive Moving Average with eXogenous inputs} (NARMAX) methodology developed in \citet{doi:10.1080/00207178908559767}, \citet{GRL:GRL13494}, \citet{GRL:GRL20944}, \citet{JGRA:JGRA18657}, \citet{balikhin:narmax}, \citet{JGRA:JGRA20661} and \citet{JGRA:JGRA50192}.
+The NARMAX methodology builds models by constructing polynomial expansions of inputs and determines the best combinations of monomials to include in the refined model by using a criterion called the \emph{error reduction ratio} (ERR).
+The parameters of the so called NARMAX OLS-ERR model are calculated using the standard \emph{ordinary least squares} (OLS) method to minimize an objective function.
 The reader may refer to \citet{billings2013nonlinear} for a detailed exposition of the NARMAX methodology.
@@
 %Talk about neural networks
-\emph{Neural Networks} have been a popular choice for building predictive models and researchers employed the standard \emph{feed forward} and the more specialized \emph{recurrent} architectures.
-\citet{Lund} proposed an \emph{Elman} recurrent network architecture called Lund $Dst$, which used the solar wind, \emph{interplanetary magnetic field} (IMF) and historical $Dst$ data.
-\citet{SWE:SWE286} originally proposed a \emph{feed forward} network for predicting the $K_p$ index which used the \emph{Boyle coupling function} (\citet{boyle1997empirical}), in \citet{SWE:SWE286} the same architecture is adapted for prediction of $Dst$, which is popularly known as the Rice $Dst$ model.
+Yet another family of forecasting methods is based on \emph{Artificial Neural Networks} (ANN) that have been a popular choice for building predictive models.
+Researchers have employed both the standard \emph{feed forward} and the more specialized \emph{recurrent} architectures.
+\citet{Lund} proposed an \emph{Elman} recurrent network architecture called Lund $Dst$, which used the solar wind velocity, \emph{interplanetary magnetic field} (IMF) and historical $Dst$ data as inputs.
+\citet{SWE:SWE286} originally proposed a \emph{feed forward} network for predicting the $K_p$ index which used the \emph{Boyle coupling function} [see \citet{boyle1997empirical}].
+The same architecture is adapted for prediction of $Dst$ in \citet{SWE:SWE286}, which is popularly known as the Rice $Dst$ model.
 \citet{pallocchia:hal-00318011} proposed a \emph{neural network} model called EDDA to predict $Dst$ using only the IMF data.
@@
 %Talk about comparison papers, number of storms covered and Persistence
-In light of the extensive list of modeling techniques employed for prediction of geomagnetic indices and in particular $Dst$, model comparison and evaluation becomes a crucial direction in advancing the research domain.
-\citet{Rastatter2013} compare several physics based (convection, kinetic and magneto hydrodynamic) and empirical prediction models such as Narmax, Rice $Dst$ on 4 storm events which occurred between 2001 and 2006.
-\citet{Amata2008} compares the EDDA and the Lund $Dst$ models over the 2003-2005 period.
-But the most extensive model comparison in terms of storm events was conducted in \citet{Ji2012} which compared seven $Dst$ models (see table \ref{table:DstModels}) on a list of 63 geomagnetic storm events of varying intensities which occurred between 1998 and 2006.
+In light of the extensive list of modeling techniques employed for prediction of the $Dst$ index, model comparison and evaluation becomes a crucial direction in advancing the research domain.
+\citet{Rastatter2013} compare several physics based (convection, kinetic and magneto hydrodynamic) and empirical prediction models such as NARMAX, Rice $Dst$ on 4 storm events which occurred between 2001 and 2006.
+\citet{Amata2008} compare the EDDA and the Lund $Dst$ models over the 2003-2005 period.
+The most extensive model comparison in terms of storm events was probably conducted in \citet{Ji2012} which compared six $Dst$ models (see table \ref{table:DstModels}) on a list of 63 geomagnetic storm events of varying intensities which occurred between 1998 and 2006.
+In the comparison done in \citet{Ji2012}, the model proposed in \citet{JGRA:JGRA16300} gives the best performance on the test set considered.
 

### 2016-05-19 swsc.tex
Updates from ShareLaTeX

@@
 %Vision
-In $Dst$ prediction a seemingly trivial yet highly informative prediction method known as the \emph{Persistence} model has been largely ignored in its crucial role in model evaluation.
-The \emph{Persistence} model is stated simply as using the previous value of $Dst$ as the prediction for the next time step ($\hat{Dst}(t) = Dst(t-1)$).
-Due to high auto correlation between $Dst(t)$ and $Dst(t-1)$ in the Omni data, the \emph{Persistence} model gives excellent predictive performance despite its apparent lack of sophistication.
-In essence it is a trivial hypothesis and hence must always be used as a base line to compare the performance of any proposed $Dst$ algorithm.
-Therefore in the context of OSA prediction, a model must outperform the \emph{Persistence} model in order to be a viable candidate for advancing the science of space weather prediction.
+In $Dst$ prediction a seemingly trivial yet highly informative prediction method is represented by the so called \emph{Persistence} model this is simply using the previous value of $Dst$ as the prediction for the next time step ($\hat{D}st(t) = Dst(t-1)$).
+Due to high correlation between $Dst$ values one hour apart ($Dst(t)$, $Dst(t-1)$), despite its apparent lack of sophistication the \emph{Persistence} model gives excellent predictive performance on error metrics.
+In essence it is a trivial hypothesis and hence it should always be used as a base line to compare the performance of any proposed $Dst$ algorithm.
+Moreover, given its zero computational cost, a model should at least outperform the \emph{Persistence} model in order to be a viable candidate for advancing the science of space weather prediction.
 
 %Assertion
-It is with the above observation in mind, we propose \emph{Gaussian Process} models for OSA prediction of $Dst$.
-We use the results of \citet{Ji2012} as a starting point and compare our proposed models with the methods outlined in the same paper while using the performance of the \emph{Persistence} model as a litmus test.
+In this paper, we propose \emph{Gaussian Process} models for OSA prediction of $Dst$.
+We use the results of \citet{Ji2012} as a starting point and compare our proposed models with the methods outlined in the same paper while using the performance of the \emph{Persistence} model as a base line.
 
 % Talk about Gaussian Process
-We use the hourly resolution Omni data set (see \cite{OmniPaper}) as extracted from NASA/GSFC's OMNI data set through OMNIWeb.
+We use the hourly resolution data as extracted from NASA/GSFC's OMNI data set through OMNIWeb (see \cite{OmniPaper}).
 The rest of the paper is organized as follows.
-Section \ref{sec:method} gives an introduction to the \emph{Gaussian Process} methodology for regression and the various issues surrounding their application.
-In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res}, bench marked against the \emph{Persistence} $Dst$ model as well as other models outlined in \citet{Ji2012}.
+Section \ref{sec:method} gives an introduction to the \emph{Gaussian Process} methodology for regression and some technical details on their application.
+In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res}, bench marked against the \emph{Persistence} $Dst$ model as well as the other models outlined in \citet{Ji2012}.
 

### 2016-05-22 swsc.tex
Updates from ShareLaTeX

@@
 
-A \emph{Gaussian Process} is a stochastic process $f(\mathbf{x}), \ \mathbf{x} \in \mathbb{R}^d$ whose finite dimensional distributions over a set of points $\mathbf{x}_1, \cdots, \mathbf{x}_n$ are multivariate Gaussian.
-Formally this is expressed as.
-
-\begin{eqnarray}
- \mathbf{f} & = & \left( \begin{array}{c} f(\mathbf{x}_1) \\ f(\mathbf{x}_2) \\ \vdots \\ f(\mathbf{x}_n) \end{array} \right) \\
- \vspace{2\baselineskip}
- \mathbf{f} | \mathbf{x}_1, \cdots, \mathbf{x}_n & \sim & \mathcal{N}\left( \mathbf{\mu}, \mathbf{\Lambda} \right)  \\
- \vspace{2\baselineskip}
- p( \mathbf{f} \ | \ \mathbf{x}_1, \cdots, \mathbf{x}_n) & = & \frac{1}{(2\pi)^{n/2} det(\mathbf{\Lambda})^{1/2}} exp \left(-\frac{1}{2} (\mathbf{f} - \mathbf{\mu})^T \mathbf{\Lambda}^{-1} (\mathbf{f} - \mathbf{\mu}) \right) \label{eq:sto}
-\end{eqnarray}
-
-From equation \ref{eq:sto}, one can observe that in order to uniquely define the distribution of the process, it is required to specify the mean vector $\mathbf{\mu}$ and the covariance matrix $\mathbf{\Lambda}$.
-For this probability density to be valid, there is a further requirements imposed on $\mathbf{\Lambda}$:
-
-\begin{enumerate}
-      \item Symmetry: $\mathbf{\Lambda}_{ij} = \mathbf{\Lambda}_{ij} \ \forall i,j \in {1, \cdots, n} $ 
-      \item Positive Semi-definiteness: $\mathbf{y}^T \mathbf{\Lambda} \mathbf{y} \geq 0 \ \forall \mathbf{y} \in \mathbb{R}^n$  
-\end{enumerate}
-
-Inspecting the individual elements of $\mathbf{\mu}$ and $\mathbf{\Lambda}$, we realise that they take the following form.
-
-\begin{eqnarray}
-      \mu_i & = & \mathbb{E}[f(\mathbf{x}_i)] := m(\mathbf{x}_i) \\
-      \Lambda_{ij} & = & \mathbb{E}[(f(\mathbf{x}_i) - \mu_i)(f(\mathbf{x}_j) - \mu_j)] := K(\mathbf{x}_i, \mathbf{x}_j)
-\end{eqnarray}
-
-Here $\mathbb{E}$ denotes the expectation (average).
-The elements of $\mathbf{\mu}$ and $\mathbf{\Lambda}$ are expressed as functions $m(\mathbf{x}_i)$ and $K(\mathbf{x}_i, \mathbf{x}_j)$ of the inputs $\mathbf{x}_i,\ \mathbf{x}_j$.
-Specifying the functions $m(\mathbf{x})$ and $K(\mathbf{x}, \mathbf{x}')$ completely specifies each element of $\mathbf{\mu}$ and $\mathbf{\Lambda}$ and subsequently the finite dimensional distribution of $\mathbf{f} | \mathbf{x}_1, \cdots, \mathbf{x}_n $.
-In machine learning literature \emph{Gaussian Processes} are represented using the following notation.
-
-\begin{equation}
-    f(\mathbf{x}) \sim \mathcal{GP}(m(\mathbf{x}), K(\mathbf{x}, \mathbf{x}'))
-\end{equation}
-
-\subsection{Gaussian Process Regression: Formulation}
+\emph{Gaussian Processes} first appeared in machine learning research in \citet{Neal:1996:BLN:525544}, as the limiting case of Bayesian inference performed on neural networks with infinitely many neurons in the hidden layers.
+Although their inception in the machine learning community is recent, their origins can be traced back to the geo-statistics research community where they are known as \emph{Kriging} methods (\citet{krige1951statistical}).
+In pure mathematics area \emph{Gaussian Processes} have been studied extensively and their existence was first proven by Kolmogorov's extension theorem (\citet{tao2011introduction}).
+The reader is referred to \cite{Rasmussen:2005:GPM:1162254} for an in depth treatment of Gaussian Processes in machine learning.
+
+Let us assume that we want to model a process in which a scalar quantity $y$ is described by $y = f(\mathbf{x}) + \epsilon$ where   $f(.): \mathbb{R}^d \rightarrow \mathbb{R}$ is an unknown scalar function of a multidimensional input vector $\mathbf{x} \in \mathbb{R}^d$, and $\epsilon \sim \mathcal{N}(0, \sigma^2)$ is Gaussian distributed noise with variance $\sigma^2$.
+
+On the basis of a set of labeled data points ${(\mathbf{x}_i, y_i); i = 1 \cdots N}$ which can be conveniently expressed by $\mathbf{X}$ a $N \times d$ data matrix and $\mathbf{y}$ a $N \times 1$ response vector as shown in equations \ref{eq:feat} and \ref{eq:labels}.
 
@@
 
-We start with a set of labeled data points $ {(\mathbf{x}_i, y_i): \forall i \in 1 \cdots N, \ \mathbf{x}_i \in \mathbb{R}^d \ y_i \in \mathbb{R}} $ condensed into a data matrix $\mathbf{X}$ and response vector $\mathbf{y}$ in equations \ref{eq:feat} and \ref{eq:labels}.
-The values $y_i$ are assumed to be measurements of an unknown function $f(\mathbf{x}_i)$ corrupted by an independent and identically distributed zero mean Gaussian noise.
-The \emph{Gaussian Process} regression framework is the defined by the following components.
+Out task is to infer the the values of the unknown function $f(.)$ based on the inputs $\mathbf{X}$ and the noisy observations $\mathbf{y}$.
+We now assume that the joint distribution of $f(\mathbf{x}_i), i = 1 \cdots N$ is multivariate Gaussian as shown in equations \ref{eq:fvalues}, \ref{eq:normal} and \ref{eq:sto}.
 
 \begin{eqnarray}
-      y & = & f(\mathbf{x}) + \epsilon  \\
-      \epsilon & \sim & \mathcal{N}(0, \sigma^2) \\
-      f(\mathbf{x}) & \sim & \mathcal{GP}(m(\mathbf{x}), K(\mathbf{x},\mathbf{x}')) 
+ \mathbf{f} & = & \left( \begin{array}{c} f(\mathbf{x}_1) \\ f(\mathbf{x}_2) \\ \vdots \\ f(\mathbf{x}_N) \end{array} \right) \label{eq:fvalues}\\
+ \vspace{2\baselineskip}
+ \mathbf{f} | \mathbf{x}_1, \cdots, \mathbf{x}_N & \sim & \mathcal{N}\left( \mathbf{\mu}, \mathbf{\Lambda} \right)  \label{eq:normal}\\
+ \vspace{2\baselineskip}
+ p( \mathbf{f} \ | \ \mathbf{x}_1, \cdots, \mathbf{x}_N) & = & \frac{1}{(2\pi)^{n/2} det(\mathbf{\Lambda})^{1/2}} exp \left(-\frac{1}{2} (\mathbf{f} - \mathbf{\mu})^T \mathbf{\Lambda}^{-1} (\mathbf{f} - \mathbf{\mu}) \right) \label{eq:sto}
 \end{eqnarray}
 
+Here $\matbf{f}$ is an $N\times 1$ vector consisting of the values $f(\mathbf{x}_i), i = q \cdots N$.
+In equation \ref{eq:normal} $\mathbf{f}|\mathbf{x}_1, \cdots, \mathbf{x}_N$ denotes that we are modeling the conditional distribution of $\mathbf{f}$ with respect to the input data (i.e.
+$\mathbf{X}$).
+$\mathcal{N}\left( \mathbf{\mu}, \mathbf{\Lambda} \right)$ represents a multivariate Gaussian distribution with mean vector $\mathbf{\mu}$ and covariance matrix $\mathbf{\Lambda}$.
+
+From equation \ref{eq:sto}, one can observe that in order to uniquely define the distribution of the process, it is required to specify $\mathbf{\mu}$ and $\mathbf{\Lambda}$.
+For this probability density to be valid, there is a further requirements imposed on $\mathbf{\Lambda}$:
+
+\begin{enumerate}
+      \item Symmetry: $\mathbf{\Lambda}_{ij} = \mathbf{\Lambda}_{ij} \ \forall i,j \in {1, \cdots, N} $ 
+      \item Positive Semi-definiteness: $\mathbf{z}^T \mathbf{\Lambda} \mathbf{z} \geq 0 \ \forall \mathbf{z} \in \mathbb{R}^N$  
+\end{enumerate}
+
+Inspecting the individual elements of $\mathbf{\mu}$ and $\mathbf{\Lambda}$, we realise that they take the following form.
+
+\begin{eqnarray}
+      \mu_i & = & \mathbb{E}[f(\mathbf{x}_i)] := m(\mathbf{x}_i) \\
+      \Lambda_{ij} & = & \mathbb{E}[(f(\mathbf{x}_i) - \mu_i)(f(\mathbf{x}_j) - \mu_j)] := K(\mathbf{x}_i, \mathbf{x}_j)
+\end{eqnarray}
+
+Here $\mathbb{E}$ denotes the expectation (average).
+The elements of $\mathbf{\mu}$ and $\mathbf{\Lambda}$ are expressed as functions $m(\mathbf{x}_i)$ and $K(\mathbf{x}_i, \mathbf{x}_j)$ of the inputs $\mathbf{x}_i,\ \mathbf{x}_j$.
+Specifying the functions $m(\mathbf{x})$ and $K(\mathbf{x}, \mathbf{x}')$ completely specifies each element of $\mathbf{\mu}$ and $\mathbf{\Lambda}$ and subsequently the finite dimensional distribution of $\mathbf{f} | \mathbf{x}_1, \cdots, \mathbf{x}_N $.
+In most practical applications of \emph{Gaussian Processes} the mean function is often defined as $m(\mathbf{x}) = 0$, which is not unreasonable if the data is standardized to have zero mean.
+\emph{Gaussian Processes} are represented in machine learning literature using the following notation.
+
+\begin{equation}
+    f(\mathbf{x}) \sim \mathcal{GP}(m(\mathbf{x}), K(\mathbf{x}, \mathbf{x}'))
+\end{equation}
+
 \subsection{Inference and Predictions} \label{sec:inference}
 
-Our aim is to generate infer the function $f(\mathbf{x})$ from the noisy training data and generate predictions $f(\mathbf{x}^{*}_i)$ for a set of test points $ {\mathbf{x}^{*}_i : \forall i \in 1, \cdots, m} $.
-We start from equation \ref{eq:sto} and use it to construct the distribution of $f(\mathbf{x})$ over the training and test points, how ever since we have noisy measurements of $f$ over the training data, we add the noise variance $\sigma^2$ to the variance of $f$ as shown in \ref{eq:dist}.
+Our aim is to infer the function $f(\mathbf{x})$ from the noisy training data and generate predictions $f(\mathbf{x}^{*}_i)$ for a set of test points $ {\mathbf{x}^{*}_i : \forall i \in 1, \cdots, M} $.
+We define $\mathbf{X}^*$ as the test data matrix whose rows are formed by $\mathbf{x}^{*}_i$ as shown in equation \ref{eq:testfeat}.
+Using the multivariate Gaussian distribution in equation \ref{eq:sto} we can construct the distribution of $f(\mathbf{x})$ over the training and test points.
+The vector of training and test outputs $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right)$ is of dimension $(N+M) \times 1$ and is constructed by appending the test set predictions $\mathbf{f}_* = \left( \begin{array}{c} f(\mathbf{x^{*}}_1) \\ f(\mathbf{x^{*}}_2) \\ \vdots \\ f(\mathbf{x^{*}}_M) \end{array} \right)$ to the observed noisy measurements $\mathbf{y}$.
 
 \begin{eqnarray}
-    \mathbf{X}_* & = & \left( \begin{array}{c} (\mathbf{x}^{*}_1)^T \\ (\mathbf{x}^{*}_2)^T \\ \vdots \\ (\mathbf{x}^{*}_m)^T \end{array} \right)_{m \times d} \\
-    \mathbf{f}_* & = & \left( \begin{array}{c} f(\mathbf{x^{*}_1}) \\ f(\mathbf{x^{*}_2}) \\ \vdots \\ f(\mathbf{x^{*}_m}) \end{array} \right)_{m \times 1} \\
+    \mathbf{X}_* & = & \left( \begin{array}{c} (\mathbf{x}^{*}_1)^T \\ (\mathbf{x}^{*}_2)^T \\ \vdots \\ (\mathbf{x}^{*}_M)^T \end{array} \right)_{M \times d} \label{eq:testfeat} \\
+    \mathbf{f}_* & = & \left( \begin{array}{c} f(\mathbf{x^{*}_1}) \\ f(\mathbf{x^{*}_2}) \\ \vdots \\ f(\mathbf{x^{*}_M}) \end{array} \right)_{M \times 1} \\
      \vspace{4\baselineskip}
@@
 
-In equation \ref{eq:dist} the block matrix components of the $n+m \times n+m$ covariance matrix have the following structure.
+Since we have noisy measurements of $f$ over the training data, we add the noise variance $\sigma^2$ to the variance of $f$ as shown in \ref{eq:dist}.
+The block matrix components of the $(N+M) \times (N+M)$ covariance matrix have the following structure.
 
@@
 
-With the multivariate normal distribution defined in equation \ref{eq:dist}, probabilistic predictions $f_*$ can be generated by constructing the conditional distribution $\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}$, since the original distribution of $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right) | \ \ \mathbf{X}, \mathbf{X}_*$ is a multivariate Gaussian, conditioning on a subset of elements $\mathbf{y}$ yields another Gaussian distribution with mean and covariance given below \ref{eq:posterior}.
+With the multivariate normal distribution defined in equation \ref{eq:dist}, probabilistic predictions $f_*$ can be generated by constructing the conditional distribution $\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}$.
+Since the original distribution of $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right) | \ \ \mathbf{X}, \mathbf{X}_*$ is a multivariate Gaussian, conditioning on a subset of elements $\mathbf{y}$ yields another Gaussian distribution whose mean and covariance can be calculated exactly as equation \ref{eq:posterior} (see \citet{Rasmussen:2005:GPM:1162254}).
 
@@
 
-\emph{Gaussian Processes} first appeared in machine learning research in \citet{Neal:1996:BLN:525544}, as the limiting case of Bayesian inference performed on neural networks with infinitely many neurons in the hidden layers.
-Although their inception in the machine learning community is recent, their origins can be traced back to the geo-statistics research community where they are known as \emph{Kriging} methods (\citet{krige1951statistical}).
-In pure mathematics area \emph{Gaussian Processes} have been studied extensively and their existence was first proven by Kolmogorov's extension theorem (\citet{tao2011introduction}).
-The reader is requested to refer to \cite{Rasmussen:2005:GPM:1162254} for an in depth treatment of Gaussian Processes in machine learning.
-
-In most practical applications of \emph{Gaussian Processes} the mean function is often defined as $m(\mathbf{x}) = 0$, which is not unreasonable if the data is standardized to have zero mean.
 The central design issue now becomes choice of the function $K(\mathbf{x}, \mathbf{x}')$.
@@
 The symmetry and positive semi-definiteness of \emph{Gaussian Process} kernels implies that they represent inner-products between some basis function representation of the data.
-This is analogous to a Fourier series decomposition in Physics.
 The interested reader is suggested to refer to \cite{Berlinet2004}, \citet{Scholkopf:2001:LKS:559923} and \citet{hofmann2008} for a thorough treatment on kernel functions and the rich theory behind them.
@@
 The quantities $l$ in the RBF, $b$ and $d$ in the polynomial kernel are known as \emph{hyper-parameters}.
-Hyper-parameters give flexibility to particular kernel structure, for example $d = 1, 2, 3, \cdots$ in the polynomial kernel represents linear, quadratic, cubic and higher order phenomena respectively.
+Hyper-parameters give flexibility to particular kernel structure, for example $d = 1, 2, 3, \cdots$ in the polynomial kernel represents linear, quadratic, cubic and higher order polynomials respectively.
 The method of assigning values to the \emph{hyper-parameter} is crucial in the model building process (see section \ref{sec:hyp} below).
@@
 
-When fitting \emph{Gaussian Process} models to data, the question to be answered following choice of a kernel function is the assignment of appropriate values to the kernel hyper-parameters and noise level ($\sigma^2$).
+When applying \emph{Gaussian Process} models to data, one must choose appropriate values to the kernel hyper-parameters and noise level ($\sigma^2$).
 This is known as the problem of \emph{model selection}.

### 2016-05-23 swsc.tex
Updates from ShareLaTeX

@@
 Here $\matbf{f}$ is an $N\times 1$ vector consisting of the values $f(\mathbf{x}_i), i = q \cdots N$.
-In equation \ref{eq:normal} $\mathbf{f}|\mathbf{x}_1, \cdots, \mathbf{x}_N$ denotes that we are modeling the conditional distribution of $\mathbf{f}$ with respect to the input data (i.e.
-$\mathbf{X}$).
-$\mathcal{N}\left( \mathbf{\mu}, \mathbf{\Lambda} \right)$ represents a multivariate Gaussian distribution with mean vector $\mathbf{\mu}$ and covariance matrix $\mathbf{\Lambda}$.
+In equation \ref{eq:normal} $\mathbf{f}|\mathbf{x}_1, \cdots, \mathbf{x}_N$ denotes the conditional distribution of $\mathbf{f}$ with respect to the input data (i.e.
+$\mathbf{X}$) and $\mathcal{N}\left( \mathbf{\mu}, \mathbf{\Lambda} \right)$ represents a multivariate Gaussian distribution with mean vector $\mathbf{\mu}$ and covariance matrix $\mathbf{\Lambda}$.
+The probability density function of this distribution $p( \mathbf{f} \ | \ \mathbf{x}_1, \cdots, \mathbf{x}_N)$ is therefore given by equation \ref{eq:sto}.
 
@@
 With the multivariate normal distribution defined in equation \ref{eq:dist}, probabilistic predictions $f_*$ can be generated by constructing the conditional distribution $\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}$.
-Since the original distribution of $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right) | \ \ \mathbf{X}, \mathbf{X}_*$ is a multivariate Gaussian, conditioning on a subset of elements $\mathbf{y}$ yields another Gaussian distribution whose mean and covariance can be calculated exactly as equation \ref{eq:posterior} (see \citet{Rasmussen:2005:GPM:1162254}).
+Since the original distribution of $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right) | \ \ \mathbf{X}, \mathbf{X}_*$ is a multivariate Gaussian, conditioning on a subset of elements $\mathbf{y}$ yields another Gaussian distribution whose mean and covariance can be calculated exactly in equation \ref{eq:posterior} (see \citet{Rasmussen:2005:GPM:1162254}).
 
@@
 
+The practical implementation of \emph{Gaussian Process} models requires the inversion of the training data kernel matrix $[\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1}$ to calculate the parameters of the predictive distribution $\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}$.
+The computational complexity of this inference is dominated by the linear problem $\mathbf{K}^T_{*} [\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y}$, this can be done via Cholesky decomposition which has time complexity $O(N^3)$.
+
 The distribution of $\mathbf{f_*}| \mathbf{X},\mathbf{y},\mathbf{X_*}$ is known in Bayesian analysis as the \emph{Posterior Predictive Distribution}.
@@
 
-The central design issue now becomes choice of the function $K(\mathbf{x}, \mathbf{x}')$.
+The central design issue in applying \emph{Gaussian Process} models is choice of the function $K(\mathbf{x}, \mathbf{x}')$.
 The same constraints that apply to $\mathbf{\Lambda}$ also apply to the function $K(.,.)$.
@@
 The method of assigning values to the \emph{hyper-parameter} is crucial in the model building process (see section \ref{sec:hyp} below).
-In this study, we construct Gaussian Process regression models with polynomial kernels for \emph{one step ahead} prediction of the $Dst$ index.
-
-\begin{comment}
-\subsubsection{Radial Basis Function (RBF)}
-
-The RBF kernel is the most commonly used kernel in machine learning, in its core it encodes the idea of \emph{smoothness} because functions drawn from its Hilbert space are infinitely differentiable.
-The RBF kernel is also an example of a \emph{stationary} kernel i.e.
-its value depends only on $||\mathbf{x} - \mathbf{y}||$.
-The quantity $l$ is known as the characteristic \emph{length scale} of this kernel, it is a free parameter (also known as hyper-parameter) which must be determined by the modeller, refer to section \ref{sec:hyp} on how suitable values of kernel hyper-parameters.
-
-\begin{equation}
-    K(\mathbf{x}, \mathbf{y}) = \frac{1}{2} exp(-||\mathbf{x} - \mathbf{y}||^2/l^2)
-\end{equation}
-
-\subsubsection{Polynomial}
-
-The polynomial kernel is also a common kernel function employed in research, it has two hyper-parameters the degree $d$ and an intercept $b$.
-The polynomial kernel formulates the function approximation in terms of polynomial expansions of the input dimensions $\mathbf{x}$, more specifically it considers all monomial combination of the input dimensions up to degree $d$.
-When $d = 1$ it corresponds to a Bayesian linear regression model.
-
-\begin{equation}
-    K(\mathbf{x}, \mathbf{y}) = (\mathbf{x}^\intercal \mathbf{y} + b)^d 
-\end{equation}
-
-\end{comment}
+In this study, we construct Gaussian Process regression models with linear polynomial kernels for \emph{one step ahead} prediction of the $Dst$ index.
 
@@
 
-The simplest way to perform \emph{model selection} is called \emph{grid search} which entails evaluating the above criteria on a fixed grid which is constructed by making a list of values for each hyper-parameter and taking the combinations of all such values.
+The simplest way to perform \emph{model selection} is called \emph{grid search} which entails evaluating the above criteria on a fixed grid which is constructed by the tensorial product of grids constructed on each hyper-parameter.
 There are more sophisticated methods for \emph{model tuning} such as \emph{coupled simulated annealing} (\citet{Xavier-De-Souza2010}) and \emph{simplex search} (\citet{Nelder1965}) but they are computationally more expensive.
@@
 The simplest auto-regressive models for \emph{OSA} prediction of $Dst$ are those that use only the history of $Dst$ to construct input features for model training.
-The input features $\mathbf{x}_{t-1}$ at each time step is the history of $Dst(t)$ until a time lag of $p$ steps.
+The input features $\mathbf{x}_{t-1}$ at each time step are the history of $Dst(t)$ until a time lag of $p$ hours.
 
@@
 Auto-regressive models can be improved by including exogenous quantities that can help in improving predictions.
-It is well known that the solar wind and interplanetary magnetic field are important factors that influence the Earth's geomagnetic state.
-
-Hourly solar wind velocity $V(t)$ and interplanetary magnetic field $B_z(t)$ recorded at the ACE satellite are used as the exogenous inputs of the model, along with the time history of $Dst$.
+It is well known that the solar wind and interplanetary magnetic field are important factors that influence the Earth's geomagnetic state (\citet{JGRA:JGRA11775}, \citet{Russell1974}).
+
+Hourly solar wind velocity $V(t)$ and interplanetary magnetic field $B_z(t)$ recorded by the ACE satellite are used as the exogenous inputs of the model, along with the time history of $Dst$.
 In this model we choose distinct time lags $p$ and $p_{ex}$ for auto-regressive and exogenous variables respectively.
+In addition we explicitly include the product $V B_z$ which is a proxy for the electric field, as an input.
 
@@
 
-Before performance bench marks for \emph{OSA} $Dst$ prediction are run on the storm events in \cite{Ji2012}, training and model selection of \emph{GP-AR} and \emph{GP-ARX} models on independent data sets must be performed.
+Before running performance bench marks for \emph{OSA} $Dst$ prediction on the storm events in \cite{Ji2012}, training and model selection of \emph{GP-AR} and \emph{GP-ARX} models on independent data sets must be performed.
 For this purpose we choose segments 00:00 1 January 2008 - 10:00 11 January 2008 for training and 00:00 15 November 2014 - 23:00 1 December 2014 for model selection.
@@
    \includegraphics[width=\textwidth]{Compare_RMSE.png}
-      \caption{Averaged RMSE performance of GP-AR, GP-ARX and the Persistence model versus results in \citet{Ji2012}.}
+      \caption{Averaged RMSE performance of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model versus results in \citet{Ji2012}.}
          \label{fig:rmse}
@@
 
-\citet{Ji2012} compare a six OSA $Dst$ models in their paper by compiling a list of 63 geomagnetic storm events of varying strengths which occurred from 1998-2006.
+\citet{Ji2012} compare six OSA $Dst$ models in their paper by compiling a list of 63 geomagnetic storm events of varying strengths which occurred from 1998-2006.
 This serves as an important stepping stone for a systematic comparison of existing and new prediction techniques because performance metrics are averaged over a large number of storm events that occurred over a 8 year period.
@@
 \begin{enumerate}
-    \item $RMSE$: The root mean square error
-    \item $CC$: Correlation coefficient between the predicted and actual value of $Dst$
-    \item $\Delta Dst_{min}$: The error in prediction of the peak value of $Dst$ for a storm event.
-    \item $|\Delta t_{peak}|$: The error in predicting the timing of a storm peak, also called \emph{timing error}. 
+    \item The root mean square error.
+    \begin{equation}
+        RMSE = \sqrt{\sum_{i=1}^{n} (y - \hat{y})^2 / n}
+    \end{equation}
+    \item Correlation coefficient between the predicted and actual value of $Dst$.
+    \begin{equation}
+        CC = Cov(y \hat{y})/\sqrt{Var(y) Var(\hat{y})}
+    \end{equation}
+    \item The error in prediction of the peak negative value of $Dst$ for a storm event.
+    \begin{equation}
+        \Delta Dst_{min} = Dst(t^*) - \hat{D}st(t^*), \ t^* = argmin(Dst(t))
+    \end{equation}
+    \item The error in predicting the timing of a storm peak, also called \emph{timing error} $|\Delta t_{peak}|$. 
 \end{enumerate}
@@
    \includegraphics[width=\textwidth]{Compare_RelProb.png}
-   \caption{Comparison of empirical distribution of relative errors for GP-AR, GP-ARX, NM and TL on the test set of \citet{Ji2012}.}
+   \caption{Comparison of empirical distribution of relative errors for \emph{GP-AR}, \emph{GP-ARX}, \emph{NM} and \emph{TL} on the test set of \citet{Ji2012}.}
    \label{fig:relprob}
@@
 
-Apart from the metrics above, we also generate the empirical distribution of absolute relative errors $|\frac{Dst - \hat{Dst}}{Dst}|$ for the \emph{GP-AR}, \emph{GP-ARX}, \emph{NM} and \emph{TL} models.
-These empirical cumulative distributions enable us to upper bound the percentage error in $Dst$ prediction and attach probability estimates for said bounds.
-
-\subsection{Models Compared}
+Apart from the metrics above, we also generate the empirical distribution of relative errors $|\frac{Dst - \hat{Dst}}{Dst}|$ for the \emph{GP-AR}, \emph{GP-ARX}, \emph{NM} and \emph{TL} models.
+
+\subsection{Model Comparison}
 

### 2016-05-23 swsc.tex
Updates from ShareLaTeX

@@
 Using the multivariate Gaussian distribution in equation \ref{eq:sto} we can construct the distribution of $f(\mathbf{x})$ over the training and test points.
-The vector of training and test outputs $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right)$ is of dimension $(N+M) \times 1$ and is constructed by appending the test set predictions $\mathbf{f}_* = \left( \begin{array}{c} f(\mathbf{x^{*}}_1) \\ f(\mathbf{x^{*}}_2) \\ \vdots \\ f(\mathbf{x^{*}}_M) \end{array} \right)$ to the observed noisy measurements $\mathbf{y}$.
+The vector of training and test outputs $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right)$ is of dimension $(N+M) \times 1$ and is constructed by appending the test set predictions $\mathbf{f}_*$ to the observed noisy measurements $\mathbf{y}$.
 
@@
 
-\citet{Ji2012} compare six OSA $Dst$ models in their paper by compiling a list of 63 geomagnetic storm events of varying strengths which occurred from 1998-2006.
+\citet{Ji2012} compare six \emph{one step ahead} $Dst$ models as listed in table \ref{table:DstModels} in their paper by compiling a list of 63 geomagnetic storm events of varying strengths which occurred from 1998-2006.
 This serves as an important stepping stone for a systematic comparison of existing and new prediction techniques because performance metrics are averaged over a large number of storm events that occurred over a 8 year period.
@@
 
-\begin{figure}
-   \centering
-   \includegraphics[width=\textwidth]{Compare_RelProb.png}
-   \caption{Comparison of empirical distribution of relative errors for \emph{GP-AR}, \emph{GP-ARX}, \emph{NM} and \emph{TL} on the test set of \citet{Ji2012}.}
-   \label{fig:relprob}
-\end{figure}
-
 Apart from the metrics above, we also generate the empirical distribution of relative errors $|\frac{Dst - \hat{Dst}}{Dst}|$ for the \emph{GP-AR}, \emph{GP-ARX}, \emph{NM} and \emph{TL} models.
@@
 
-Table \ref{table:DstModels} gives a brief enumeration of the models compared, we use the performance benchmark results published in \citet{Ji2012} and to them add results obtained from testing \emph{GP-AR}, \emph{GP-ARX} as well as the \emph{Persistence} model on the same list of storms.
-
-For the purpose of generating the cumulative probability plots of the model errors, we must obtain hourly predictions for all the storm events using the \emph{NARMAX} and \emph{TL} models.
+Table \ref{table:DstModels} gives a brief enumeration of the models compared.
+We use the same performance benchmark results published in \citet{Ji2012} and we add the results obtained from testing \emph{GP-AR}, \emph{GP-ARX} as well as the \emph{Persistence} model on the same list of storms.
+
+For the purpose of generating the cumulative probability plots of the model errors, we need to obtain hourly predictions for all the storm events using the \emph{NARMAX} and \emph{TL} models.
 In the case of the \emph{NM} model, the formula outlined in \citet{balikhin:narmax} is used to generate hourly predictions.
 For the \emph{TL} model we use real time predictions listed on their website (\url{http://lasp.colorado.edu/home/spaceweather/}) corresponding to all the storm events.
-The real time predictions are at a frequency of ten minutes, hourly predictions are generated by averaging the $Dst$ predictions for each hour.
+The real time predictions are at a frequency of ten minutes and hourly predictions are generated by averaging the $Dst$ predictions for each hour.
 
@@
    \includegraphics[width=\textwidth]{Compare_CC.png}
-      \caption{Averaged cross correlation performance of GP-AR, GP-ARX and the Persistence model versus results in \citet{Ji2012}.}
+      \caption{Averaged cross correlation performance of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model versus results in \citet{Ji2012}.}
          \label{fig:cc}
@@
 
-With respect to the \emph{TL} model one important concern must be expressed.
+With respect to the \emph{TL} model one important caveat must be noted.
 The training of the \emph{TL} model was done on Omni data from 1996-2002 which overlaps with a large number of the storm events, therefore the experimental procedure has a strong bias towards \emph{TL} model.
 
+\section{Results} \label{sec:res}
+%%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages
+
 \begin{figure}
@@
    \includegraphics[width=\textwidth]{Compare_deltaDst.png}
-      \caption{Averaged $\Delta Dst_{min}$ of GP-AR, GP-ARX and the Persistence model versus results in \citet{Ji2012}.}
+      \caption{Averaged $\Delta Dst_{min}$ of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model versus results in \citet{Ji2012}.}
          \label{fig:deltaDst}
@@
 
-\section{Results} \label{sec:res}
-%%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages
-
 Figures \ref{fig:rmse}, \ref{fig:cc}, \ref{fig:deltaDst} and \ref{fig:timingErr} compare the performance of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model with the existing results of \citet{Ji2012}.
-Figure \ref{fig:rmse} compares the average RMSE of the models over the list of 63 storms, we can see that the \emph{GP-ARX} model gives an improvement of approximately $38\%$ in RMSE with respect to the \emph{TL} model and $60\%$ improvement with respect to the \emph{NM} model.
-
-Figure \ref{fig:relprob} compares the empirical distributions of absolute relative model errors across all storm events in the experiments.
-If one chooses a $95\%$ probability bound then it can be seen that the \emph{GP-ARX} model bounds the relative error to $20\%$ of the actual $Dst$ value, while in the case of the \emph{NM} model it bounds the relative error to $50\%$ of the actual $Dst$ value.
-Therefore the \emph{GP-ARX} model gives an approximate improvement of $25\%$ over the other compared models on the test data set chosen.
+Figure \ref{fig:rmse} compares the average RMSE of the predictions over the list of 63 storms.
+One can see that the \emph{GP-ARX} model gives an improvement of approximately $38\%$ in RMSE with respect to the \emph{TL} model and $60\%$ improvement with respect to the \emph{NM} model.
+It can also be seen that the \emph{Persistence} model gives better \emph{RMSE} performance than the models compared in \citet{Ji2012}.
 
@@
    \includegraphics[width=\textwidth]{Compare_timingerr.png}
-      \caption{Averaged absolute timing error of GP-AR, GP-ARX and the Persistence model versus results in \citet{Ji2012}.}
+      \caption{Averaged absolute timing error of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model versus results in \citet{Ji2012}.}
          \label{fig:timingErr}
@@
 This is also the case in \ref{fig:relprob} where we average predictions generated from a pre-trained instance of the \emph{TL} model.
-Even so the \emph{GP-ARX} model gives comparable correlation coefficient to \emph{TL} a testimony to its performance given that its training set has no overlap with the storm events.
+Even so the \emph{GP-ARX} model gives a comparable correlation coefficient to \emph{TL}, although the training set of the \emph{GP-ARX} model has no overlap with the storm events which is not the case for \emph{TL}.
 
@@
    \centering
-   \includegraphics[width=\textwidth]{Compare_pred.png}
-      \caption{Comparison of OSA predictions generated by the NM, TL and GP-ARX models for the storm event $9^{th}$ November $2004$ 11:00 UTC - $11^{th}$ November $2004$ 09:00 UTC}
-         \label{fig:predictions}
+   \includegraphics[width=\textwidth]{Compare_RelProb.png}
+   \caption{Comparison of empirical distribution of relative errors for \emph{GP-AR}, \emph{GP-ARX}, \emph{NM} and \emph{TL} on the test set of \citet{Ji2012}.}
+   \label{fig:relprob}
 \end{figure}
@@
 
-Figure \ref{fig:timingErr} compares the time discrepancy of predicting the storm peak, again as observed above the \emph{Persistence} model will always have $\Delta t_{peak} = -1$, the GP-AR and GP-ARX models give better performance than the other models in terms of timing error in prediction of storm peaks.
-
-In figure \ref{fig:predictions} we show the OSA predictions generated by the TL, NM and GP-ARX models for one event from the list of storms in \citet{Ji2012}.
-This storm is interesting due to its strength ($Dst_{min} = -289 nT$) and the presence of two distinct peaks.
-In this particular event the GP-ARX model approximates the twin peaks as well as the onset and decay phases quite faithfully which is lacking in the NM and TL model predictions.
-The TL model recognises the presence of twin peaks but over estimates the second one as well as the decay phase, while the NM model is much delayed in the prediction of the initial peak and fails to approximate the time and intensity of the second peak.
+Figure \ref{fig:timingErr} compares the time discrepancy of predicting the storm peak.
+By definition, the \emph{Persistence} model will always have $\Delta t_{peak} = -1$.
+The \emph{GP-AR} and \emph{GP-ARX} models give better performance than the other models in terms of timing error in prediction of storm peaks.
+
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Compare_pred.png}
+      \caption{Comparison of OSA predictions generated by the \emph{NM}, \emph{TL} and \emph{GP-ARX} models for the storm event $9^{th}$ November $2004$ 11:00 UTC - $11^{th}$ November $2004$ 09:00 UTC}
+         \label{fig:predictions}
+\end{figure}
+
+Figure \ref{fig:relprob} compares the empirical cumulative distribution functions of the absolute value of the relative errors across all storm events in the experiments.
+One can see that $95\%$ of the hourly predictions have a relative error smaller than $25\%$ for \emph{GP-ARX}, $50\%$ for \emph{NM} and $60\%$ for \emph{TL}.
+
+In figure \ref{fig:predictions} we show the OSA predictions generated by the \emph{TL}, \emph{NM} and \emph{GP-ARX} models for one event from the list of storms in \citet{Ji2012}.
+This storm is interesting due to its strength ($Dst_{min} = -289 \ nT$) and the presence of two distinct peaks.
+In this particular event the \emph{GP-ARX} model approximates the twin peaks as well as the onset and decay phases quite faithfully, contrary the \emph{NM} and \emph{TL} model predictions.
+The \emph{TL} model recognises the presence of twin peaks but overestimates the second one as well as the decay phase, while the \emph{NM} model is much delayed in the prediction of the initial peak and fails to approximate the time and intensity of the second peak.
 
 As noted in section \ref{sec:inference}, \emph{Gaussian Process} models generate probabilistic predictions instead of point predictions, giving the modeller the ability to specify error bars on the predictions.
-In figure \ref{fig:predictionswitherrorbar} OSA predictions along with lower and upper error bars generated by the GP-ARX model are depicted for the event used as a reference in figure \ref{fig:predictions}.
+In figure \ref{fig:predictionswitherrorbar} \emph{OSA} predictions along with lower and upper error bars generated by the \emph{GP-ARX} model are depicted for the event shown in figure \ref{fig:predictions}.
 The error bars are calculated at one standard deviation on either side of the mean prediction, using the posterior predictive distribution in equations \ref{eq:posteriormean} and \ref{eq:posteriorcov}.
+We remind that, since the model produces Gaussian output distributions, one standard deviation on either side of the mean corresponds to a confidence interval of $68\%$.
 
@@
    \includegraphics[width=\textwidth]{Compare_pred_err_bar.png}
-      \caption{OSA predictions with error bars $\pm 1$ standard deviation, generated by the GP-ARX model for the storm event $9^{th}$ November $2004$ 11:00 UTC - $11^{th}$ November $2004$ 09:00 UTC}
+      \caption{OSA predictions with error bars $\pm 1$ standard deviation, generated by the \emph{GP-ARX} model for the storm event $9^{th}$ November $2004$ 11:00 UTC - $11^{th}$ November $2004$ 09:00 UTC}
          \label{fig:predictionswitherrorbar}
@@
 
-We conclude this paper with three key conclusions/insights learned from our research so far.
+In this paper, we proposed two \emph{Gaussian Process} auto-regressive models, \emph{GP-ARX} and \emph{GP-AR}, to generate hourly predictions of $Dst$.
+We compared the performance of our proposed models with the \emph{Persistence} model and six existing model benchmarks reported in \citet{Ji2012}.
+Predictive performance was compared on \emph{RMSE}, \emph{CC}, $\Delta Dst_{min}$ and $|\Delta t_{peak}|$ calculated on a list of 63 geomagnetic storms from \cite{Ji2012}.
+
+Our results can be summarized as follows.
    \begin{enumerate}
-      \item \emph{Persistence} model must be central in the model evaluation process in the context of \emph{One Step Ahead} prediction of the $Dst$ index.
+      \item \emph{Persistence} model must be central in the model evaluation process in the context of \emph{one step ahead} prediction of the $Dst$ index.
 It is clear that the \emph{persistence} behavior in the $Dst$ values is very strong i.e.
-the trivial predictive model $\hat{Dst}(t) = Dst(t-1)$ gives excellent performance.
-In fact it can be seen in \ref{sec:res} that the \emph{Persistence} model is outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $Dst$.
-Therefore any model that is proposed to tackle the OSA prediction problem for $Dst$ must be compared to the \emph{Persistence} model and show visible gains above it.
+the trivial predictive model $\hat{Dst}(t) = Dst(t-1)$ gives excellent performance according to the metrics chosen.
+In fact it can be seen in figure \ref{sec:res} that the \emph{Persistence} model outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $Dst$.
+Therefore any model that is proposed to tackle the OSA prediction problem for $Dst$ should be compared to the \emph{Persistence} model and show visible gains above it.
 
@@
 
-      If one compares the training sets used by the models, one can appreciate that they are able to learn robust models for prediction with relatively small training and validations sets (training set contains 250 instances, while the validation set contains 432 instances) on the other hand the \emph{NM} model uses one year (8760 instances) of data to learn the formula outlined in \citet{balikhin:narmax} and the \emph{TL} model uses six years of data (52560 instances) for training.
+      If one compares the training sets used by the models, one can appreciate that they are able to learn robust models for prediction with relatively small training and validations sets: the training set contains 250 instances, while the validation set contains 432 instances.
+On the other hand the \emph{NM} model uses one year (8760 instances) of data to learn the formula outlined in \citet{balikhin:narmax} and the \emph{TL} model uses six years of data (52560 instances) for training.
 
@@
 Since the GP models generate predictive distributions for test data and not just point predictions they lend themselves to the requirements of space weather prediction very well because of the need to generate error bars on predictions.
-
-      \item Multi-step ahead prediction should be the preferred gold standard with respect to forecasting of geomagnetic indices.
+   \end{enumerate}
+
+\subsection{Future Research Directions}
+
+Multi-step ahead prediction should be the preferred gold standard with respect to forecasting of geomagnetic indices.
 Also important is the usage of more informative performance metrics for evaluating and comparing empirical predictive models.
-   \end{enumerate}
 

### 2016-05-23 swsc.tex
Updates from ShareLaTeX

@@
 \subsection{Gaussian Process Auto-Regressive with eXogenous inputs (GP-ARX)} \label{sec:gparx}
+
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{energyLandscape.png}
+      \caption{Contour plot of model selection metric (\emph{RMSE}) for \emph{GP-ARX} over the validation data.}
+         \label{fig:energy}
+   \end{figure}
 
@@
 With respect to the \emph{TL} model one important caveat must be noted.
-The training of the \emph{TL} model was done on Omni data from 1996-2002 which overlaps with a large number of the storm events, therefore the experimental procedure has a strong bias towards \emph{TL} model.
+The training of the \emph{TL} model was performed on the OMNI data from 1996-2002 which overlaps with a large number of the storm events, therefore the experimental procedure has a strong bias towards \emph{TL} model.
 

### 2016-05-24 swsc.tex
Updates from ShareLaTeX

@@
           \inst{1}
+          \and
+          S. Wing \inst{1}
           \and
@@
 Model selection of \emph{GP-AR} and \emph{GP-ARX} is performed by grid search methodology and the root mean square error (\emph{RMSE}) on the validation set is used to select the values of the model hyper-parameters.
+Figure \ref{fig:energy} shows a contour plot of the \emph{RMSE} computed on the validation data for the \emph{GP-ARX} model, the x axis represents the intercept $b$ of the polynomial kernel while the y axis represents $\sigma$ the standard deviation of the noise of the GP model as seen in section \ref{sec:inference}.
+In the experiments performed the best performing values of the hyper-parameters outputted by the grid search routing are $b = 0.002,\ \sigma = 1.994$.
+
 The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$ are arrived by experimenting with values starting from one and choosing the value of $p$ yielding best performance on the 63 events in the bench mark.

### 2016-06-05 swsc.tex
Language changes proposed by Simon

@@
           \and
+          E. Camporeale\inst{1}
+          \and 
           S. Wing \inst{1}
-          \and
-          E. Camporeale\inst{1}
           }
@@
              }
+
+   %\institute{Johns Hopkins University, Applied Physics Laboratory, \\ 11100 Johns Hopkins Road, Laurel, Maryland 20723}
 
@@
 
-Geomagnetic indices come in various forms, they may take continuous or discrete values, indices may be calculated at varying time frequencies.
+Geomagnetic indices come in various forms, they may take continuous or discrete values, indices may be calculated at varying time resolutions.
 Their values are often calculated by averaging or combining a number of readings taken by instruments around the Earth.
@@
 \begin{enumerate}
-    \item $K_p$: The Kp-index is a discrete valued global geomagnetic storm index and is based on 3 hour measurements of the K-indices, for a given value, for each of the past days.
+    \item $K_p$: The Kp-index is a discrete valued global geomagnetic storm index and is based on 3 hour measurements of the K-indices \cite{Bartels}.
 The K-index itself is a three hour long quasi-logarithmic local index of the geomagnetic activity, relative to a calm day curve for the given location.
 
-    \item $AE$: The Auroral Electrojet Index, $AE$, is designed to provide a global, quantitative measure of auroral zone magnetic activity produced by enhanced Ionospheric currents flowing below and within the auroral oval.
+    \item $AE$: The Auroral Electrojet Index, $AE$, is designed to provide a global, quantitative measure of auroral zone magnetic activity produced by enhanced Ionospheric currents flowing below and within the auroral oval \cite{AEIndex}.
 It is a continuous index which is calculated every hour.
 
-    \item $Dst$: A continuous hourly index which measures the weakening of the Earths magnetic field due to ring currents. 
+    \item $Dst$: A continuous hourly index which measures the weakening of the Earths magnetic field due to ring currents and the strength of geomagnetic storms \cite{DesslerAndParker}. 
 \end{enumerate}
@@
 %Talk about Burton and friends
-We focus on prediction of the hourly $Dst$ index which is a straightforward indicator of geomagnetic storms.
-More specifically, we focus on the \emph{one step ahead} (OSA) prediction of $Dst$ because it is the first piece towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions.
+For the present study, we focus on prediction of the hourly $Dst$ index which is a straightforward indicator of geomagnetic storms.
+More specifically, we focus on the \emph{one step ahead} (OSA), in this case one hour ahead prediction of $Dst$ because it is the first piece towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions.
 
 The $Dst$ OSA prediction problem has been the subject of several modeling efforts in the literature.
-One the earliest models has been presented by \citet{JGR:JGR10260} who calculated $Dst(t)$ as the solution of an \emph{Ordinary Differential Equation} (ODE) which expressed the rate of change of $Dst(t)$ as a combination of two terms: decay and injection $\frac{d Dst(t)}{dt} = Q(t) - \frac{Dst(t)}{\tau}$.
+One the earliest models has been presented by \citet{JGR:JGR10260} who calculated $Dst(t)$ as the solution of an \emph{Ordinary Differential Equation} (ODE) which expressed the rate of change of $Dst(t)$ as a combination of two terms: decay and injection $\frac{d Dst(t)}{dt} = Q(t) - \frac{Dst(t)}{\tau}$, where Q(t) relates to the particle injection from the plasma sheet into the inner magnetosphere.
 
@@
 \citet{Lund} proposed an \emph{Elman} recurrent network architecture called Lund $Dst$, which used the solar wind velocity, \emph{interplanetary magnetic field} (IMF) and historical $Dst$ data as inputs.
+\citet{JGRA:JGRA17461} used the recurrent neural networks to predict $Kp$.
 \citet{SWE:SWE286} originally proposed a \emph{feed forward} network for predicting the $K_p$ index which used the \emph{Boyle coupling function} [see \citet{boyle1997empirical}].
@@
 %Vision
-In $Dst$ prediction a seemingly trivial yet highly informative prediction method is represented by the so called \emph{Persistence} model this is simply using the previous value of $Dst$ as the prediction for the next time step ($\hat{D}st(t) = Dst(t-1)$).
+In $Dst$ prediction, a seemingly trivial yet highly informative prediction method is represented by the so called \emph{Persistence} model, which uses the previous value of $Dst$ as the prediction for the next time step ($\hat{D}st(t) = Dst(t-1)$).
 Due to high correlation between $Dst$ values one hour apart ($Dst(t)$, $Dst(t-1)$), despite its apparent lack of sophistication the \emph{Persistence} model gives excellent predictive performance on error metrics.
@@
 
-Here $\matbf{f}$ is an $N\times 1$ vector consisting of the values $f(\mathbf{x}_i), i = q \cdots N$.
+Here $\mathbf{f}$ is an $N\times 1$ vector consisting of the values $f(\mathbf{x}_i), i = q \cdots N$.
 In equation \ref{eq:normal} $\mathbf{f}|\mathbf{x}_1, \cdots, \mathbf{x}_N$ denotes the conditional distribution of $\mathbf{f}$ with respect to the input data (i.e.
@@
 Auto-regressive models can be improved by including exogenous quantities that can help in improving predictions.
-It is well known that the solar wind and interplanetary magnetic field are important factors that influence the Earth's geomagnetic state (\citet{JGRA:JGRA11775}, \citet{Russell1974}).
+It is well known that the solar wind and interplanetary magnetic field are important factors that influence the Earth's geomagnetic state (\citet{JGRA:JGRA11775}, \citet{Russell1974}, \citet{JGRA:JGRA17574}).
 

### 2016-06-05 swsc.tex
Simons changes 2

@@
 
-   \authorrunning{Chandorkar and Camporeale}
+   \authorrunning{Chandorkar et. al}
 
@@
           \and 
-          S. Wing \inst{1}
-          }
-
-   \institute{Centrum Wiskunde Informatica (CWI), Amsterdam,
+          S. Wing \inst{2}
+          }
+
+   \institute{\inst{1} Centrum Wiskunde Informatica (CWI), Amsterdam,
               1098XG Amsterdam\\
-              \email{\href{mailto:m.h.chandorkar@cwi.nl}{m.h.chandorkar@cwi.nl} \href{mailto:e.camporeale@cwi.nl}{e.camporeale@cwi.nl}} 
-             }
-
-   %\institute{Johns Hopkins University, Applied Physics Laboratory, \\ 11100 Johns Hopkins Road, Laurel, Maryland 20723}
+              \ \email{\href{mailto:m.h.chandorkar@cwi.nl}{m.h.chandorkar@cwi.nl} \href{mailto:e.camporeale@cwi.nl}{e.camporeale@cwi.nl}}\\
+              \inst{2} The Johns Hopkins University Applied Physics Laboratory, 
+              Laurel, Maryland, 20723, USA
+             }
 

### 2016-06-07 swsc.tex
Added Simon's references

@@
           \and 
-          S. Wing \inst{2}
+          S. Wing \inst{1,2}
           }
@@
 Auto-regressive models can be improved by including exogenous quantities that can help in improving predictions.
-It is well known that the solar wind and interplanetary magnetic field are important factors that influence the Earth's geomagnetic state (\citet{JGRA:JGRA11775}, \citet{Russell1974}, \citet{JGRA:JGRA17574}).
+$Dst$ gives a measure of ring currents, which are modulated by plasma sheet particle injections into the inner magnetosphere during sub-storms.
+Studies have shown that the substorm occurrence rate increases with solar wind velocity (high speed streams) (\citet{Kissinger2011}, \citet{Newell2016}).
+Prolonged southward interplanetary magnetic field (IMF) z-component ($B_z$) is needed for sub-storm to occur (\citet{McPherron1986}).
+An increase in the solar wind electric field, $VB_z$, can increase the dawn-dusk electric field in the magnetotail, which in turn determines the amount of plasma sheet particle that move to the inner magnetosphere (\citet{Friedel2001}).
+Therefore, our exogenous parameters consist of solar wind velocity, IMF $B_z$, and $VB_z$.
+
+%It is well known that the solar wind and interplanetary magnetic field are important factors that influence the Earth's geomagnetic state (\citet{JGRA:JGRA11775}, \citet{Russell1974}, \citet{JGRA:JGRA17574}).
 
@@
 \begin{thebibliography}{}
-
-   \bibitem[Baker(1966)]{baker} Baker, N., Solar Evolution,
-      in Stellar Evolution, ed.\ R. F. Stein and A. G. W. Cameron,
-      Plenum, New York, 333, 1966.
-
-   \bibitem[Balluch(1988)]{balluch} Balluch, M., A paper on something, 
-      {\it Astron. Astrophys}, {\bf 200}, 58-63, 1988.
-
-   \bibitem[Cox(1980)]{cox} Cox, J. P.,
-      Theory of Stellar Pulsation,
-      Princeton University Press, Princeton, 1980.
-
-   \bibitem[Cox and Stewart(1969)]{cox69} Cox, A. N., and J. N. Stewart,
-      Academia Nauk, {\it Scientific Information}, {\bf 15}, 1-11, 1969.
-
-   \bibitem[Dumm and Kopf(2014)]{dummkopf} Dumm, Z., and Z. Kopf, A new self-inconsistent 
-      core stability model, 
-      {\it J. Space Weather Space Clim.}, {\bf 4}, A00, 2014, DOI: 10.1051/swsc/2014000,
-      \url{http://dx.doi.org/10.1051/swsc/2014000}.
-
-   \bibitem[Fool(2012)]{fool} Fool, X. Y., All earlier results are wrong, 
-      {\it J. Space Weather Space Clim.}, {\bf 2}, A00, 2012, DOI: 10.1051/swsc/2012000,
-      \url{http://dx.doi.org/10.1051/swsc/2012000}.
-
-   \bibitem[Fou(2013)]{fou} Fou, Y. X., No earlier results were right, 
-      {\it J. Space Weather Space Clim.}, {\bf 3}, A00, 2013, DOI: 10.1051/swsc/2013000,
-      \url{http://dx.doi.org/10.1051/swsc/2013000}.
-
-   \bibitem[Kr\"ugel et al.(1971)]{krugel71} Kr\"ugel, A. H., A. F. Davidsen, D. Tytler, and G. A. Kriss, 
-      Summary of Project Achievements, Technical Report 08-11, 
-      Fake Company, Nowherecity, Nowhereland, 1971.
-
-   \bibitem[Mizuno(1980)]{mizuno} Mizuno, H., A further step backward in scientific understanding, 
-      {\it Prog. Theor. Phys.}, {\bf 64}, 544-545, 1980.
-
-   \bibitem[Tscharnuter(1987)]{tscharnuter} Tscharnuter, W. M., Who knows about astronomy?, 
-      {\it Astron. Astrophys}, {\bf 188}, 55-57, 1987.
-
-   \bibitem[Yorke(1980a)]{yorke80a} Yorke, H. W., A paper on nothing, 
-      {\it Astron. Astrophys}, {\bf 86}, 286, 1980a.
-
-   \bibitem[Yorke(1980b)]{yorke80b} Yorke, H. W., A paper on everything, 
-      {\it Astron. Astrophys}, {\bf 86}, 291, 1980b.
-
 \end{thebibliography}

### 2016-06-14 swsc.tex
Updates from ShareLaTeX

@@
 Hyper-parameters give flexibility to particular kernel structure, for example $d = 1, 2, 3, \cdots$ in the polynomial kernel represents linear, quadratic, cubic and higher order polynomials respectively.
-The method of assigning values to the \emph{hyper-parameter} is crucial in the model building process (see section \ref{sec:hyp} below).
-In this study, we construct Gaussian Process regression models with linear polynomial kernels for \emph{one step ahead} prediction of the $Dst$ index.
-
-\subsection{Optimizing Hyper-parameters} \label{sec:hyp}
-
-When applying \emph{Gaussian Process} models to data, one must choose appropriate values to the kernel hyper-parameters and noise level ($\sigma^2$).
-This is known as the problem of \emph{model selection}.
-
-Common examples of criteria used for \emph{model selection} include data likelihood, cross validation performance scores and validation set performance scores among others.
-Since model selection criteria are non-convex and non-smooth in general, gradient free approaches are used to find approximate values of local optima.
-
-The simplest way to perform \emph{model selection} is called \emph{grid search} which entails evaluating the above criteria on a fixed grid which is constructed by the tensorial product of grids constructed on each hyper-parameter.
-There are more sophisticated methods for \emph{model tuning} such as \emph{coupled simulated annealing} (\citet{Xavier-De-Souza2010}) and \emph{simplex search} (\citet{Nelder1965}) but they are computationally more expensive.
-
-In this study we employ \emph{grid search} to find the best performing set of hyper-parameter assignments inside a predefined grid of possible hyper-parameter values.
+The method of assigning values to the \emph{hyper-parameter} is crucial in the model building process.
+
+In this study, we construct Gaussian Process regression models with linear kernels as shown in equation \ref{eq:usedKernel}, for \emph{one step ahead} prediction of the $Dst$ index.
+Our choice of kernel leaves us with only one adjustable hyper-parameter, the model noise $\sigma$.
+
+\begin{equation}\label{eq:usedKernel}
+    K(\mathbf{x}, \mathbf{y}) = \mathbf{x}^\intercal \mathbf{y}
+\end{equation}
+
+We initialize a grid of values for $\sigma$ the model noise and use the error on a predefined validation data set to choose the best performing value of $\sigma$.
+While constructing the grid of possible values for $\sigma$ it must be ensured that $\sigma > 0$ so that the kernel matrix constructed on the training data is non-singular.
 
@@
 
-\begin{figure}
-   \centering
-   \includegraphics[width=\textwidth]{energyLandscape.png}
-      \caption{Contour plot of model selection metric (\emph{RMSE}) for \emph{GP-ARX} over the validation data.}
-         \label{fig:energy}
-   \end{figure}
-
 Auto-regressive models can be improved by including exogenous quantities that can help in improving predictions.
@@
 Model selection of \emph{GP-AR} and \emph{GP-ARX} is performed by grid search methodology and the root mean square error (\emph{RMSE}) on the validation set is used to select the values of the model hyper-parameters.
-Figure \ref{fig:energy} shows a contour plot of the \emph{RMSE} computed on the validation data for the \emph{GP-ARX} model, the x axis represents the intercept $b$ of the polynomial kernel while the y axis represents $\sigma$ the standard deviation of the noise of the GP model as seen in section \ref{sec:inference}.
-In the experiments performed the best performing values of the hyper-parameters outputted by the grid search routing are $b = 0.002,\ \sigma = 1.994$.
+In the experiments performed the best performing values of the hyper-parameters outputted by the grid search routine are $\sigma = 0.45$ for the \emph{GP-ARX} and $\sigma = 1.173$ for \emph{GP-AR}.
 

### 2016-06-20 swsc.tex
Draft 2016/06/20

@@
           \and 
-          S. Wing \inst{1,2}
+          S. Wing \inst{2}
           }
@@
    \abstract{
-    We present two models that improve the state of the art of \emph{One Step Ahead}(OSA) prediction of the $Dst$ geomagnetic activity index.
-In the domain of OSA prediction, it is well known that $Dst(t)$ values correlate strongly with values one hour in the past (i.e.
-$Dst(t-1)$).
-While a number of forecasting models have been proposed and compared, none of them have been compared with the \emph{Persistence} model to quantify their performance improvements above this most trivial of predictors.
-In this study we train Gaussian Process regression models for One Step Ahead (OSE) prediction of the Disturbance Storm Time ($Dst$) geomagnetic index.
-We propose two variants of the Gaussian Process model.
-    \begin{enumerate*}
-      \item Gaussian Process Auto regressive (GP-AR) 
-      \item Gaussian Process Auto regressive with eXogenous inputs (GP-ARX). 
-    \end{enumerate*}
+    We present two models, based on Gaussian Processes, for the \emph{One Step Ahead}(OSA) prediction of the $Dst$ geomagnetic activity index. 
+    The models are Auto regressive, with or without eXogenous inputs (GP-ARX and GP-AR, respectively).
     We compare the performance of these models with the current state of the art in one step ahead $Dst$ prediction on a set of 
-    63 benchmark storms from 1998-2006. The proposed Gaussian Process models not only outperform the rest but also the \emph{Persistence} model.
+    63 benchmark storms from 1998-2006, previously analyzed in an earlier study.
+    We show that, despite its lack of sophistication, the so-called Persistence model represents an important test to evaluate the performance of a one hour ahead $Dst$ model.
+    Contrary to the state-of-the-art models compared in the literature, our models consistently outperform the Persistence model and represent a substantial improvement in the field, when evaluated on standard metrics.
+    Finally, an important feature of the new models is that they naturally provide confidence intervals for their forecast.
     }
@@
 
-Geomagnetic indices come in various forms, they may take continuous or discrete values, indices may be calculated at varying time resolutions.
+Geomagnetic indices come in various forms, they may take continuous or discrete values and may be defined with varying time resolutions.
 Their values are often calculated by averaging or combining a number of readings taken by instruments around the Earth.
-Each geomagnetic index is a proxy for a particular kind of phenomena.
+Each geomagnetic index is a proxy for a particular kind of phenomenon.
 Some popular indices are the $K_p$, $Dst$ and the $AE$ index.
@@
 \begin{enumerate}
-    \item $K_p$: The Kp-index is a discrete valued global geomagnetic storm index and is based on 3 hour measurements of the K-indices \cite{Bartels}.
+    \item $K_p$: The Kp-index is a discrete valued global geomagnetic storm index and is based on 3 hour measurements of the K-indices \citep{Bartels}.
 The K-index itself is a three hour long quasi-logarithmic local index of the geomagnetic activity, relative to a calm day curve for the given location.
 
-    \item $AE$: The Auroral Electrojet Index, $AE$, is designed to provide a global, quantitative measure of auroral zone magnetic activity produced by enhanced Ionospheric currents flowing below and within the auroral oval \cite{AEIndex}.
+    \item $AE$: The Auroral Electrojet Index, $AE$, is designed to provide a global, quantitative measure of auroral zone magnetic activity produced by enhanced Ionospheric currents flowing below and within the auroral oval \citep{AEIndex}.
 It is a continuous index which is calculated every hour.
 
-    \item $Dst$: A continuous hourly index which measures the weakening of the Earths magnetic field due to ring currents and the strength of geomagnetic storms \cite{DesslerAndParker}. 
+    \item $Dst$: A continuous hourly index which measures the weakening of the Earths magnetic field due to ring currents and the strength of geomagnetic storms \citep{DesslerAndParker}. 
 \end{enumerate}
@@
 For the present study, we focus on prediction of the hourly $Dst$ index which is a straightforward indicator of geomagnetic storms.
-More specifically, we focus on the \emph{one step ahead} (OSA), in this case one hour ahead prediction of $Dst$ because it is the first piece towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions.
+More specifically, we focus on the \emph{one step ahead} (OSA), in this case one hour ahead prediction of $Dst$ because it is the simplest model towards building long term predictions of geomagnetic response of the Earth to changing space weather conditions.
 
 The $Dst$ OSA prediction problem has been the subject of several modeling efforts in the literature.
-One the earliest models has been presented by \citet{JGR:JGR10260} who calculated $Dst(t)$ as the solution of an \emph{Ordinary Differential Equation} (ODE) which expressed the rate of change of $Dst(t)$ as a combination of two terms: decay and injection $\frac{d Dst(t)}{dt} = Q(t) - \frac{Dst(t)}{\tau}$, where Q(t) relates to the particle injection from the plasma sheet into the inner magnetosphere.
+One of the earliest models has been presented by \citet{JGR:JGR10260} who calculated $Dst(t)$ as the solution of an \emph{Ordinary Differential Equation} (ODE) which expressed the rate of change of $Dst(t)$ as a combination of two terms: decay and injection $\frac{d Dst(t)}{dt} = Q(t) - \frac{Dst(t)}{\tau}$, where Q(t) relates to the particle injection from the plasma sheet into the inner magnetosphere.
 
@@
 The NARMAX methodology builds models by constructing polynomial expansions of inputs and determines the best combinations of monomials to include in the refined model by using a criterion called the \emph{error reduction ratio} (ERR).
-The parameters of the so called NARMAX OLS-ERR model are calculated using the standard \emph{ordinary least squares} (OLS) method to minimize an objective function.
+The parameters of the so called NARMAX OLS-ERR model are calculated by solving the \emph{ordinary least squares} (OLS) problem arising from a quadratic objective objective function.
 The reader may refer to \citet{billings2013nonlinear} for a detailed exposition of the NARMAX methodology.
@@
 \citet{Lund} proposed an \emph{Elman} recurrent network architecture called Lund $Dst$, which used the solar wind velocity, \emph{interplanetary magnetic field} (IMF) and historical $Dst$ data as inputs.
-\citet{JGRA:JGRA17461} used the recurrent neural networks to predict $Kp$.
-\citet{SWE:SWE286} originally proposed a \emph{feed forward} network for predicting the $K_p$ index which used the \emph{Boyle coupling function} [see \citet{boyle1997empirical}].
-The same architecture is adapted for prediction of $Dst$ in \citet{SWE:SWE286}, which is popularly known as the Rice $Dst$ model.
+\citet{JGRA:JGRA17461} used recurrent neural networks to predict $Kp$.
+\citet{SWE:SWE286} originally proposed a \emph{feed forward} network for predicting the $K_p$ index which used the \emph{Boyle coupling function} \citep{boyle1997empirical}.
+The same architecture is adapted for prediction of $Dst$ in \citet{SWE:SWE286}, popularly known as the Rice $Dst$ model.
 \citet{pallocchia:hal-00318011} proposed a \emph{neural network} model called EDDA to predict $Dst$ using only the IMF data.
@@
 %Talk about comparison papers, number of storms covered and Persistence
-In light of the extensive list of modeling techniques employed for prediction of the $Dst$ index, model comparison and evaluation becomes a crucial direction in advancing the research domain.
-\citet{Rastatter2013} compare several physics based (convection, kinetic and magneto hydrodynamic) and empirical prediction models such as NARMAX, Rice $Dst$ on 4 storm events which occurred between 2001 and 2006.
-\citet{Amata2008} compare the EDDA and the Lund $Dst$ models over the 2003-2005 period.
+In light of the extensive list of modeling techniques employed for prediction of the $Dst$ index, model comparison and evaluation becomes a crucial step for advancing the research domain.
+\citet{Rastatter2013} compared several physics based (convection, kinetic and magneto hydrodynamic) and empirical prediction models such as NARMAX, Rice $Dst$ on 4 storm events which occurred between 2001 and 2006.
+\citet{Amata2008} compared the EDDA and the Lund $Dst$ models over the 2003-2005 period.
 The most extensive model comparison in terms of storm events was probably conducted in \citet{Ji2012} which compared six $Dst$ models (see table \ref{table:DstModels}) on a list of 63 geomagnetic storm events of varying intensities which occurred between 1998 and 2006.
@@
 In $Dst$ prediction, a seemingly trivial yet highly informative prediction method is represented by the so called \emph{Persistence} model, which uses the previous value of $Dst$ as the prediction for the next time step ($\hat{D}st(t) = Dst(t-1)$).
-Due to high correlation between $Dst$ values one hour apart ($Dst(t)$, $Dst(t-1)$), despite its apparent lack of sophistication the \emph{Persistence} model gives excellent predictive performance on error metrics.
+Due to high correlation between $Dst$ values one hour apart ($Dst(t)$, $Dst(t-1)$), the \emph{Persistence} model gives excellent predictive performance on error metrics, despite its apparent lack of sophistication.
 In essence it is a trivial hypothesis and hence it should always be used as a base line to compare the performance of any proposed $Dst$ algorithm.
 Moreover, given its zero computational cost, a model should at least outperform the \emph{Persistence} model in order to be a viable candidate for advancing the science of space weather prediction.
+An obvious critique to the Persistence model is that it is not really predictive, in the sense that it cannot forecast a storm, until it has already commenced.
+However, most of the literature has used global metrics, such as the Root Mean Square Error (RMSE, see section \ref{sec:exp}) to evaluate models.
+It is in this context that we argue that the Persistence model should be regarded as the first candidate to outperform.
 
@@
 In this paper, we propose \emph{Gaussian Process} models for OSA prediction of $Dst$.
-We use the results of \citet{Ji2012} as a starting point and compare our proposed models with the methods outlined in the same paper while using the performance of the \emph{Persistence} model as a base line.
+We use the results of \citet{Ji2012} as a starting point and compare our proposed models with the methods evaluated in that paper, while using the performance of the \emph{Persistence} model as a base line.
 
 % Talk about Gaussian Process
-We use the hourly resolution data as extracted from NASA/GSFC's OMNI data set through OMNIWeb (see \cite{OmniPaper}).
+We use hourly resolution measurements extracted from NASA/GSFC's OMNI data set through OMNIWeb \citep{OmniPaper}.
 The rest of the paper is organized as follows.
-Section \ref{sec:method} gives an introduction to the \emph{Gaussian Process} methodology for regression and some technical details on their application.
-In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res}, bench marked against the \emph{Persistence} $Dst$ model as well as the other models outlined in \citet{Ji2012}.
+Section \ref{sec:method} gives an introduction to the \emph{Gaussian Process} methodology for regression and some technical details about their application.
+In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res}, bench marked against the \emph{Persistence} as well as the other models outlined in \citet{Ji2012}.
 
@@
 
-On the basis of a set of labeled data points ${(\mathbf{x}_i, y_i); i = 1 \cdots N}$ which can be conveniently expressed by $\mathbf{X}$ a $N \times d$ data matrix and $\mathbf{y}$ a $N \times 1$ response vector as shown in equations \ref{eq:feat} and \ref{eq:labels}.
+A set of labeled data points ${(\mathbf{x}_i, y_i); i = 1 \cdots N}$ can be conveniently expressed by a $N \times d$ data matrix $\mathbf{X}$ and a $N \times 1$ response vector $\mathbf{y}$, as shown in equations (\ref{eq:feat}) and (\ref{eq:labels}).
 
@@
 
-Out task is to infer the the values of the unknown function $f(.)$ based on the inputs $\mathbf{X}$ and the noisy observations $\mathbf{y}$.
-We now assume that the joint distribution of $f(\mathbf{x}_i), i = 1 \cdots N$ is multivariate Gaussian as shown in equations \ref{eq:fvalues}, \ref{eq:normal} and \ref{eq:sto}.
+Our task is to infer the values of the unknown function $f(.)$ based on the inputs $\mathbf{X}$ and the noisy observations $\mathbf{y}$.
+We now assume that the joint distribution of $f(\mathbf{x}_i), i = 1 \cdots N$ is a multivariate Gaussian as shown in equations (\ref{eq:fvalues}), (\ref{eq:normal}) and (\ref{eq:sto}).
 
@@
 
-Here $\mathbf{f}$ is an $N\times 1$ vector consisting of the values $f(\mathbf{x}_i), i = q \cdots N$.
-In equation \ref{eq:normal} $\mathbf{f}|\mathbf{x}_1, \cdots, \mathbf{x}_N$ denotes the conditional distribution of $\mathbf{f}$ with respect to the input data (i.e.
-$\mathbf{X}$) and $\mathcal{N}\left( \mathbf{\mu}, \mathbf{\Lambda} \right)$ represents a multivariate Gaussian distribution with mean vector $\mathbf{\mu}$ and covariance matrix $\mathbf{\Lambda}$.
-The probability density function of this distribution $p( \mathbf{f} \ | \ \mathbf{x}_1, \cdots, \mathbf{x}_N)$ is therefore given by equation \ref{eq:sto}.
+Here $\mathbf{f}$ is a $N\times 1$ vector consisting of the values $f(\mathbf{x}_i), i = 1 \cdots N$.
+In equation (\ref{eq:normal}), $\mathbf{f}|\mathbf{x}_1, \cdots, \mathbf{x}_N$ denotes the conditional distribution of $\mathbf{f}$ with respect to the input data (i.e., $\mathbf{X}$) and $\mathcal{N}\left( \mathbf{\mu}, \mathbf{\Lambda} \right)$ represents a multivariate Gaussian distribution with mean vector $\mathbf{\mu}$ and covariance matrix $\mathbf{\Lambda}$.
+The probability density function of this distribution $p( \mathbf{f} \ | \ \mathbf{x}_1, \cdots, \mathbf{x}_N)$ is therefore given by equation (\ref{eq:sto}).
 
-From equation \ref{eq:sto}, one can observe that in order to uniquely define the distribution of the process, it is required to specify $\mathbf{\mu}$ and $\mathbf{\Lambda}$.
-For this probability density to be valid, there is a further requirements imposed on $\mathbf{\Lambda}$:
+From equation (\ref{eq:sto}), one can observe that in order to uniquely define the distribution of the process, it is required to specify $\mathbf{\mu}$ and $\mathbf{\Lambda}$.
+For this probability density to be valid, there are further requirements imposed on $\mathbf{\Lambda}$:
 
 \begin{enumerate}
-      \item Symmetry: $\mathbf{\Lambda}_{ij} = \mathbf{\Lambda}_{ij} \ \forall i,j \in {1, \cdots, N} $ 
+      \item Symmetry: $\mathbf{\Lambda}_{ij} = \mathbf{\Lambda}_{ji} \ \forall i,j \in {1, \cdots, N} $ 
       \item Positive Semi-definiteness: $\mathbf{z}^T \mathbf{\Lambda} \mathbf{z} \geq 0 \ \forall \mathbf{z} \in \mathbb{R}^N$  
@@
 In most practical applications of \emph{Gaussian Processes} the mean function is often defined as $m(\mathbf{x}) = 0$, which is not unreasonable if the data is standardized to have zero mean.
-\emph{Gaussian Processes} are represented in machine learning literature using the following notation.
+\emph{Gaussian Processes} are represented in machine learning literature using the following notation:
 
@@
 Our aim is to infer the function $f(\mathbf{x})$ from the noisy training data and generate predictions $f(\mathbf{x}^{*}_i)$ for a set of test points $ {\mathbf{x}^{*}_i : \forall i \in 1, \cdots, M} $.
-We define $\mathbf{X}^*$ as the test data matrix whose rows are formed by $\mathbf{x}^{*}_i$ as shown in equation \ref{eq:testfeat}.
-Using the multivariate Gaussian distribution in equation \ref{eq:sto} we can construct the distribution of $f(\mathbf{x})$ over the training and test points.
+We define $\mathbf{X}^*$ as the test data matrix whose rows are formed by $\mathbf{x}^{*}_i$ as shown in equation (\ref{eq:testfeat}).
+\begin{equation}
+    \mathbf{X}_* = \left( \begin{array}{c} (\mathbf{x}^{*}_1)^T \\ (\mathbf{x}^{*}_2)^T \\ \vdots \\ (\mathbf{x}^{*}_M)^T \end{array} \right)_{M \times d} \label{eq:testfeat} 
+\end{equation}
+Using the multivariate Gaussian distribution in equation (\ref{eq:sto}) we can construct the joint distribution of $f(\mathbf{x})$ over the training and test points.
 The vector of training and test outputs $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right)$ is of dimension $(N+M) \times 1$ and is constructed by appending the test set predictions $\mathbf{f}_*$ to the observed noisy measurements $\mathbf{y}$.
@@
 \begin{eqnarray}
-    \mathbf{X}_* & = & \left( \begin{array}{c} (\mathbf{x}^{*}_1)^T \\ (\mathbf{x}^{*}_2)^T \\ \vdots \\ (\mathbf{x}^{*}_M)^T \end{array} \right)_{M \times d} \label{eq:testfeat} \\
     \mathbf{f}_* & = & \left( \begin{array}{c} f(\mathbf{x^{*}_1}) \\ f(\mathbf{x^{*}_2}) \\ \vdots \\ f(\mathbf{x^{*}_M}) \end{array} \right)_{M \times 1} \\
@@
 
-Since we have noisy measurements of $f$ over the training data, we add the noise variance $\sigma^2$ to the variance of $f$ as shown in \ref{eq:dist}.
+Since we have noisy measurements of $f$ over the training data, we add the noise variance $\sigma^2$ to the variance of $f$ as shown in (\ref{eq:dist}).
 The block matrix components of the $(N+M) \times (N+M)$ covariance matrix have the following structure.
@@
       \item $\mathbf{I}$: The $n \times n$ identity matrix.
-      \item $\mathbf{K} = [K(\mathbf{x}_i, \mathbf{x}_j)], \ i,j \in 1,\cdots,n$ : Kernel matrix constructed from all couples of the training data.
+      \item $\mathbf{K} = [K(\mathbf{x}_i, \mathbf{x}_j)], \ i,j \in 1,\cdots,n$ : Kernel matrix constructed from all couples obtained from the training data.
       \item $\mathbf{K}_{*} = [K(\mathbf{x}_i, \mathbf{x}^{*}_j)], \ i \in 1,\cdots,n ; j \in 1,\cdots,m$ : Cross kernel matrix constructed from all couples between training and test data points.
-      \item $\mathbf{K}_{**} = [K(\mathbf{x}^{*}_i, \mathbf{x}^{*}_j)], \ i,j \in 1,\cdots,m$: Kernel matrix constructed from all couples of the test data.
+      \item $\mathbf{K}_{**} = [K(\mathbf{x}^{*}_i, \mathbf{x}^{*}_j)], \ i,j \in 1,\cdots,m$: Kernel matrix constructed from all couples obtained from the test data.
 \end{enumerate}
 
-With the multivariate normal distribution defined in equation \ref{eq:dist}, probabilistic predictions $f_*$ can be generated by constructing the conditional distribution $\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}$.
-Since the original distribution of $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right) | \ \ \mathbf{X}, \mathbf{X}_*$ is a multivariate Gaussian, conditioning on a subset of elements $\mathbf{y}$ yields another Gaussian distribution whose mean and covariance can be calculated exactly in equation \ref{eq:posterior} (see \citet{Rasmussen:2005:GPM:1162254}).
-
+With the multivariate normal distribution defined in equation (\ref{eq:dist}), probabilistic predictions $f_*$ can be generated by constructing the conditional distribution $\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}$.
+Since the original distribution of $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right) | \ \ \mathbf{X}, \mathbf{X}_*$ is a multivariate Gaussian, conditioning on a subset of elements: $\mathbf{y}$ yields another Gaussian distribution whose mean and covariance can be calculated exactly, as in equation (\ref{eq:posterior}) (see \citet{Rasmussen:2005:GPM:1162254}).
+\begin{equation}
+    \mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*} \sim \mathcal{N}(\mathbf{\bar{f}_*}, \Sigma_*)  \label{eq:posterior},
+\end{equation}
+where
 \begin{eqnarray}
-    \mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*} & \sim & \mathcal{N}(\mathbf{\bar{f}_*}, \Sigma_*)  \label{eq:posterior} \\
     \mathbf{\bar{f}_*} & = & \mathbf{K}^T_{*} [\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y} \label{eq:posteriormean} \\
@@
 The practical implementation of \emph{Gaussian Process} models requires the inversion of the training data kernel matrix $[\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1}$ to calculate the parameters of the predictive distribution $\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}$.
-The computational complexity of this inference is dominated by the linear problem $\mathbf{K}^T_{*} [\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y}$, this can be done via Cholesky decomposition which has time complexity $O(N^3)$.
+The computational complexity of this inference is dominated by the linear problem $\mathbf{K}^T_{*} [\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y}$, which can be solved via Cholesky decomposition, with a time complexity of $O(N^3)$, where $N$ is the number of data points.
 
 The distribution of $\mathbf{f_*}| \mathbf{X},\mathbf{y},\mathbf{X_*}$ is known in Bayesian analysis as the \emph{Posterior Predictive Distribution}.
-This illustrates a key difference between \emph{Gaussian Processes} and other regression models such as \emph{Neural Networks}, \emph{Linear Models} and \emph{Support Vector Machines}, that a \emph{Gaussian Process} model does not generate point predictions for new data but outputs a predictive distribution for the quantity sought, this allows us to construct error bars on the predictions.
-This property of Bayesian models such as \emph{Gaussian Processes} makes them suitable for Space Weather forecasting applications.
+This illustrates a key difference between \emph{Gaussian Processes} and other regression models such as \emph{Neural Networks}, \emph{Linear Models} and \emph{Support Vector Machines}: a \emph{Gaussian Process} model does not generate point predictions for new data but outputs a predictive distribution for the quantity sought, thus allowing to construct error bars on the predictions.
+This property of Bayesian models such as \emph{Gaussian Processes} makes them very appealing for Space Weather forecasting applications.
 
-The central design issue in applying \emph{Gaussian Process} models is choice of the function $K(\mathbf{x}, \mathbf{x}')$.
-The same constraints that apply to $\mathbf{\Lambda}$ also apply to the function $K(.,.)$.
+The central design issue in applying \emph{Gaussian Process} models is the choice of the function $K(\mathbf{x}, \mathbf{x}')$.
+The same constraints that apply to $\mathbf{\Lambda}$ also apply to the function $K$.
 In machine learning, these symmetric positive definite functions of two variables are known as \emph{kernels}.
@@
 
-For the success of a Gaussian Process model an appropriate choice of kernel function is paramount.
+For the success of a \emph{Gaussian Process} model an appropriate choice of kernel function is paramount.
 The symmetry and positive semi-definiteness of \emph{Gaussian Process} kernels implies that they represent inner-products between some basis function representation of the data.
-The interested reader is suggested to refer to \cite{Berlinet2004}, \citet{Scholkopf:2001:LKS:559923} and \citet{hofmann2008} for a thorough treatment on kernel functions and the rich theory behind them.
+The interested reader is suggested to refer to \cite{Berlinet2004}, \citet{Scholkopf:2001:LKS:559923} and \citet{hofmann2008} for a thorough treatment of kernel functions and the rich theory behind them.
 Some common kernel functions used in machine learning include the radial basis function (RBF) kernel $K(\mathbf{x}, \mathbf{y}) = \frac{1}{2} exp(-||\mathbf{x} - \mathbf{y}||^2/l^2)$ and the polynomial kernel $K(\mathbf{x}, \mathbf{y}) = (\mathbf{x}^\intercal \mathbf{y} + b)^d$.
 
-The quantities $l$ in the RBF, $b$ and $d$ in the polynomial kernel are known as \emph{hyper-parameters}.
+The quantities $l$ in the RBF, and $b$ and $d$ in the polynomial kernel are known as \emph{hyper-parameters}.
 Hyper-parameters give flexibility to particular kernel structure, for example $d = 1, 2, 3, \cdots$ in the polynomial kernel represents linear, quadratic, cubic and higher order polynomials respectively.
-The method of assigning values to the \emph{hyper-parameter} is crucial in the model building process.
+The method of assigning values to the \emph{hyper-parameters} is crucial in the model building process.
 
-In this study, we construct Gaussian Process regression models with linear kernels as shown in equation \ref{eq:usedKernel}, for \emph{one step ahead} prediction of the $Dst$ index.
+In this study, we construct Gaussian Process regression models with linear kernels as shown in equation (\ref{eq:usedKernel}), for \emph{one step ahead} prediction of the $Dst$ index.
 Our choice of kernel leaves us with only one adjustable hyper-parameter, the model noise $\sigma$.
@@
 
-We initialize a grid of values for $\sigma$ the model noise and use the error on a predefined validation data set to choose the best performing value of $\sigma$.
+We initialize a grid of values for the model noise $\sigma$ and use the error on a predefined validation data set to choose the best performing value of $\sigma$.
 While constructing the grid of possible values for $\sigma$ it must be ensured that $\sigma > 0$ so that the kernel matrix constructed on the training data is non-singular.
@@
 
-Below in equations \ref{eq:Dst}, \ref{eq:GPNoise}, \ref{eq:DstGP}, \ref{eq:GPPoly} we outline a \emph{Gaussian Process} formulation for \emph{OSA} prediction of $Dst$.
-A vector of features $x_{t-1}$ is used as input to an unknown function $f(x_{t-1})$ which drawn from a \emph{Gaussian Process} driven by a linear polynomial kernel.
+Below in equations (\ref{eq:Dst}) - (\ref{eq:GPPoly}) we outline a \emph{Gaussian Process} formulation for \emph{OSA} prediction of $Dst$.
+A vector of features $x_{t-1}$ is used as input to an unknown function $f(x_{t-1})$.
 
-The features $x_{t-1}$ can be any collection of quantities in the hourly resolution Omni data set.
-Generally $x_{t-1}$ are time histories of $Dst$ or other important variables examples being plasma pressure $p(t)$, solar wind speed $V(t)$, z component of the interplanetary magnetic field $B_z(t)$.
+The features $x_{t-1}$ can be any collection of quantities in the hourly resolution OMNI data set.
+Generally $x_{t-1}$ are time histories of $Dst$ and other important variables such as plasma pressure $p(t)$, solar wind speed $V(t)$, z component of the interplanetary magnetic field $B_z(t)$.
 
@@
 
-In \ref{sec:gpar} and \ref{sec:gparx} below we consider two choices for the input features $\mathbf{x}_{t-1}$ leading to two variants of \emph{Gaussian Process} regression for $Dst$ time series prediction.
+In the following we consider two choices for the input features $\mathbf{x}_{t-1}$ leading to two variants of \emph{Gaussian Process} regression for $Dst$ time series prediction.
 
@@
 \begin{eqnarray*}
-    \mathbf{x}_t & = & \left(Dst(t), Dst(t-1), \cdots , Dst(t-p+1)\right)
+    \mathbf{x}_{t-1} & = & \left(Dst(t-1), \cdots , Dst(t-p+1)\right)
 \end{eqnarray*}
@@
 
-Auto-regressive models can be improved by including exogenous quantities that can help in improving predictions.
+Auto-regressive models can be augmented by including exogenous quantities in the inputs $\mathbf{x}_{t-1}$ at each time step, in order to improve predictive accuracy.
 $Dst$ gives a measure of ring currents, which are modulated by plasma sheet particle injections into the inner magnetosphere during sub-storms.
-Studies have shown that the substorm occurrence rate increases with solar wind velocity (high speed streams) (\citet{Kissinger2011}, \citet{Newell2016}).
-Prolonged southward interplanetary magnetic field (IMF) z-component ($B_z$) is needed for sub-storm to occur (\citet{McPherron1986}).
-An increase in the solar wind electric field, $VB_z$, can increase the dawn-dusk electric field in the magnetotail, which in turn determines the amount of plasma sheet particle that move to the inner magnetosphere (\citet{Friedel2001}).
+Studies have shown that the substorm occurrence rate increases with solar wind velocity (high speed streams) \citep{Kissinger2011,Newell2016}.
+Prolonged southward interplanetary magnetic field (IMF) z-component ($B_z$) is needed for sub-storms to occur \citep{McPherron1986}.
+An increase in the solar wind electric field, $VB_z$, can increase the dawn-dusk electric field in the magnetotail, which in turn determines the amount of plasma sheet particle that move to the inner magnetosphere \citep{Friedel2001}.
 Therefore, our exogenous parameters consist of solar wind velocity, IMF $B_z$, and $VB_z$.
@@
 
-Hourly solar wind velocity $V(t)$ and interplanetary magnetic field $B_z(t)$ recorded by the ACE satellite are used as the exogenous inputs of the model, along with the time history of $Dst$.
 In this model we choose distinct time lags $p$ and $p_{ex}$ for auto-regressive and exogenous variables respectively.
@@
 \begin{eqnarray*}
-        & \mathbf{x}_t & = (Dst(t), Dst(t-1), \cdots , Dst(t-p+1), \\
-        & \ \ \ \ \ & V(t), V(t-1), \cdots, V(t-p_{ex}+1),\\
-        & \ \ \ \ \ & B_{z}(t), B_{z}(t-1), \cdots, B_{z}(t-p_{ex}+1),\\
-        & \ \ \ \ \ & V B_{z}(t), V B_{z}(t-1), \cdots, V B_{z}(t-p_{ex}+1))
+        & \mathbf{x}_{t-1} & = (Dst(t-1), \cdots , Dst(t-p+1), \\
+        & \ \ \ \ \ & V(t-1), \cdots, V(t-p_{ex}+1),\\
+        & \ \ \ \ \ & B_{z}(t-1), \cdots, B_{z}(t-p_{ex}+1),\\
+        & \ \ \ \ \ & V B_{z}(t-1), \cdots, V B_{z}(t-p_{ex}+1))
 \end{eqnarray*}
@@
 
-\citet{Ji2012} compare six \emph{one step ahead} $Dst$ models as listed in table \ref{table:DstModels} in their paper by compiling a list of 63 geomagnetic storm events of varying strengths which occurred from 1998-2006.
+\citet{Ji2012} compare six \emph{one step ahead} $Dst$ models as listed in table \ref{table:DstModels} in their paper by compiling a list of 63 geomagnetic storm events of varying strengths which occurred in the period 1998-2006.
 This serves as an important stepping stone for a systematic comparison of existing and new prediction techniques because performance metrics are averaged over a large number of storm events that occurred over a 8 year period.
@@
     \begin{equation}
-        RMSE = \sqrt{\sum_{i=1}^{n} (y - \hat{y})^2 / n}
+        RMSE = \sqrt{\sum_{t=1}^{n} (Dst(t) - \hat{D}st(t))^2 / n}
     \end{equation}
@@
     \begin{equation}
-        CC = Cov(y \hat{y})/\sqrt{Var(y) Var(\hat{y})}
+        CC = Cov(Dst, \hat{D}st)/\sqrt{Var(y) Var(\hat{D}st)}
     \end{equation}
@@
     \begin{equation}
-        \Delta Dst_{min} = Dst(t^*) - \hat{D}st(t^*), \ t^* = argmin(Dst(t))
+        \Delta Dst_{min} = Dst(t^*) - \hat{D}st(t^{**}), \ t^* = argmin(Dst(t)),\ t^{**} = argmin(\hat{D}st(t))
     \end{equation}
     \item The error in predicting the timing of a storm peak, also called \emph{timing error} $|\Delta t_{peak}|$.
+    \begin{equation}
+        \Delta t_{peak} = \mathit{argmin}_{t}(D_{st}(t)) - \mathit{argmin}_t(\hat{D}_{st}(t)), \ \ t \in \{1, \cdots, n\}
+    \end{equation}
 \end{enumerate}
@@
 Apart from the metrics above, we also generate the empirical distribution of relative errors $|\frac{Dst - \hat{Dst}}{Dst}|$ for the \emph{GP-AR}, \emph{GP-ARX}, \emph{NM} and \emph{TL} models.
+
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Compare_CC.png}
+      \caption{Averaged cross correlation performance of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model versus results in \citet{Ji2012}.}
+         \label{fig:cc}
+\end{figure}
 
@@
 
-\begin{figure}
-   \centering
-   \includegraphics[width=\textwidth]{Compare_CC.png}
-      \caption{Averaged cross correlation performance of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model versus results in \citet{Ji2012}.}
-         \label{fig:cc}
-\end{figure}
-
 With respect to the \emph{TL} model one important caveat must be noted.
 The training of the \emph{TL} model was performed on the OMNI data from 1996-2002 which overlaps with a large number of the storm events, therefore the experimental procedure has a strong bias towards \emph{TL} model.
-
-\section{Results} \label{sec:res}
-%%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages
 
@@
 
+\section{Results} \label{sec:res}
+%%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages
+
 Figures \ref{fig:rmse}, \ref{fig:cc}, \ref{fig:deltaDst} and \ref{fig:timingErr} compare the performance of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model with the existing results of \citet{Ji2012}.
 Figure \ref{fig:rmse} compares the average RMSE of the predictions over the list of 63 storms.
-One can see that the \emph{GP-ARX} model gives an improvement of approximately $38\%$ in RMSE with respect to the \emph{TL} model and $60\%$ improvement with respect to the \emph{NM} model.
+One can see that the \emph{GP-ARX} model gives an improvement of approximately $40\%$ in RMSE with respect to the \emph{TL} model and $62\%$ improvement with respect to the \emph{NM} model.
 It can also be seen that the \emph{Persistence} model gives better \emph{RMSE} performance than the models compared in \citet{Ji2012}.
@@
 From the results of \citet{Ji2012}, the \emph{TL} model gives the highest correlation coefficient of $94\%$, but that is not surprising considering the fact that there is a large overlap between the training data used to train it and storm events.
-This is also the case in \ref{fig:relprob} where we average predictions generated from a pre-trained instance of the \emph{TL} model.
 Even so the \emph{GP-ARX} model gives a comparable correlation coefficient to \emph{TL}, although the training set of the \emph{GP-ARX} model has no overlap with the storm events which is not the case for \emph{TL}.
@@
 
-Figure \ref{fig:deltaDst} shows how the different models compare with respect to accuracy in predicting the peak value of $Dst$ during storm events ($\Delta Dst_{min}$ averaged over all storms).
-In the context of this metric the \emph{GP-ARX} model gives a $31\%$ improvement over the \emph{TL} model and $40\%$ over the \emph{NM} model.
-It is trivial to note that by definition, the \emph{Persistence} model ($\hat{Dst}(t) = Dst(t-1)$) will have $\Delta Dst_{min} = 0$ for every storm instance.
+Figure \ref{fig:deltaDst} compares the different models with respect to accuracy in predicting the peak value of $Dst$ during storm events ($\Delta Dst_{min}$ averaged over all storms).
+In the context of this metric the \emph{GP-ARX} model gives a $32\%$ improvement over the \emph{TL} model and $41\%$ over the \emph{NM} model.
+It is trivial to note that by definition, the \emph{Persistence} model ($\hat{Dst}(t) = Dst(t-1)$) will have $\Delta Dst_{min} = 0$ for every storm event.
 
@@
 
-Figure \ref{fig:relprob} compares the empirical cumulative distribution functions of the absolute value of the relative errors across all storm events in the experiments.
-One can see that $95\%$ of the hourly predictions have a relative error smaller than $25\%$ for \emph{GP-ARX}, $50\%$ for \emph{NM} and $60\%$ for \emph{TL}.
-
-In figure \ref{fig:predictions} we show the OSA predictions generated by the \emph{TL}, \emph{NM} and \emph{GP-ARX} models for one event from the list of storms in \citet{Ji2012}.
-This storm is interesting due to its strength ($Dst_{min} = -289 \ nT$) and the presence of two distinct peaks.
-In this particular event the \emph{GP-ARX} model approximates the twin peaks as well as the onset and decay phases quite faithfully, contrary the \emph{NM} and \emph{TL} model predictions.
-The \emph{TL} model recognises the presence of twin peaks but overestimates the second one as well as the decay phase, while the \emph{NM} model is much delayed in the prediction of the initial peak and fails to approximate the time and intensity of the second peak.
-
-As noted in section \ref{sec:inference}, \emph{Gaussian Process} models generate probabilistic predictions instead of point predictions, giving the modeller the ability to specify error bars on the predictions.
-In figure \ref{fig:predictionswitherrorbar} \emph{OSA} predictions along with lower and upper error bars generated by the \emph{GP-ARX} model are depicted for the event shown in figure \ref{fig:predictions}.
-The error bars are calculated at one standard deviation on either side of the mean prediction, using the posterior predictive distribution in equations \ref{eq:posteriormean} and \ref{eq:posteriorcov}.
-We remind that, since the model produces Gaussian output distributions, one standard deviation on either side of the mean corresponds to a confidence interval of $68\%$.
-
 \begin{figure}
@@
 
-\begin{comment}
-\begin{enumerate}
-    \item Persistence Behavior: It is clear that the \emph{persistence} behavior in the $Dst$ values is very strong i.e.
-the trivial predictive model $\hat{Dst}(t) = Dst(t-1)$ gives excellent performance.
-In fact it can be seen in these figures that the \emph{Persistence} model is outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $Dst$.
-Therefore any model that is proposed to tackle the OSA prediction problem for $Dst$ must be compared to and show visible gains over the \emph{Persistence} model.
-    \item Assessment of Gaussian Process Models: The results indicate that there are many benefits to be gained by the application of Gaussian Process models for prediction of $Dst$.
-The \emph{GP-AR} and \emph{GP-ARX} not only outperform the models compared in \citet{Ji2012} but also the \emph{Persistence} model.
-\end{enumerate}
-\end{comment}
+Figure \ref{fig:relprob} compares the empirical cumulative distribution functions of the absolute value of the relative errors across all storm events in the experiments.
+One can see that $95\%$ of the hourly predictions have a relative error smaller than $20\%$ for \emph{GP-ARX}, $50\%$ for \emph{NM} and $60\%$ for \emph{TL}.
+
+In Figure \ref{fig:predictions} we show the OSA predictions generated by the \emph{TL}, \emph{NM} and \emph{GP-ARX} models for one event from the list of storms in \citet{Ji2012}.
+This storm is interesting due to its strength ($Dst_{min} = -289 \ nT$) and the presence of two distinct peaks.
+In this particular event the \emph{GP-ARX} model approximates the twin peaks as well as the onset and decay phases quite faithfully, contrary the \emph{NM} and \emph{TL} model predictions.
+The \emph{TL} model recognises the presence of twin peaks but overestimates the second one as well as the decay phase, while the \emph{NM} model is much delayed in the prediction of the initial peak and fails to approximate the time and intensity of the second peak.
+
+As noted in section \ref{sec:inference}, \emph{Gaussian Process} models generate probabilistic predictions instead of point predictions, giving the modeller the ability to specify error bars on the predictions.
+In Figure \ref{fig:predictionswitherrorbar}, \emph{OSA} predictions along with lower and upper error bars generated by the \emph{GP-ARX} model are depicted for the event shown in Figure \ref{fig:predictions}.
+The error bars are calculated at one standard deviation on either side of the mean prediction, using the posterior predictive distribution in equations (\ref{eq:posteriormean}) and (\ref{eq:posteriorcov}).
+Since the predictive distributions are Gaussian, one standard deviation on either side of the mean corresponds to a confidence interval of $68\%$.
 
@@
 the trivial predictive model $\hat{Dst}(t) = Dst(t-1)$ gives excellent performance according to the metrics chosen.
-In fact it can be seen in figure \ref{sec:res} that the \emph{Persistence} model outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $Dst$.
+In fact it can be seen in Figure \ref{fig:rmse} that the \emph{Persistence} model outperforms every model compared by \citet{Ji2012} in the context of OSA prediction of $Dst$.
 Therefore any model that is proposed to tackle the OSA prediction problem for $Dst$ should be compared to the \emph{Persistence} model and show visible gains above it.
@@
 
-      If one compares the training sets used by the models, one can appreciate that they are able to learn robust models for prediction with relatively small training and validations sets: the training set contains 250 instances, while the validation set contains 432 instances.
+      If one compares the training sets used by all the models, one can appreciate that the models presented here need relatively small training and validations sets: the training set contains 250 instances, while the validation set contains 432 instances.
 On the other hand the \emph{NM} model uses one year (8760 instances) of data to learn the formula outlined in \citet{balikhin:narmax} and the \emph{TL} model uses six years of data (52560 instances) for training.
@@
 
-\subsection{Future Research Directions}
+%\subsection{Future Research Directions}
 
-Multi-step ahead prediction should be the preferred gold standard with respect to forecasting of geomagnetic indices.
+%Multi-step ahead prediction should be the preferred gold standard with respect to forecasting of geomagnetic indices.
 Also important is the usage of more informative performance metrics for evaluating and comparing empirical predictive models.
@@
       We acknowledge use of NASA/GSFC's Space Physics Data Facility's OMNIWeb (or CDAWeb or ftp) service, and OMNI data.
+We also acknowledge the authors of \citet{JGRA:JGRA16300} for their scientific inputs towards understanding the \emph{TL} model and assistance in access of its generated predictions.
 \end{acknowledgements}
@@
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%%   Below are examples of how to use particular document layout features
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%  Examples for figures using graphicx
-%%  A guide "Using Imported Graphics in LaTeX2e"  (Keith Reckdahl)
-%%  is available on a lot of LaTeX public servers or ctan mirrors.
-%%  The file is : epslatex.pdf 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%_____________________________________________________________
-%                 A figure as large as the width of the column
-%-------------------------------------------------------------
-   \begin{figure}
-   \centering
-   \includegraphics[width=\textwidth]{empty.eps}
-      \caption{Vibrational stability equation of state
-               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
-               $>0$ means vibrational stability.
-              }
-         \label{FigVibStab}
-   \end{figure}
-%
-%_____________________________________________________________
-%                                    One column rotated figure
-%-------------------------------------------------------------
-   \begin{figure}
-   \centering
-   \includegraphics[angle=-90,width=3cm]{empty.eps}
-      \caption{Vibrational stability equation of state
-               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
-               $>0$ means vibrational stability.
-              }
-         \label{FigVibStab}
-   \end{figure}
-%
-%_____________________________________________________________
-%                        Figure with caption on the right side 
-%-------------------------------------------------------------
-   \begin{figure}
-   \centering
-   \includegraphics[width=3cm]{empty.eps}
-      \caption{Vibrational stability equation of state
-               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
-               $>0$ means vibrational stability.
-              }
-         \label{FigVibStab}
-   \end{figure}
-%
-%_____________________________________________________________
-%
-%_____________________________________________________________
-%                                Figure with a new BoundingBox 
-%-------------------------------------------------------------
-   \begin{figure}
-   \centering
-   \includegraphics[bb=10 20 100 300,width=3cm,clip]{empty.eps}
-      \caption{Vibrational stability equation of state
-               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
-               $>0$ means vibrational stability.
-              }
-         \label{FigVibStab}
-   \end{figure}
-%
-%_____________________________________________________________
-%
-%_____________________________________________________________
-%                                      The "resizebox" command 
-%-------------------------------------------------------------
-   \begin{figure}
-   \resizebox{\textwidth}{!}
-            {\includegraphics[bb=10 20 100 300,clip]{empty.eps}
-      \caption{Vibrational stability equation of state
-               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
-               $>0$ means vibrational stability.
-              }
-         \label{FigVibStab}
-   \end{figure}
-%
-%______________________________________________________________
-%
-%_____________________________________________________________
-%                                             Simple A&A Table
-%_____________________________________________________________
-%
-\begin{table}
-\caption{Nonlinear Model Results}             % title of Table
-\label{table:1}      % is used to refer this table in the text
-\centering                          % used for centering table
-\begin{tabular}{c c c c}        % centered columns (4 columns)
-\hline\hline                 % inserts double horizontal lines
-HJD & $E$ & Method\#2 & Method\#3 \\    % table heading 
-\hline                        % inserts single horizontal line
-   1 & 50 & $-837$ & 970 \\      % inserting body of the table
-   2 & 47 & 877    & 230 \\
-   3 & 31 & 25     & 415 \\
-   4 & 35 & 144    & 2356 \\
-   5 & 45 & 300    & 556 \\ 
-\hline                                   %inserts single line
-\end{tabular}
-\end{table}
-%
-%_____________________________________________________________
-%                                             Two column Table 
-%_____________________________________________________________
-%
-\begin{table*}
-\caption{Nonlinear Model Results}             
-\label{table:1}      
-\centering          
-\begin{tabular}{c c c c l l l }     % 7 columns 
-\hline\hline       
-                      % To combine 4 columns into a single one 
-HJD & $E$ & Method\#2 & \multicolumn{4}{c}{Method\#3}\\ 
-\hline                    
-   1 & 50 & $-837$ & 970 & 65 & 67 & 78\\  
-   2 & 47 & 877    & 230 & 567& 55 & 78\\
-   3 & 31 & 25     & 415 & 567& 55 & 78\\
-   4 & 35 & 144    & 2356& 567& 55 & 78 \\
-   5 & 45 & 300    & 556 & 567& 55 & 78\\
-\hline                  
-\end{tabular}
-\end{table*}
-%
-%-------------------------------------------------------------
-%                                          Table with notes 
-%-------------------------------------------------------------
-%
-% A single note
-\begin{table}[h]
-\caption{\label{t7}Spectral types and photometry for stars in the
-  region.}
-\centering
-\begin{tabular}{lccc}
-\hline\hline
-Star&Spectral type&RA(J2000)&Dec(J2000)\\
-\hline
-69           &B1\,V     &09 15 54.046 & $-$50 00 26.67\\
-49           &B0.7\,V   &*09 15 54.570& $-$50 00 03.90\\
-LS~1267~(86) &O8\,V     &09 15 52.787&11.07\\
-24.6         &7.58      &1.37 &0.20\\
-\hline
-LS~1262      &B0\,V     &09 15 05.17&11.17\\
-MO 2-119     &B0.5\,V   &09 15 33.7 &11.74\\
-LS~1269      &O8.5\,V   &09 15 56.60&10.85\\
-\hline
-\end{tabular}
-\tablefoot{ The top panel shows likely members of Pismis~11. The second
-panel contains likely members of Alicante~5. The bottom panel
-displays stars outside the clusters.}
-\end{table}
-%
-% More notes
-%
-\begin{table}[h]
-\caption{\label{t7}Spectral types and photometry for stars in the
-  region.}
-\centering
-\begin{tabular}{lccc}
-\hline\hline
-Star&Spectral type&RA(J2000)&Dec(J2000)\\
-\hline
-69           &B1\,V     &09 15 54.046 & $-$50 00 26.67\\
-49           &B0.7\,V   &*09 15 54.570& $-$50 00 03.90\\
-LS~1267~(86) &O8\,V     &09 15 52.787&11.07\tablefootmark{a}\\
-24.6         &7.58\tablefootmark{1}&1.37\tablefootmark{a}   &0.20\tablefootmark{a}\\
-\hline
-LS~1262      &B0\,V     &09 15 05.17&11.17\tablefootmark{b}\\
-MO 2-119     &B0.5\,V   &09 15 33.7 &11.74\tablefootmark{c}\\
-LS~1269      &O8.5\,V   &09 15 56.60&10.85\tablefootmark{d}\\
-\hline
-\end{tabular}
-\tablefoot{ The top panel shows likely members of Pismis~11. The second
-panel contains likely members of Alicante~5. The bottom panel
-displays stars outside the clusters.\\
-\tablefoottext{a}{Photometry for MF13, LS~1267 and HD~80077 from
-Dupont et al.}
-\tablefoottext{b}{Photometry for LS~1262, LS~1269 from
-Durand et al.}
-\tablefoottext{c}{Photometry for MO2-119 from
-Mathieu et al.}
-}
-\end{table}
-%
-%-------------------------------------------------------------
-%                                       Table with references 
-%-------------------------------------------------------------
-%
-\begin{table*}[h]
- \caption[]{\label{nearbylistaa2}List of nearby SNe used in this work.}
-\begin{tabular}{lccc}
- \hline \hline
-  SN name &
-  Epoch &
- Bands &
-  References \\
- &
-  (with respect to $B$ maximum) &
- &
- \\ \hline
-1981B   & 0 & {\it UBV} & 1\\
-1986G   &  $-$3, $-$1, 0, 1, 2 & {\it BV}  & 2\\
-1989B   & $-$5, $-$1, 0, 3, 5 & {\it UBVRI}  & 3, 4\\
-1990N   & 2, 7 & {\it UBVRI}  & 5\\
-1991M   & 3 & {\it VRI}  & 6\\
-\hline
-\noalign{\smallskip}
-\multicolumn{4}{c}{ SNe 91bg-like} \\
-\noalign{\smallskip}
-\hline
-1991bg   & 1, 2 & {\it BVRI}  & 7\\
-1999by   & $-$5, $-$4, $-$3, 3, 4, 5 & {\it UBVRI}  & 8\\
-\hline
-\noalign{\smallskip}
-\multicolumn{4}{c}{ SNe 91T-like} \\
-\noalign{\smallskip}
-\hline
-1991T   & $-$3, 0 & {\it UBVRI}  &  9, 10\\
-2000cx  & $-$3, $-$2, 0, 1, 5 & {\it UBVRI}  & 11\\ %
-\hline
-\end{tabular}
-\tablebib{(1)~\citet{branch83};
-(2) \citet{phillips87}; (3) \citet{barbon90}; (4) \citet{wells94};
-(5) \citet{mazzali93}; (6) \citet{gomez98}; (7) \citet{kirshner93};
-(8) \citet{patat96}; (9) \citet{salvo01}; (10) \citet{branch03};
-(11) \citet{jha99}.
-}
-\end{table}
-%_____________________________________________________________
-%                                 A rotated Table in landscape  
-%  In the preamble, use:   \usepackage{lscape}
-%-------------------------------------------------------------
-\begin{landscape}
-\begin{table*}
-\caption{Summary for ISOCAM sources with mid-IR excess 
-(YSO candidates).}\label{YSOtable}
-\centering
-\begin{tabular}{crrlcl} 
-\hline\hline             
-ISO-L1551 & $F_{6.7}$~[mJy] & $\alpha_{6.7-14.3}$ 
-& YSO type$^{d}$ & Status & Comments\\
-\hline
-  \multicolumn{6}{c}{\it New YSO candidates}\\ % To combine 6 columns into a single one
-\hline
-  1 & 1.56 $\pm$ 0.47 & --    & Class II$^{c}$ & New & Mid\\
-  2 & 0.79:           & 0.97: & Class II ?     & New & \\
-  3 & 4.95 $\pm$ 0.68 & 3.18  & Class II / III & New & \\
-  5 & 1.44 $\pm$ 0.33 & 1.88  & Class II       & New & \\
-\hline
-  \multicolumn{6}{c}{\it Previously known YSOs} \\
-\hline
-  61 & 0.89 $\pm$ 0.58 & 1.77 & Class I & \object{HH 30} & Circumstellar disk\\
-  96 & 38.34 $\pm$ 0.71 & 37.5& Class II& MHO 5          & Spectral type\\
-\hline
-\end{tabular}
-\end{table*}
-\end{landscape}
-%
-%_____________________________________________________________
-%                              Table longer than a single page  
-%  In the preamble, use:              \usepackage{longtable}
-%-------------------------------------------------------------
-%          All long tables have to be placed at the end, after 
-%                                        \end{thebibliography}
-%
-% In the text, at the place where the large table should appear
-% add the command:
-\addtocounter{table}{1}
-% Tables counters will be well numbered.
-%
-\end{thebibliography}
-% If table 2
-\longtab{2}{
-\begin{longtable}{lllrrr}
-\caption{\label{kstars} Sample stars with absolute magnitude}\\
-\hline\hline
-Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
-\hline
-\endfirsthead
-\caption{continued.}\\
-\hline\hline
-Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
-\hline
-\endhead
-\hline
-\endfoot
-%%
-Gl 33    & 6.37 & K2 V & 7.46 & S & 0.043170\\
-Gl 66AB  & 6.26 & K2 V & 8.15 & S & 0.260478\\
-Gl 68    & 5.87 & K1 V & 7.47 & P & 0.026610\\
-         &      &      &      & H & 0.008686\\
-Gl 86 
-\footnote{Source not included in the HRI catalog. See Sect.~5.4.2 for details.}
-         & 5.92 & K0 V & 10.91& S & 0.058230\\
-\end{longtable}
-}% End \longtab
-%
-%_____________________________________________________________
-%                              Table longer than a single page
-%                                             and in landscape 
-%  In the preamble, use:       \usepackage{longtable,lscape}
-%-------------------------------------------------------------
-%          All long tables have to be placed at the end, after
-%                                        \end{thebibliography}
-%
-% In the text, at the place where the large table should appear
-% add the command:                        
-\addtocounter{table}{1} 
-% Tables counters will be well numbered.
-%
-\end{thebibliography}
-% If table 2
-\longtabL{2}{
-\begin{landscape}
-\begin{longtable}{lllrrr}
-\caption{\label{kstars} Sample stars with absolute magnitude}\\
-\hline\hline
-Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
-\hline
-\endfirsthead
-\caption{continued.}\\
-\hline\hline
-Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
-\hline
-\endhead
-\hline
-\endfoot
-%%
-Gl 33    & 6.37 & K2 V & 7.46 & S & 0.043170\\
-Gl 66AB  & 6.26 & K2 V & 8.15 & S & 0.260478\\
-Gl 68    & 5.87 & K1 V & 7.47 & P & 0.026610\\
-         &      &      &      & H & 0.008686\\
-Gl 86
-\footnote{Source not included in the HRI catalog. See Sect.~5.4.2 for details.}
-         & 5.92 & K0 V & 10.91& S & 0.058230\\
-\end{longtable}
-\end{landscape}
-}% End \longtabL
-%
-% Online Material
-%_____________________________________________________________
-%        Online appendices have to be placed at the end, after
-%                                        \end{thebibliography}
-%-------------------------------------------------------------
-\end{thebibliography}
-
-\Online
-
-\begin{appendix} %First online appendix
-\section{Background galaxy number counts and shear noise-levels}
-Because the optical images used in this analysis...
-
-\begin{figure*}
-\centering
-\includegraphics[width=16.4cm,clip]{1787f24.ps}
-\caption{Plotted above...}
-\label{appfig}
-\end{figure*}
-
-Because the optical images...
-\end{appendix}
-
-\begin{appendix} %Second online appendix
-These studies, however, have faced...
-\end{appendix}
-
-\end{document}
-%
-%_____________________________________________________________
-%        Some tables or figures are in the printed version and
-%                      some are only in the electronic version
-%-------------------------------------------------------------
-%
-% Leave all the tables or figures in the text, at their right place 
-% and use the commands \onlfig{}{} and \onltab{}{}. These elements
-% will be automatically placed at the end, in the section
-% Online material.
-
-\documentclass{swsc}
-...
-\begin{document}
-text of the paper...
-\begin{figure*}%f1
-\includegraphics[width=10.9cm]{1787f01.eps}
-\caption{Shown in greyscale is a...}
-\label{cl12301}}
-\end{figure*}
-...
-from the intrinsic ellipticity distribution.
-% Figure 2 available electronically only
-\onlfig{2}{
-\begin{figure*}%f2
-\includegraphics[width=11.6cm]{1787f02.eps}
-\caption {Shown in greyscale...}
-\label{cl1018}
-\end{figure*}
-}
-
-% Figure 3 available electronically only
-\onlfig{3}{
-\begin{figure*}%f3
-\includegraphics[width=11.2cm]{1787f03.eps}
-\caption{Shown in panels...}
-\label{cl1059}
-\end{figure*}
-}
-
-\begin{figure*}%f4
-\includegraphics[width=10.9cm]{1787f04.eps}
-\caption{Shown in greyscale is...}
-\label{cl1232}}
-\end{figure*}
-
-\begin{table}%t1
-\caption{Complexes characterisation.}\label{starbursts}
-\centering
-\begin{tabular}{lccc}
-\hline \hline
-Complex & $F_{60}$ & 8.6 &  No. of  \\
-...
-\hline
-\end{tabular}
-\end{table}
-The second method produces...
-
-% Figure 5 available electronically only
-\onlfig{5}{
-\begin{figure*}%f5
-\includegraphics[width=11.2cm]{1787f05.eps}
-\caption{Shown in panels...}
-\label{cl1238}}
-\end{figure*}
-}
-
-As can be seen, in general the deeper...
-% Table 2 available electronically only
-\onltab{2}{
-\begin{table*}%t2
-\caption{List of the LMC stellar complexes...}\label{Properties}
-\centering
-\begin{tabular}{lccccccccc}
-\hline  \hline
-Stellar & RA & Dec & ...
-...
-\hline
-\end{tabular}
-\end{table*}
-}
-
-% Table 3 available electronically only
-\onltab{3}{
-\begin{table*}%t3
-\caption{List of the derived...}\label{IrasFluxes}
-\centering
-\begin{tabular}{lcccccccccc}
-\hline \hline
-Stellar & $f12$ & $L12$ &...
-...
-\hline
-\end{tabular}
-\end{table*}
-}
-%
-%-------------------------------------------------------------
-%     For the online material, table longer than a single page
-%                 In the preamble, use: \usepackage{longtable}
-%       or for landscape option: \usepackage{longtable,lscape}
-%-------------------------------------------------------------
-\documentclass{swsc}
-\usepackage[varg]{txfonts}
-\usepackage{graphicx}
-\usepackage{longtable}
-
-\begin{document}
-text of the paper
-% Table will be print automatically at the end, in the section Online material.
-\onllongtab{3}{
-\begin{longtable}{lrcrrrrrrrrl}
-\caption{Line data and abundances ...}\\
-\hline
-\hline
-Def & mol & Ion & $\lambda$ & $\chi$ & $\log gf$ & N & e &  rad & $\delta$ & $\delta$ 
-red & References \\
-\hline
-\endfirsthead
-\caption{Continued.} \\
-\hline
-Def & mol & Ion & $\lambda$ & $\chi$ & $\log gf$ & B & C &  rad & $\delta$ & $\delta$ 
-red & References \\
-\hline
-\endhead
-\hline
-\endfoot
-\hline
-\endlastfoot
-A & CH & 1 &3638 & 0.002 & $-$2.551 &  &  &  & $-$150 & 150 &  Jorgensen et al. (1996) \\                    
-\end{longtable}
-}% End onllongtab
-
-% Or for landscape, large table:
-
-\onllongtabL{3}{
-\begin{landscape}
-\begin{longtable}{lrcrrrrrrrrl}
-...
-\end{longtable}
-\end{landscape}
-}% End onllongtabL

### 2016-11-01 swsc.tex
Draft 22 June 2016
Updates from ShareLaTeX
Revision[1] [WIP]
Revision[1]: 2016-11-1

@@
 
-%It is well known that the solar wind and interplanetary magnetic field are important factors that influence the Earth's geomagnetic state (\citet{JGRA:JGRA11775}, \citet{Russell1974}, \citet{JGRA:JGRA17574}).
-
 In this model we choose distinct time lags $p$ and $p_{ex}$ for auto-regressive and exogenous variables respectively.
@@
 
+Although the training and model selection data sets both do not have a geomagnetic storm in them, this would not degrade the performance of \emph{GP-AR} and \emph{GP-ARX} because the linear polynomial kernel describes a non stationary and self similar Gaussian Process.
+This implies that for two events where the time histories of $Dst$, $V$ and $B_z$ are not close to each other but can be expressed as a diagonal rescaling of time histories observed in the training data, the predictive distribution is a linearly rescaled version of the training data $Dst$ distribution.
+
+The computational complexity of calculation of the predictive distribution is $O(N^3)$ as per section \label{sec:inference} this can limit the size of the covariance matrix constructed from the training data.
+Note that this computation overhead is paid for every unique assignment to the model hyper-parameters.
+
 Model selection of \emph{GP-AR} and \emph{GP-ARX} is performed by grid search methodology and the root mean square error (\emph{RMSE}) on the validation set is used to select the values of the model hyper-parameters.
@@
 The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$ are arrived by experimenting with values starting from one and choosing the value of $p$ yielding best performance on the 63 events in the bench mark.
+We do not feel that this approach for choosing the time lags leads to bias towards models which only perform well on the data set of \citet{Ji2012} because Gaussian Processes are not easily susceptible to overfitting for situations when the number of model hyper-parameters is small or leave one out cross-validation is not performed.
 
@@
     \begin{equation}
-        CC = Cov(Dst, \hat{D}st)/\sqrt{Var(y) Var(\hat{D}st)}
+        CC = Cov(Dst, \hat{D}st)/\sqrt{Var(Dst) Var(\hat{D}st)}
     \end{equation}
@@
 \section{Results} \label{sec:res}
-%%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages
 
@@
 
-%\subsection{Future Research Directions}
-
-%Multi-step ahead prediction should be the preferred gold standard with respect to forecasting of geomagnetic indices.
-Also important is the usage of more informative performance metrics for evaluating and comparing empirical predictive models.
-
 \begin{acknowledgements}
@@
 We also acknowledge the authors of \citet{JGRA:JGRA16300} for their scientific inputs towards understanding the \emph{TL} model and assistance in access of its generated predictions.
+Simon Wing acknowledges supports from CWI and NSF Grant AGS-1058456 and NASA Grants (NNX13AE12G, NNX15AJ01G, NNX16AC39G).
 \end{acknowledgements}
@@
 
-\end{linenumbers}
+%\end{linenumbers}
 
-\end{document}
+%\end{document}
 
@@
 
-\begin{thebibliography}{}
-\end{thebibliography}
+%\begin{thebibliography}{}
+%\end{thebibliography}
 

### 2016-11-01 Response-Referee9.tex
Referee Response Documents: [WIP]

\documentclass{article}

\usepackage{graphicx}
\usepackage{txfonts}
\usepackage{subfigure}
\usepackage{epstopdf}
\usepackage{lineno}
\usepackage[authoryear,round]{natbib}
\usepackage[backref]{hyperref}
\usepackage{url}
\usepackage[inline]{enumitem}
\usepackage{verbatim}

\begin{document}
\title{Manuscript Ref: swsc160029}
\maketitle
\section{Revisions}

\section{Response}

\end{document}

### 2016-11-01 Response-Referee9.tex
Revisions[2]

@@
 \usepackage{verbatim}
+\usepackage{csquotes}
+\bibliographystyle{swsc.bst}
 
@@
 \maketitle
-\section{Revisions}
+\section*{Response}
 
-\section{Response}
+\begin{enumerate}
+\item \emph{Comparison with existing Dst models}
+
+\textbf{Remarks}:
+
+The authors would like to bring to attention figures \ref{fig:rmseNew} and \ref{fig:ccNew} in which we test the performance NARMAX (\emph{NM}) model as specified in \cite{balikhin:narmax} on the 63 storms benchmark set of \cite{Ji2012}, but use previously predicted $Dst$ values as suggested by the referee instead of actual $Dst$ data.
+%Beginning from the year $1998$ we generate predictions for the \emph{NM} model till the year $2006$.
+This is denoted by \emph{NM Variant} in the figures. As observed in figures \ref{fig:rmseNew} and \ref{fig:ccNew} this leads to severe degradation of performance.
+
+This leads us to conclude that the following statement from \cite{Ji2012} applies only to the testing of the \emph{TL} model: \blockquote{When run the Dst model program for calculating the prediction value of Dst, we do not use the observed Dst data.
+We use the only previously predicted Dst data.}
+
+Furthermore, the authors do not agree that comparisons should be made only on models that input the same input parameters.
+For practical reasons, one may be interested in comparing models’ predictions, irrespective of the input parameters.
+Space weather users care only about the accuracy of the model’s prediction and should not care what parameters the model inputs.
+
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Compare_RMSE_New.png}
+      \caption{
+      Averaged RMSE performance of \emph{NARMAX Variant} and the 
+      \emph{Persistence} model versus results in \cite{Ji2012}.
+      }
+      \label{fig:rmseNew}
+\end{figure}
+
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Compare_CC_New.png}
+      \caption{Averaged cross correlation performance of \emph{NARMAX Variant} and the 
+      \emph{Persistence} model versus results in \citet{Ji2012}.}
+         \label{fig:ccNew}
+\end{figure}
+
+\item \emph{Differences between quicklook, provisional and final $Dst$}
+
+\textbf{Remarks}: If the differences among final, quicklook, and provisional $Dst$ are the quite days base line values (as the reviewer suggests), then this would not affect the timing of the storms only the magnitude due to the subtraction of different offsets.
+Also, this would not change the evaluation of delta $Dst$, CC, RMSE, etc., provided that we consistently use the same type of Dst (e.g, we use only provisional or quicklook $Dst$s).
+
+\end{enumerate}
+
+\section*{Acknowledgements}
+
+The authors would like to thank the reviewer for dedicating their time to provide constructive and valuable feedback.
+
+\bibliography{swsc}
 

### 2016-11-02 Response-Referees.tex
Added common response document

\documentclass{article}

\usepackage{graphicx}
\usepackage{txfonts}
\usepackage{subfigure}
\usepackage{epstopdf}
\usepackage{lineno}
\usepackage[authoryear,round]{natbib}
\usepackage[backref]{hyperref}
\usepackage{url}
\usepackage[inline]{enumitem}
\usepackage{verbatim}
\usepackage{csquotes}

\usepackage[authoryear,round]{natbib}
\usepackage[backref]{hyperref}
\usepackage{url}
%%    This version assumes using bibtex with the swsc bibliography style file
\bibliographystyle{swsc.bst}

\begin{document}
\title{Manuscript Ref: swsc160029}
\maketitle

\section*{Response to referees}
We thank the referees for their valuable feedback and critical assessment which helped to improve the quality of the article. We have revised the manuscript accordingly. In the following, the referees' comments are in italics, and our replies in normal text.\\

\textbf{Referee \#6}
\begin{enumerate}
\item \emph{Was the size of the training set of N=250 hourly values of Dst for the GP-AR model (and, additionally, V, Bz, VBz for the GP-ARX model) limited by the numerical effort to perform the matrix inversion? If yes, what resources did the inversion require?}

It is certainly true that Gaussian Process models become prohibitively expensive for very large training sets. Their computational complexity scales as $O(N^3)$, and this was commented in the manuscript on lines $161-165$.
However, a size of 250 is still very much below the computational limits of the method and, in our case, it has been decided by studying the performance of the models for increasing training sets. We have noticed that using sets with more than 250 values does not increase the overall performance.

We have made this point explicit on lines...

\item \emph{Both the training and model selection data sets did not include strong geomagnetic storms (Dst was between -30 and +37 in the training set and in a similar range in the model selection set).
What were the reasons for selecting these time periods in 2008 and 2014.
Why wasn't a stronger storm event included in the training or the model selection samples?
What justifies applying the models to stronger storms (the example shown in Figures 6 and 7 has Dst values down to -289nT)?}

Additions were made in lines $229-234$ \blockquote{Although the training and model selection data sets both do not have a geomagnetic storm in them, this would not degrade the performance of \emph{GP-AR} and \emph{GP-ARX} because the linear polynomial kernel describes a non stationary and self similar Gaussian Process. This implies that for two events where the time histories of $Dst$, $V$ and $B_z$ are not close to each other but can be expressed as a diagonal rescaling of time histories observed in the training data, the predictive distribution is a linearly rescaled version of the training data $Dst$ distribution.}

\item \emph{The time lag parameters 'p' for the two models were selected by comparing the model to the 63 storm events. Shouldn't the parameter have been selected using an independent set of storms instead in order not to favor these new models against the other models?}

Additions were made in lines $242-247$ in section 4. \blockquote{The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$. These vales have been decided by experimenting with values starting from one and choosing the value of $p$ yielding best performance on the 63 events in the bench mark. We do not feel that this approach for choosing the time lags leads to bias towards models which only perform well on the data set of \cite{Ji2012} because Gaussian Processes are not easily susceptible to overfitting for situations when the number of model hyper-parameters is small or leave one out cross-validation is not performed.}

\item \emph{Rastaetter et al. please correct the first initial in "A. Glocer".}

Corrections made in lines $437-439$

\end{enumerate}
\textbf{Referee \#9}:

\begin{enumerate}
\item \emph{In comparing their models with the existing six models, the authors claim that all these models are ONE STEP AHEAD models (Line 374 of the manuscript). They quote the results of Ji et al., (2012) for comparison. But when Ji et al. compared the six models, they did not run the six models in real OSA approach. As stated in Ji et al., (2012),  'When run the Dst model program for calculating the prediction value of Dst, we do not use the observed Dst data. We use the only previously predicted Dst data.' This means that all the six models take solar wind parameters exclusively as inputs. Thus their verification results for the six models are not in real ONE-STEP-AHEAD approach. The time advance of their predictions relies on the solar wind observation (or forecast). Such comparison is not convincing, and it is inappropriate to conclude that the new models are superior than all the existing models.}

We believe that the referee has misinterpreted the sentence quoted from Ji et al. (2012), and that all of the six models in that paper are indeed run as one-step-ahead models. Our interpretation of that statement is that it refers only to the TL model. That is the only model that indeed does not use the observed DST data as input. It uses some quantities, called $dst*$, that are calculated and advanced in time, by using observed solar wind quantities only. 
However, we agree that that statement is confusing and that it could lead to the referee's interpretation. We present two strong counter-arguments that should settle the issue.

The first argument is that we do not see any reason or benefit for comparing the models in the way the referee suggests they were compared, that is using the predicted $Dst$ values instead of the real and available $Dst$ data. It is obvious that by doing so the performance of the models would degrade, in ways that are difficult to foresee. The main and only purpose of the Ji et al. paper, however, is to compare the performance of different models. Why would one like to compare less-than-optimal results? That paper would be completely flawed if the authors had made what the referee suggests.

We realize that the first argument is simply based on scientific common sense, and that the referee might not be persuaded. The following is a more scientific and indisputable proof that the referee's suggestion is wrong.
In the Ji et al. paper, NARMAX and TL are the most performing models. Since the referee's suggestion is irrelevant for TL (because it does not use $Dst$ by construction), we have tested it on NARMAX.
We have taken the NARMAX model as specified in Boynton et al. (2011) and re-tested it on the 63 storms, by using the previously predicted $Dst$ values instead of the actual $Dst$ data. This model is denoted as \emph{NM Variant} in
Figures \ref{fig:rmseNew} and \ref{fig:ccNew}. Here, we show the Root Mean Square Error and the Correlation Coefficient (same as Figure 2b and 2a in Ji et al. (2012)). By using predicted $Dst$, the RMSE changes from 23.4 to 54.8 and the CC goes from 0.87 to 0.04 (which means no-correlation). 
Clearly, the values reported in that paper could not have been obtained in the way the referee suggests.

As a final remark to the comment \emph{all the six models take solar wind parameters exclusively as inputs}, and the fact that TL indeed does not use $DST$ as input, we reckon that comparisons should not only be made on models that input the same input parameters. Space weather users care only about the accuracy of the model’s prediction, regardless of what parameters the model inputs. Comparisons are fair as long as the same amount of information is potentially available to all models. However, it is a model designer choice which information to use and which to disregard.

\begin{figure}[h]
   \centering
   \includegraphics[width=\textwidth]{Compare_RMSE_New.png}
      \caption{
      Averaged RMSE performance of \emph{NARMAX Variant} and the 
      \emph{Persistence} model versus results in \cite{Ji2012}.
      }
      \label{fig:rmseNew}
\end{figure}

\begin{figure}
   \centering
   \includegraphics[width=\textwidth]{Compare_CC_New.png}
      \caption{Averaged cross correlation performance of \emph{NARMAX Variant} and the 
      \emph{Persistence} model versus results in \citet{Ji2012}.}
         \label{fig:ccNew}
\end{figure}

\item \emph{We know that Dst data are published by WDC Kyoto. The baseline for the Kyoto Dst is determined by measurements made during statistically quiet days. This results in a sequence of adjustments from “Quicklook” Dst values to “Provisional” (several years delay) and eventually “Final” values (after about 5 years) as the quiet days are determined by hindsight. By checking the WDC kyoto website it can be found that currently final data are available for years before 2011, provisional data for years 2012-2014, and Quicklook Dst after 2015. From literatures it was reported that there were large differences between provisional and final Dst (bigger than average error of some of the existing Dst models). Given this, when applied to real time operations, OSA models utilizing history of observed Dst (thus the Quicklook Dst) as input may introduce large errors. This may be one reason that existing Dst models utilize only solar wind data as input.}

First, it is not true that existing $Dst$ models utilize only solar wind data as input. NARMAX (one of the best performing model) uses the time history of $Dst$. Furthermore, since the differences among final, quicklook, and provisional $Dst$ are the quite days base line values, this does not affect the timing of the storms, but only the magnitude due to the subtraction of different offsets. Also, this would not change the evaluation of delta $Dst$, CC, RMSE, etc., provided that we consistently use the same type of Dst (e.g, we use only provisional or quicklook $Dst$s).

\item \emph{To test the strengths of the Gaussian Processes in predictive modeling, it is FINE to evaluate the model by taking FINAL Dst as input, as the authors did. But in comparing them with existing models to test the strength of a new modeling approach compared with other methods, all these models should be run in the same way (whether or not to input observed Dst). Once the new models outperform existing models when they are run in the same way, we can say that Gaussian Processes show stronger potential in predictive modeling.}

Unless we misunderstand the referee's comments, this is a repetition of point 1. 
\end{enumerate}

\bibliography{swsc}

\end{document}

### 2016-11-08 Response-Referees.tex
Referee #6 changes contd.

@@
 Their computational complexity scales as $O(N^3)$, and this was commented in the manuscript on lines $161-165$.
+
 However, a size of 250 is still very much below the computational limits of the method and, in our case, it has been decided by studying the performance of the models for increasing training sets.
@@
 
-We have made this point explicit on lines...
+We have made this point explicit on lines $235-241$: \blockquote{The computational complexity of calculation of the predictive distribution is $O(N^3)$ as per section \ref{sec:inference}, this can limit the size of the covariance matrix constructed from the training data.
+Note that this computation overhead is paid for every unique assignment to the model hyper-parameters.
+However, our chosen training set has a size of 250 which is still very much below the computational limits of the method and in our case, it has been decided by studying the performance of the models for increasing training sets.
+We have noticed that using training sets with more than 250 values does not increase the overall performance in \emph{OSA} prediction of Dst.
+}
 
+\newpage
 \item \emph{Both the training and model selection data sets did not include strong geomagnetic storms (Dst was between -30 and +37 in the training set and in a similar range in the model selection set).
@@
 
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Model_RMSE_validationStorms}
+      \caption{Averaged RMSE performance of \emph{GP-ARX} models for different values of auto-regressive order $p$}
+         \label{fig:prmse}
+ \end{figure}
+
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Model_corr_validationStorms}
+      \caption{Averaged Correlation Coefficient of \emph{GP-ARX} models for different values of auto-regressive order $p$}
+         \label{fig:pcorr}
+ \end{figure}
+
 \item \emph{The time lag parameters 'p' for the two models were selected by comparing the model to the 63 storm events.
@@
 
-Additions were made in lines $242-247$ in section 4.
-\blockquote{The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$.
-These vales have been decided by experimenting with values starting from one and choosing the value of $p$ yielding best performance on the 63 events in the bench mark.
-We do not feel that this approach for choosing the time lags leads to bias towards models which only perform well on the data set of \cite{Ji2012} because Gaussian Processes are not easily susceptible to overfitting for situations when the number of model hyper-parameters is small or leave one out cross-validation is not performed.}
+Additions were made in lines $246-253$ in section 4.
+\blockquote{The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$ are arrived by experimenting with values starting from one and choosing the value of $p$ yielding best performance on a separate set of storm events during the periods $1995-1997$ and $2011-2014$.
+Figures \ref{fig:prmse} and \ref{fig:pcorr} show the performance of \emph{GP-ARX} models having values of $p={3,4,5,6,7}$.
+We observe that the median \emph{RMSE} and \emph{CC} degrade slightly from $p = 3$ to $p = 7$ and models with $p = 6$ have a much smaller variance in performance making them more robust to outliers.
+The choice of $p = 6$ thus gives a good trade off between model robustness and complexity.
+
+We feel that this approach for choosing the time lags does not lead to bias towards models which only perform well on the data set of \citet{Ji2012} because the storms used have no overlap with it.
+Also, Gaussian Processes are not easily susceptible to overfitting when the number of model hyper-parameters is small or leave one out cross-validation is not performed.
+}
 

### 2016-11-08 swsc.tex
Referee #6 changes contd.

@@
 
-The computational complexity of calculation of the predictive distribution is $O(N^3)$ as per section \label{sec:inference} this can limit the size of the covariance matrix constructed from the training data.
+The computational complexity of calculation of the predictive distribution is $O(N^3)$ as per section \ref{sec:inference}, this can limit the size of the covariance matrix constructed from the training data.
 Note that this computation overhead is paid for every unique assignment to the model hyper-parameters.
+However, our chosen training set has a size of 250 which is still very much below the computational limits of the method and in our case, it has been decided by studying the performance of the models for increasing training sets.
+We have noticed that using training sets with more than 250 values does not increase the overall performance in \emph{OSA} prediction of Dst.
+
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Model_RMSE_validationStorms}
+      \caption{Averaged RMSE performance of \emph{GP-ARX} models for different values of auto-regressive order $p$}
+         \label{fig:prmse}
+ \end{figure}
+
+\begin{figure}
+   \centering
+   \includegraphics[width=\textwidth]{Model_corr_validationStorms}
+      \caption{Averaged Correlation Coefficient of \emph{GP-ARX} models for different values of auto-regressive order $p$}
+         \label{fig:pcorr}
+ \end{figure}
 
@@
 
-The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$ are arrived by experimenting with values starting from one and choosing the value of $p$ yielding best performance on the 63 events in the bench mark.
-We do not feel that this approach for choosing the time lags leads to bias towards models which only perform well on the data set of \citet{Ji2012} because Gaussian Processes are not easily susceptible to overfitting for situations when the number of model hyper-parameters is small or leave one out cross-validation is not performed.
+The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$ are arrived by experimenting with values starting from one and choosing the value of $p$ yielding best performance on a separate set of storm events during the periods $1995-1997$ and $2011-2014$.
+Figures \ref{fig:prmse} and \ref{fig:pcorr} show the performance of \emph{GP-ARX} models having values of $p={3,4,5,6,7}$.
+We observe that the median \emph{RMSE} and \emph{CC} degrade slightly from $p = 3$ to $p = 7$ and models with $p = 6$ have a much smaller variance in performance making them more robust to outliers.
+The choice of $p = 6$ thus gives a good trade off between model robustness and complexity.
+
+We feel that this approach for choosing the time lags does not lead to bias towards models which only perform well on the data set of \citet{Ji2012} because the storms used have no overlap with it.
+Also, Gaussian Processes are not easily susceptible to overfitting when the number of model hyper-parameters is small or leave one out cross-validation is not performed.
 

### 2016-11-10 Response-Referees.tex
Updates from ShareLaTeX

@@
 Additions were made in lines $246-253$ in section 4.
-\blockquote{The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$ are arrived by experimenting with values starting from one and choosing the value of $p$ yielding best performance on a separate set of storm events during the periods $1995-1997$ and $2011-2014$.
-Figures \ref{fig:prmse} and \ref{fig:pcorr} show the performance of \emph{GP-ARX} models having values of $p={3,4,5,6,7}$.
+\blockquote{The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} they are $p=6$, $p_{ex} = 1$.
+These values are arrived at by building \emph{GP-AR} and \emph{GP-ARX} models for different values of $p$ and choosing the model yielding the best performance on a set of storm events during the periods $1995-1997$ and $2011-2014$.
+
+Box plots \ref{fig:prmse} and \ref{fig:pcorr} show the performance of \emph{GP-ARX} models having values of $p={3,4,5,6,7}$.
 We observe that the median \emph{RMSE} and \emph{CC} degrade slightly from $p = 3$ to $p = 7$ and models with $p = 6$ have a much smaller variance in performance making them more robust to outliers.

### 2016-11-10 Letter-Editor.tex
Updates from ShareLaTeX

\documentclass{letter}
\usepackage{hyperref}
\signature{Mandar Chandorkar \\ Enrico Camporeale \\ Simon Wing}
%\address{21 Bridge Street \\ Smallville \\ Dunwich DU3 4WE}
\begin{document}

\begin{letter}{Editor \\ Journal of Space Weather and Space Climate}
\opening{Dear Dr. Consuelo Cid:}

We are submitting the revised version of the manuscript, where all the points raised by the referees have been carefully taken into account.

However, we would like to point out that the main comment by referee \#9 revolves around a paragraph in the cited paper by Ji et al. (2012), that we believe has led the referee to a complete misunderstanding of the approach used in that paper.

Of course, this is an important point to clarify given that our results are based on the tests performed by Ji et al. The referee argues that Ji et al. have tested the models not by using the standard one-hour-ahead approach, but rather by propagating the Dst values predicted by the models. This approach would lead to degraded (i.e. less than optimal) performance, and we do not understand what would be the rationale of doing so, in a paper whose first objective is to compare the accuracy of different models.

In any case, to completely clear up any doubt about this important concern, we have tested the NARMAX model (which is one of the best performing models) using both approaches (real data or model predicted data for Dst). The result is shown in the reply to the referees, and it shows that the Ji et al. paper could not have possibly used the approach suggested by the referee. Indeed, not using the real Dst data severely degrades the performance of NARMAX.
We hope that this would settle the argument and, therefore,  that the paper will be deemed publishable in the current form.

\closing{Yours Faithfully,}

\end{letter}
\end{document}

### 2016-11-10 swsc.tex
Updates from ShareLaTeX

@@
 
-The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$ are arrived by experimenting with values starting from one and choosing the value of $p$ yielding best performance on a separate set of storm events during the periods $1995-1997$ and $2011-2014$.
-Figures \ref{fig:prmse} and \ref{fig:pcorr} show the performance of \emph{GP-ARX} models having values of $p={3,4,5,6,7}$.
+The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} they are $p=6$, $p_{ex} = 1$.
+These values are arrived at by building \emph{GP-AR} and \emph{GP-ARX} models for different values of $p$ and choosing the model yielding the best performance on a set of storm events during the periods $1995-1997$ and $2011-2014$.
+
+Box plots \ref{fig:prmse} and \ref{fig:pcorr} show the performance of \emph{GP-ARX} models having values of $p={3,4,5,6,7}$.
 We observe that the median \emph{RMSE} and \emph{CC} degrade slightly from $p = 3$ to $p = 7$ and models with $p = 6$ have a much smaller variance in performance making them more robust to outliers.

### 2016-11-15 Response-Referees.tex
Revisions Nov 16

@@
 
-We have made this point explicit on lines $235-241$: \blockquote{The computational complexity of calculation of the predictive distribution is $O(N^3)$ as per section \ref{sec:inference}, this can limit the size of the covariance matrix constructed from the training data.
+We have made this point explicit on lines $235-243$: \blockquote{The computational complexity of calculation of the predictive distribution is $O(N^3)$, as discussed in section 2.1.
+This can limit the size of the covariance matrix constructed from the training data.
 Note that this computation overhead is paid for every unique assignment to the model hyper-parameters.
-However, our chosen training set has a size of 250 which is still very much below the computational limits of the method and in our case, it has been decided by studying the performance of the models for increasing training sets.
+However, our chosen training set has a size of 250 which is still very much below the computational limits of the method and in our case.
+Solving equation 13 on a laptop computer takes less than a second for the training set considered in our analysis.
+
+The size of the training set has been decided by studying the performance of the models for increasing training sets.
 We have noticed that using training sets with more than 250 values does not increase the overall performance in \emph{OSA} prediction of Dst.
@@
 
+\newpage
 \item \emph{The time lag parameters 'p' for the two models were selected by comparing the model to the 63 storm events.
@@
 
-Additions were made in lines $246-253$ in section 4.
+Additions were made in lines $248-262$ in section 4.
 \blockquote{The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} they are $p=6$, $p_{ex} = 1$.
@@
 Box plots \ref{fig:prmse} and \ref{fig:pcorr} show the performance of \emph{GP-ARX} models having values of $p={3,4,5,6,7}$.
+In these figures lower and upper ends of the vertical lines denote the minimum and maximum respectively, the horizontal line inside a box represents the median, lower and upper ends of a box represent the first and third quartile respectively while dots if present represent outlier points.
+
 We observe that the median \emph{RMSE} and \emph{CC} degrade slightly from $p = 3$ to $p = 7$ and models with $p = 6$ have a much smaller variance in performance making them more robust to outliers.
@@
 
-We feel that this approach for choosing the time lags does not lead to bias towards models which only perform well on the data set of \citet{Ji2012} because the storms used have no overlap with it.
-Also, Gaussian Processes are not easily susceptible to overfitting when the number of model hyper-parameters is small or leave one out cross-validation is not performed.
+This approach for choosing the time lags does not lead to bias towards models which only perform well on the data set of \citet{Ji2012} because the storms used to choose the model order have no overlap with it.
+Moreover, Gaussian Processes are not easily susceptible to overfitting when the number of model hyper-parameters is small.
 }
@@
 
-Corrections made in lines $437-439$
+Corrections made in lines $452-454$
 
 \end{enumerate}
+
+\newpage
 \textbf{Referee \#9}:
@@
 
-We realize that the first argument is simply based on scientific common sense, and that the referee might not be persuaded.
+We realize that this argument is simply based on scientific common sense, and that the referee might not be persuaded.
 The following is a more scientific and indisputable proof that the referee's suggestion is wrong.
-In the Ji et al. paper, NARMAX and TL are the most performing models. Since the referee's suggestion is irrelevant for TL (because it does not use $Dst$ by construction), we have tested it on NARMAX.
+In the Ji et al. paper, NARMAX and TL are the best performing models. Since the referee's suggestion is irrelevant for TL (because it does not use $Dst$ by construction), we have tested it on NARMAX.
 We have taken the NARMAX model as specified in Boynton et al.

### 2016-11-15 swsc.tex
Revisions Nov 16

@@
    \abstract{
-    We present two models, based on Gaussian Processes, for the \emph{One Step Ahead}(OSA) prediction of the $Dst$ geomagnetic activity index. 
+    We present two models, based on Gaussian Processes, for the \emph{One Step Ahead} (OSA) prediction of the $Dst$ geomagnetic activity index. 
     The models are Auto regressive, with or without eXogenous inputs (GP-ARX and GP-AR, respectively).
@@
 The NARMAX methodology builds models by constructing polynomial expansions of inputs and determines the best combinations of monomials to include in the refined model by using a criterion called the \emph{error reduction ratio} (ERR).
-The parameters of the so called NARMAX OLS-ERR model are calculated by solving the \emph{ordinary least squares} (OLS) problem arising from a quadratic objective objective function.
+The parameters of the so called NARMAX OLS-ERR model are calculated by solving the \emph{ordinary least squares} (OLS) problem arising from a quadratic objective function.
 The reader may refer to \citet{billings2013nonlinear} for a detailed exposition of the NARMAX methodology.
@@
 In $Dst$ prediction, a seemingly trivial yet highly informative prediction method is represented by the so called \emph{Persistence} model, which uses the previous value of $Dst$ as the prediction for the next time step ($\hat{D}st(t) = Dst(t-1)$).
-Due to high correlation between $Dst$ values one hour apart ($Dst(t)$, $Dst(t-1)$), the \emph{Persistence} model gives excellent predictive performance on error metrics, despite its apparent lack of sophistication.
+Due to high correlation between $Dst$ values one hour apart ($Dst(t)$, $Dst(t-1)$), the \emph{Persistence} model gives excellent predictive performance on error metrics, despite its lack of sophistication.
 In essence it is a trivial hypothesis and hence it should always be used as a base line to compare the performance of any proposed $Dst$ algorithm.
@@
 Section \ref{sec:method} gives an introduction to the \emph{Gaussian Process} methodology for regression and some technical details about their application.
-In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res}, bench marked against the \emph{Persistence} as well as the other models outlined in \citet{Ji2012}.
+In section \ref{sec:osa} two different \emph{Gaussian Process} auto-regressive models are proposed and subsequently in sections \ref{sec:exp} and \ref{sec:res}, benchmarked against the \emph{Persistence} as well as the other models outlined in \citet{Ji2012}.
 
@@
 With the multivariate normal distribution defined in equation (\ref{eq:dist}), probabilistic predictions $f_*$ can be generated by constructing the conditional distribution $\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}$.
-Since the original distribution of $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right) | \ \ \mathbf{X}, \mathbf{X}_*$ is a multivariate Gaussian, conditioning on a subset of elements: $\mathbf{y}$ yields another Gaussian distribution whose mean and covariance can be calculated exactly, as in equation (\ref{eq:posterior}) (see \citet{Rasmussen:2005:GPM:1162254}).
+Since the original distribution of $\left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right) | \ \ \mathbf{X}, \mathbf{X}_*$ is a multivariate Gaussian, conditioning on a subset of elements $\mathbf{y}$ yields another Gaussian distribution whose mean and covariance can be calculated exactly, as in equation (\ref{eq:posterior}) (see \citet{Rasmussen:2005:GPM:1162254}).
 \begin{equation}
@@
 The practical implementation of \emph{Gaussian Process} models requires the inversion of the training data kernel matrix $[\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1}$ to calculate the parameters of the predictive distribution $\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}$.
-The computational complexity of this inference is dominated by the linear problem $\mathbf{K}^T_{*} [\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y}$, which can be solved via Cholesky decomposition, with a time complexity of $O(N^3)$, where $N$ is the number of data points.
+The computational complexity of this inference is dominated by the linear problem in Eq.
+(\ref{eq:posteriormean}), which can be solved via Cholesky decomposition, with a time complexity of $O(N^3)$, where $N$ is the number of data points.
 
@@
 Below in equations (\ref{eq:Dst}) - (\ref{eq:GPPoly}) we outline a \emph{Gaussian Process} formulation for \emph{OSA} prediction of $Dst$.
-A vector of features $x_{t-1}$ is used as input to an unknown function $f(x_{t-1})$.
+A vector of features $\mathbf{x}_{t-1}$ is used as input to an unknown function $f(\mathbf{x}_{t-1})$.
 
-The features $x_{t-1}$ can be any collection of quantities in the hourly resolution OMNI data set.
-Generally $x_{t-1}$ are time histories of $Dst$ and other important variables such as plasma pressure $p(t)$, solar wind speed $V(t)$, z component of the interplanetary magnetic field $B_z(t)$.
+The features $\mathbf{x}_{t-1}$ can be any collection of quantities in the hourly resolution OMNI data set.
+Generally $\mathbf{x}_{t-1}$ are time histories of $Dst$ and other important variables such as plasma pressure $p(t)$, solar wind speed $V(t)$, $z$ component of the interplanetary magnetic field $B_z(t)$.
 
@@
 Studies have shown that the substorm occurrence rate increases with solar wind velocity (high speed streams) \citep{Kissinger2011,Newell2016}.
-Prolonged southward interplanetary magnetic field (IMF) z-component ($B_z$) is needed for sub-storms to occur \citep{McPherron1986}.
+Prolonged southward interplanetary magnetic field (IMF) $z$-component ($B_z$) is needed for sub-storms to occur \citep{McPherron1986}.
 An increase in the solar wind electric field, $VB_z$, can increase the dawn-dusk electric field in the magnetotail, which in turn determines the amount of plasma sheet particle that move to the inner magnetosphere \citep{Friedel2001}.
@@
 
-The computational complexity of calculation of the predictive distribution is $O(N^3)$ as per section \ref{sec:inference}, this can limit the size of the covariance matrix constructed from the training data.
+The computational complexity of calculation of the predictive distribution is $O(N^3)$, as discussed in section \ref{sec:inference}.
+This can limit the size of the covariance matrix constructed from the training data.
 Note that this computation overhead is paid for every unique assignment to the model hyper-parameters.
-However, our chosen training set has a size of 250 which is still very much below the computational limits of the method and in our case, it has been decided by studying the performance of the models for increasing training sets.
+However, our chosen training set has a size of 250 which is still very much below the computational limits of the method and in our case.
+Solving equation \ref{eq:posteriormean} on a laptop computer takes less than a second for the training set considered in our analysis.
+
+The size of the training set has been decided by studying the performance of the models for increasing training sets.
 We have noticed that using training sets with more than 250 values does not increase the overall performance in \emph{OSA} prediction of Dst.
@@
 
-We feel that this approach for choosing the time lags does not lead to bias towards models which only perform well on the data set of \citet{Ji2012} because the storms used have no overlap with it.
-Also, Gaussian Processes are not easily susceptible to overfitting when the number of model hyper-parameters is small or leave one out cross-validation is not performed.
+This approach for choosing the time lags does not lead to bias towards models which only perform well on the data set of \citet{Ji2012} because the storms used to choose the model order have no overlap with it.
+Moreover, Gaussian Processes are not easily susceptible to overfitting when the number of model hyper-parameters is small.
 
@@
 With respect to the \emph{TL} model one important caveat must be noted.
-The training of the \emph{TL} model was performed on the OMNI data from 1996-2002 which overlaps with a large number of the storm events, therefore the experimental procedure has a strong bias towards \emph{TL} model.
+The training of the \emph{TL} model was performed on data from 1996-2002 which overlaps with a large number of the storm events, therefore the experimental procedure has a strong bias towards \emph{TL} model.
 
@@
 From the results of \citet{Ji2012}, the \emph{TL} model gives the highest correlation coefficient of $94\%$, but that is not surprising considering the fact that there is a large overlap between the training data used to train it and storm events.
-Even so the \emph{GP-ARX} model gives a comparable correlation coefficient to \emph{TL}, although the training set of the \emph{GP-ARX} model has no overlap with the storm events which is not the case for \emph{TL}.
+Even so the \emph{GP-ARX} model gives a comparable correlation coefficient to \emph{TL}, although the training set of the \emph{GP-ARX} model has no overlap with the storm events, which is not the case for \emph{TL}.
 
